### 2014-09-23 usability-study/usability-study.tex
work on paper

\documentclass[a4paper, 11pt]{article}

\begin{document}

\title{Usability Metrics for Mobile Map Applications}

\author{Rachel Rivera\\
  Loyola Marymount University}

\date{September 25, 2014}

% Create title page with no page number

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

%\singlespacing

\maketitle

\vspace{-.2in}
\begin{abstract}
\noindent This investigation aimed to record some usability metrics for two mobile map applications as well as explore \textit{why} these applications performed the way that they did. The two map applications examined were Google Maps\footnote{https://maps.google.com} and Apple Maps\footnote{https://www.apple.com/ios/maps}. Users that took part in this study were given iPhone devices and told to execute three concrete tasks with each application; measurements were taken throughout the execution. Usability metrics were then used in order to make a judgment call on which map application performed best.
\end{abstract}

\medskip
\medskip

\noindent \textit{Professor}: Dr. Dionisio

\medskip
\noindent \textit{Group Members}: Katarina Klask, J.B. Morris, Alex Schneider, Khalid Seirafi, and Ronald Uy

\thispagestyle{empty}

\clearpage

%\onehalfspacing
\setcounter{footnote}{0}
\renewcommand{\thefootnote}{\arabic{footnote}}
\setcounter{page}{1}

\section{Introduction}

Interaction design is a field of study that has grown and matured a considerable amount in recent years. The current consensus of the field measures the performance of systems based on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\footnote[1]{\textit{Vocabulary specifically from Jakob Nielsen}: Nielsen, Jakob. Usability Engineering. Boston: Academic, 1993. Print.} Talkin more about stuff.
\section{The Model} \label{sec:Model}

Model for testing users.

\subsection{Results}
blah blah blah
\begin{table}[ht]
\centering 
\begin{tabular}{l c c c c c} % The final bracket specifies the number of columns in the table along with left and right borders which are specified using vertical bars (|); each column can be left, right or center-justified using l, r or c. To specify a precise width, use p{width}, e.g. p{5cm}
& \multicolumn{5}{c}{Growth Media} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
Strain & 1 & 2 & 3 & 4 & 5\\ % Column names row
GDS1002 & 0.962 & 0.821 & 0.356 & 0.682 & 0.801\\ % Content row 1
NWN652 & 0.981 & 0.891 & 0.527 & 0.574 & 0.984\\ % Content row 2
PPD234 & 0.915 & 0.936 & 0.491 & 0.276 & 0.965\\ % Content row 3
JSB126 & 0.828 & 0.827 & 0.528 & 0.518 & 0.926\\ % Content row 4
JSB724 & 0.916 & 0.933 & 0.482 & 0.644 & 0.937\\ % Content row 5

Average Rate & 0.920 & 0.882 & 0.477 & 0.539 & 0.923\\ % Summary/total row

\end{tabular}
\caption{Table caption text}
\end{table}

\end{document}

### 2014-09-23 usability-study/usability-study.tex
practicing with latex

@@
-\documentclass[a4paper, 11pt]{article}
-
+\documentclass[11pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage{graphicx}
+\usepackage{multicol}
+\setlength{\columnsep}{1cm}
+\bibliographystyle{unsrt}
 \begin{document}
@@
 \noindent This investigation aimed to record some usability metrics for two mobile map applications as well as explore \textit{why} these applications performed the way that they did.
-The two map applications examined were Google Maps\footnote{https://maps.google.com} and Apple Maps\footnote{https://www.apple.com/ios/maps}.
+The two map applications examined were Google Maps \cite{Pan} and Apple Maps \footnote[2]{https://www.apple.com/ios/maps}.
 Users that took part in this study were given iPhone devices and told to execute three concrete tasks with each application; measurements were taken throughout the execution.
@@
 
+When Apple Maps was first released as part of iOS6 in September 2012, they received a signifcant amount of negative feedback.
+The Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people in September 2012.
+However, Apple Maps slowly began gaining traction and by September 2013, the numer of users of Google Maps dropped to around 58.7 million.
 Interaction design is a field of study that has grown and matured a considerable amount in recent years.
-The current consensus of the field measures the performance of systems based on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\footnote[1]{\textit{Vocabulary specifically from Jakob Nielsen}: Nielsen, Jakob.
-Usability Engineering.
-Boston: Academic, 1993.
-Print.} Talkin more about stuff.
+The current consensus of the field measures the performance of systems based on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen}
 \section{The Model} \label{sec:Model}
@@
 
+\subsection{How to Include Figures}
+
+First you have to upload the image file (JPEG, PNG or PDF) from your computer to writeLaTeX using the upload link the project menu.
+Then use the includegraphics command to include it in your document.
+Use the figure environment and the caption command to add a number and a caption to your figure.
+See the code for Figure \ref{fig:frog} in this section for an example.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.3\textwidth]{frog.jpg}
+\caption{\label{fig:frog}This frog was uploaded to writeLaTeX via the project menu.}
+\end{figure}
+
 \subsection{Results}
 blah blah blah
+
 \begin{table}[ht]
@@
 & \multicolumn{5}{c}{Growth Media} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
-Strain & 1 & 2 & 3 & 4 & 5\\ % Column names row
+Strain & 11 & 2 & 3 & 4 & 5\\ % Column names row
 GDS1002 & 0.962 & 0.821 & 0.356 & 0.682 & 0.801\\ % Content row 1
@@
 
+\begin{table}
+\centering
+\begin{tabular}{l|r}
+Item & Quantity \\\hline
+Widgets & 42 \\
+Gadgets & 13
+\end{tabular}
+\caption{\label{tab:widgets}An example table.}
+\end{table}
+
+\begin{multicols}{2}
+[
+\section{First Section}
+All human things are subject to decay. And when fate summons, Monarchs must obey.
+]
+Hello, here is some text without a meaning.  This text should show what 
+a printed text will look like at this place.
+If you read this text, you will get no information.  Really?  Is there 
+no information?  Is there...
+\end{multicols}
+\clearpage
+\begin{thebibliography}{100} % 100 is a random guess of the total number of 
+%references
+\bibitem{Nielsen} Nielsen, Jakob, \emph{Usability Engineering}, Boston: Academic, 1993.
+\bibitem{Boney96} Boney, L., Tewfik, A.H., and Hamdy, K.N., ``Digital 
+Watermarks for Audio Signals," \emph{Proceedings of the Third IEEE 
+International Conference on Multimedia}, pp. 473-480, June 1996.
+\bibitem{MG} Goossens, M., Mittelbach, F., Samarin, \emph{A LaTeX 
+Companion}, Addison-Wesley, Reading, MA, 1994.
+\bibitem{HK} Kopka, H., Daly P.W., \emph{A Guide to LaTeX},
+Addison-Wesley, Reading, MA, 1999.
+\bibitem{Pan} Pan, D., ``A Tutorial on MPEG/Audio Compression," \emph{IEEE 
+Multimedia}, Vol.2, pp.60-74, Summer 1998.
+\end{thebibliography}
+%%%%%%%%%%%%% end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2014-09-24 usability-study/usability-study.tex
completed abstract and intro

@@
-\documentclass[11pt]{article}
+\documentclass[a4paper; 11pt]{article}
 \usepackage[utf8]{inputenc}
@@
 
-\title{Usability Metrics for Mobile Map Applications}
+\title{Usability of Mobile Map Applications}
 
 \author{Rachel Rivera\\
+Professor: Dr. Dionisio\\
+CMSI 370: Interaction Design\\
   Loyola Marymount University}
@@
 \vspace{-.2in}
+\begin{center}
 \begin{abstract}
-\noindent This investigation aimed to record some usability metrics for two mobile map applications as well as explore \textit{why} these applications performed the way that they did.
-The two map applications examined were Google Maps \cite{Pan} and Apple Maps \footnote[2]{https://www.apple.com/ios/maps}.
-Users that took part in this study were given iPhone devices and told to execute three concrete tasks with each application; measurements were taken throughout the execution.
-Usability metrics were then used in order to make a judgment call on which map application performed best.
+\noindent This investigation is motivated by the increasing significance of the usability of systems in today's competitive marketplace.
+Usability testing and evaluations have become important tools in the assessment of user interfaces.
+This paper presents empirical mobile usability studies of two distinct map applications: Google Maps and Apple Maps.
+Participants of this investigation were asked to carry out three tasks with each application.
+The results of this examination include (a) the contextual factors studied; (b) the core usability metrics measured; (c) quantitative data collected from each task; and (d) a brief qualitative review from each participant.
+This paper not only presents the usability metrics collected by the study but also presents a heuristic evaluation of each map application.
 \end{abstract}
+\end{center}
 
@@
 
-\noindent \textit{Professor}: Dr. Dionisio
-
 \medskip
-\noindent \textit{Group Members}: Katarina Klask, J.B. Morris, Alex Schneider, Khalid Seirafi, and Ronald Uy
+\noindent \textit{Group Members}: Katarina Klask, J.B. Morris, Alex Schneider, Khalid Seirafi, and Ronald Uy.
 
@@
 \section{Introduction}
-
-When Apple Maps was first released as part of iOS6 in September 2012, they received a signifcant amount of negative feedback.
-The Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people in September 2012.
-However, Apple Maps slowly began gaining traction and by September 2013, the numer of users of Google Maps dropped to around 58.7 million.
-Interaction design is a field of study that has grown and matured a considerable amount in recent years.
-The current consensus of the field measures the performance of systems based on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen}
+When the Apple Maps mobile application was first released as part of iOS 6 in September 2012, it received a significant amount of negative feedback.
+Walt Mossberg of the \textit{Wall Street Journal} was one of many who thought that Apple Maps was the biggest drawback of the iPhone 5.\cite{Mossberg} At this time, the Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people.\cite{ComScore} However, Apple Maps grew, matured, and began gaining traction.
+By September 2013, Apple Maps gained 35 million regular users.
+The number of users of Google Maps dropped to around 58.7 million at this time.\cite{ComScore} Although the Apple Maps mobile app is newer than the Google Maps mobile app,\footnote{The initial release of Google Maps for Mobile was in September of 2008.} the two map applications are comparable with respect to features they provide as well as popularity nowadays.
+\medskip
+\par
+Since both Google Maps apps and Apple Maps apps are widely known and used, it is important that the usability metrics of the two systems are examined.
+Thus, this study aims to empirically investigate the usability of each application.
+The current consensus within the field of interaction design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics that this particular study focuses on are efficiency, errors, and satisfaction.
 \section{The Model} \label{sec:Model}
@@
 \subsection{How to Include Figures}
-
-First you have to upload the image file (JPEG, PNG or PDF) from your computer to writeLaTeX using the upload link the project menu.
-Then use the includegraphics command to include it in your document.
-Use the figure environment and the caption command to add a number and a caption to your figure.
-See the code for Figure \ref{fig:frog} in this section for an example.
-
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 \bibitem{Nielsen} Nielsen, Jakob, \emph{Usability Engineering}, Boston: Academic, 1993.
-\bibitem{Boney96} Boney, L., Tewfik, A.H., and Hamdy, K.N., ``Digital 
-Watermarks for Audio Signals," \emph{Proceedings of the Third IEEE 
-International Conference on Multimedia}, pp. 473-480, June 1996.
-\bibitem{MG} Goossens, M., Mittelbach, F., Samarin, \emph{A LaTeX 
-Companion}, Addison-Wesley, Reading, MA, 1994.
+\bibitem{Mossberg} Mossberg, Walt., ``The iPhone Takes to the Big Screen," \emph{The Wall Street Journal}, September 2012.
+\bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore,Inc}, 2012.
 \bibitem{HK} Kopka, H., Daly P.W., \emph{A Guide to LaTeX},

### 2014-09-24 usability-study/usability-study.tex
adding graphs for results

@@
 \usepackage{multicol}
+\makeatletter
+\setlength{\@fptop}{0pt}
+\makeatother
 \setlength{\columnsep}{1cm}
@@
 \begin{abstract}
-\noindent This investigation is motivated by the increasing significance of the usability of systems in today's competitive marketplace.
-Usability testing and evaluations have become important tools in the assessment of user interfaces.
+\noindent This investigation is motivated by the increasing significance of the usability of systems in today's competitive technology market.
+Usability testing and usability evaluations have become important tools in the assessment of user interfaces.
 This paper presents empirical mobile usability studies of two distinct map applications: Google Maps and Apple Maps.
-Participants of this investigation were asked to carry out three tasks with each application.
+Participants of this investigation were asked to carry out three tasks with each mobile application.
 The results of this examination include (a) the contextual factors studied; (b) the core usability metrics measured; (c) quantitative data collected from each task; and (d) a brief qualitative review from each participant.
@@
 The current consensus within the field of interaction design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics that this particular study focuses on are efficiency, errors, and satisfaction.
-\section{The Model} \label{sec:Model}
+\section{The Experiment} \label{sec:Experiment}
+Each person who participated in this study had to carry out three concrete tasks with each of the map applications.
+\begin{enumerate}
+  \item Save a pin anywhere on the map and get driving directions to that pin.
+  \item Find the location of the nearest Vons.
+  \item Find the fastest route from the current location to LACMA.
+\end{enumerate}
+In order to keep the study as controlled as possible, all testing was done on iOS devices (no map applications of android devices were tested).
+As test subjects executed the tasks, information was collected pertaining to their efficiency, errors, and satisfaction.
 
-Model for testing users.
+\subsection{Contextual Factors}
+Prior to having each test subject perform the tasks, a little contextual data about the users was first collected.
+Users were asked about their previous experience with Apple Maps, their previous experience with Google Maps, and their overall technological proficiency.
+\clearpage
 
-\subsection{How to Include Figures}
+One problem that arose when conducting this study was that almost every test subject had significantly more Google Maps experience than Apple Maps experience.
+This is not necessarily surprising since Apple Maps is a newer application.
+However, this study was aiming to test efficiency, and not learnability.
+Thus if a test subject did not have any experience with an application, the test subject would not be asked to execute tasks on that application.
+
 \begin{figure}[ht]
-\centering
-\includegraphics[width=0.3\textwidth]{frog.jpg}
-\caption{\label{fig:frog}This frog was uploaded to writeLaTeX via the project menu.}
+\includegraphics[keepaspectratio, width=1\textwidth ]{user_context.pdf}
 \end{figure}
@@
 \subsection{Results}
-blah blah blah
+\textit{The design of system icons is simple, modern, friendly, and sometimes quirky.
+Each icon is reduced to its minimal form, with every idea edited to its essence.
+The designs ensure readability and clarity even at small sizes}
+\begin{figure}[ht]
+\includegraphics[keepaspectratio, width=1\textwidth ]{task1.pdf}
+\end{figure}
 
@@
 \centering 
-\begin{tabular}{l c c c c c} % The final bracket specifies the number of columns in the table along with left and right borders which are specified using vertical bars (|); each column can be left, right or center-justified using l, r or c.
+\begin{tabular}{l c c c c c c c} % The final bracket specifies the number of columns in the table along with left and right borders which are specified using vertical bars (|); each column can be left, right or center-justified using l, r or c.
 To specify a precise width, use p{width}, e.g.
-p{5cm}
-& \multicolumn{5}{c}{Growth Media} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
-Strain & 11 & 2 & 3 & 4 & 5\\ % Column names row
-GDS1002 & 0.962 & 0.821 & 0.356 & 0.682 & 0.801\\ % Content row 1
+p{8cm}
+& \multicolumn{5}{c}{Map Application Tasks} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
+& Task1 & Error1 & Task2 & Error2 & Task3 & Error3 \\
+User1 & 0.962 & 0.821 & 0.356 & 0.682 & 0.801\\ % Content row 1
 NWN652 & 0.981 & 0.891 & 0.527 & 0.574 & 0.984\\ % Content row 2

### 2014-09-25 usability-study/usability-study.tex
new task graphs

@@
 Participants of this investigation were asked to carry out three tasks with each mobile application.
-The results of this examination include (a) the contextual factors studied; (b) the core usability metrics measured; (c) quantitative data collected from each task; and (d) a brief qualitative review from each participant.
+The results of this examination include (a) the contextual factors studied; (b) quantitative data collected from each task; (c) a brief qualitative review from each participant; and (d) the core usability metrics measured.
 This paper not only presents the usability metrics collected by the study but also presents a heuristic evaluation of each map application.
@@
 \end{figure}
+\begin{figure}[ht]
+\includegraphics[keepaspectratio, width=1\textwidth ]{task2.pdf}
+\end{figure}
+\begin{figure}[ht]
+\includegraphics[keepaspectratio, width=1\textwidth ]{task3.pdf}
+\end{figure}
 

### 2014-09-25 usability-study/usability-study.tex

--- 2014-09-24 usability-study/usability-study.tex
@@
 This paper presents empirical mobile usability studies of two distinct map applications: Google Maps and Apple Maps.
-Participants of this investigation were asked to carry out three tasks with each mobile application.
-The results of this examination include (a) the contextual factors studied; (b) the core usability metrics measured; (c) quantitative data collected from each task; and (d) a brief qualitative review from each participant.
+Participants of this investigation were asked to carry out three tasks with each mobile map application.
+The results of this examination include (a) the contextual factors studied; (b) quantitative data collected from each task; (c) a brief qualitative review of the applications from each participant; and (d) the core usability metrics measured.
 This paper not only presents the usability metrics collected by the study but also presents a heuristic evaluation of each map application.
@@
 The current consensus within the field of interaction design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics that this particular study focuses on are efficiency, errors, and satisfaction.
+\section{Contextual Factors}
+Prior to having each test subject perform the tasks, a little contextual data about the users was first collected.
+Users were asked about their previous experience with Apple Maps, their previous experience with Google Maps, and their overall technological proficiency.
+\vspace{-.2in}
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[keepaspectratio, width=.7\textwidth ]{user_context.pdf}
+\end{center}
+\end{figure}
+\clearpage
+\par
+The fact that nearly every test subject had significantly more Google Maps experience than Apple Maps experience was a challenge in the conducting of this study.
+Since this study aimed to test the usability metric of efficiency, and not learnability, we did not test subjects if they had no prior experience with the application.
+For future studies, however, it might be worthwhile to give inexperienced users a short tutorial and then still test them.
+
 \section{The Experiment} \label{sec:Experiment}
@@
 
-\subsection{Contextual Factors}
-Prior to having each test subject perform the tasks, a little contextual data about the users was first collected.
-Users were asked about their previous experience with Apple Maps, their previous experience with Google Maps, and their overall technological proficiency.
-\clearpage
+\section{Usability Metrics}
+\subsection{Efficiency}
+\par
+\textbf{First Task: }Drop a pin on the map and get driving directions to that pin.
+\vspace{-.4in}
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[keepaspectratio, width=.8\textwidth ]{task1.pdf}
+\end{center}
+\end{figure}
+\begin{center}
+\par
+Average time for task on Apple Maps app: approx. \textbf{19 seconds}
+\par
+Average time for task on Google Maps app: approx. \textbf{21 seconds}
+\end{center}
+\par
+\noindent
+The results suggest that users are more efficient using Apple Maps.
+\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it is important to note that there were a couple of anomalies within the set of Google Maps users.}
+\medskip
+\medskip
+\par
+\noindent
+\textbf{Second Task: }Find the location of the nearest Vons.
+\begin{figure}[ht]
+\begin{center}
+\vspace{-.5in}
+\includegraphics[keepaspectratio, width=.8\textwidth ]{task2.pdf}
+\end{center}
+\end{figure}
+\begin{center}
+\vspace{-.6in}
+\par
+Average time for task on Apple Maps app: approx. \textbf{24 seconds}
+\par
+Average time for task on Google Maps app: approx. \textbf{21 seconds}
+\end{center}
+\par
+\noindent
+The results suggest that users are more efficient using Google Maps. 
+\medskip
+\medskip
+\par
+\noindent
+\textbf{Third Task: }Find the fastest route from the current location to LACMA.
+\begin{figure}[ht]
+\begin{center}
+\vspace{-.4in}
+\includegraphics[keepaspectratio, width=.8\textwidth ]{task3.pdf}
+\end{center}
+\end{figure}
+\begin{center}
+\vspace{-.4in}
+\par
+Average time for task on Apple Maps app: approx. \textbf{36 seconds}
+\par
+Average time for task on Google Maps app: approx. \textbf{17 seconds}
+\end{center}
+\par
+\noindent
+The results suggest that users are more efficient using Google Maps.
 
-One problem that arose when conducting this study was that almost every test subject had significantly more Google Maps experience than Apple Maps experience.
-This is not necessarily surprising since Apple Maps is a newer application.
-However, this study was aiming to test efficiency, and not learnability.
-Thus if a test subject did not have any experience with an application, the test subject would not be asked to execute tasks on that application.
-
-\begin{figure}[ht]
-\includegraphics[keepaspectratio, width=1\textwidth ]{user_context.pdf}
-\end{figure}
-
-\subsection{Results}
-\textit{The design of system icons is simple, modern, friendly, and sometimes quirky.
-Each icon is reduced to its minimal form, with every idea edited to its essence.
-The designs ensure readability and clarity even at small sizes}
-\begin{figure}[ht]
-\includegraphics[keepaspectratio, width=1\textwidth ]{task1.pdf}
-\end{figure}
+\subsection{Errors}
+\par
+The only errors accounted for here were incidents wherein the user did something whose result is not what he or she anticipated. The following table shows you some examples of errors from this study.
 
@@
 \centering
-\begin{tabular}{l c c c c c c c} % The final bracket specifies the number of columns in the table along with left and right borders which are specified using vertical bars (|); each column can be left, right or center-justified using l, r or c.
-To specify a precise width, use p{width}, e.g.
-p{8cm}
-& \multicolumn{5}{c}{Map Application Tasks} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
-& Task1 & Error1 & Task2 & Error2 & Task3 & Error3 \\
-User1 & 0.962 & 0.821 & 0.356 & 0.682 & 0.801\\ % Content row 1
-NWN652 & 0.981 & 0.891 & 0.527 & 0.574 & 0.984\\ % Content row 2
-PPD234 & 0.915 & 0.936 & 0.491 & 0.276 & 0.965\\ % Content row 3
-JSB126 & 0.828 & 0.827 & 0.528 & 0.518 & 0.926\\ % Content row 4
-JSB724 & 0.916 & 0.933 & 0.482 & 0.644 & 0.937\\ % Content row 5
-
-Average Rate & 0.920 & 0.882 & 0.477 & 0.539 & 0.923\\ % Summary/total row
+\begin{tabular}{l|c|c}
+Task & Application & Specific Error\\\hline
+Task1 & Apple Maps & hit 'Share' button; no response from tapping \\
+Task1 & Google Maps & hit 'Near you' to try to drop pin\\
+Task2 & Apple Maps & tried to manually search for 'Nearest Vons'\\
+Task2 & Google Maps & voice command instead of search\\
+Task3 & Apple Maps & hit 'Current Location' for directions\\
+Task3 & Google Maps & could not find navigate button
+\end{tabular}
+\end{table}
+\par
+The \textit{percentage} of subjects with errors is analyzed rather than solely the numbers of errors since more subjects tested the Google Maps app than Apple Maps app.
+\begin{table}[ht]
+\begin{tabular}{l c c c c}
+& \multicolumn{4}{c}{}\\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
+& \textbf{Task1} & \textbf{Task2} & \textbf{Task3} \\
+Percentage of Apple Users with Errors & 40\% & 50\%  & 20\% \\
+Percentage of Google Users with Errors & 29.4\% & 17.6\%  & 11.7\% \\
 
 \end{tabular}
-\caption{Table caption text}
+
 \end{table}
+\par
+The results suggest that Google Maps is more usable with respect to errors.
 
-\begin{table}
-\centering
-\begin{tabular}{l|r}
-Item & Quantity \\\hline
-Widgets & 42 \\
-Gadgets & 13
-\end{tabular}
-\caption{\label{tab:widgets}An example table.}
-\end{table}
+\subsection{Satisfaction}
+\begin{figure}[ht]
+\begin{center}
+\vspace{-.6in}
+\includegraphics[keepaspectratio, width=.8\textwidth ]{satisfaction.pdf}
+\end{center}
+\end{figure}
+\par
+\vspace{-.4in}
+Satisfaction ratings were on a scale from 1 to 10.
+\par
+Average rating for Apple Maps: approx. \textbf{6.18}
+\par
+Average rating for Google Maps: approx. \textbf{7.72}
 
@@
 [
-\section{First Section}
+\section{Heuristic Evaluation}
 All human things are subject to decay. And when fate summons, Monarchs must obey.
@@
 If you read this text, you will get no information.  Really?  Is there 
-no information?  Is there...
+no information?  Is there
 \end{multicols}

### 2014-09-25 usability-study/usability-study.tex

@@
 \usepackage{multicol}
+\usepackage{amsmath}
 \makeatletter
@@
 The fact that nearly every test subject had significantly more Google Maps experience than Apple Maps experience was a challenge in the conducting of this study.
-Since this study aimed to test the usability metric of efficiency, and not learnability, we did not test subjects if they had no prior experience with the application.
-For future studies, however, it might be worthwhile to give inexperienced users a short tutorial and then still test them.
+Since this study aimed to test the usability metric of efficiency, and not that of learnability, we did not test subjects if they had no prior experience with the application.
+For future studies, however, it might be worthwhile to give inexperienced users a short tutorial so it is possible to test them thereafter.
 
@@
 \begin{center}
+\vspace{-.1in}
 \includegraphics[keepaspectratio, width=.8\textwidth ]{task1.pdf}
@@
 \begin{center}
+\vspace{-.6in}
 \par
-Average time for task on Apple Maps app: approx. \textbf{19 seconds}
+Average time for task \#1 on Apple Maps app: approx. \textbf{19 seconds}
 \par
-Average time for task on Google Maps app: approx. \textbf{21 seconds}
+Average time for task \#1 on Google Maps app: approx. \textbf{21 seconds}
 \end{center}
@@
 \noindent
-The results suggest that users are more efficient using Apple Maps.
+This task's results suggest that users are more efficient using Apple Maps.
 \footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it is important to note that there were a couple of anomalies within the set of Google Maps users.}
@@
 \par
-Average time for task on Apple Maps app: approx. \textbf{24 seconds}
+Average time for task \#2 on Apple Maps app: approx. \textbf{24 seconds}
 \par
-Average time for task on Google Maps app: approx. \textbf{21 seconds}
+Average time for task \#2 on Google Maps app: approx. \textbf{21 seconds}
 \end{center}
@@
 \noindent
-The results suggest that users are more efficient using Google Maps. 
+This task's results suggest that users are more efficient using Google Maps. 
 \medskip
@@
 \par
-Average time for task on Apple Maps app: approx. \textbf{36 seconds}
+Average time for task \#3 on Apple Maps app: approx. \textbf{36 seconds}
 \par
-Average time for task on Google Maps app: approx. \textbf{17 seconds}
+Average time for task \#3 on Google Maps app: approx. \textbf{17 seconds}
 \end{center}
@@
 \noindent
-The results suggest that users are more efficient using Google Maps.
+This task's results suggest that users are more efficient using Google Maps.
 
@@
 \par
-The only errors accounted for here were incidents wherein the user did something whose result is not what he or she anticipated. The following table shows you some examples of errors from this study.
+The only errors collected by this study were incidents wherein the user did something whose result is not what he or she anticipated.
+The table on the following page shows you some examples of errors from this investigation.
 
@@
 Task & Application & Specific Error\\\hline
-Task1 & Apple Maps & hit 'Share' button; no response from tapping \\
-Task1 & Google Maps & hit 'Near you' to try to drop pin\\
-Task2 & Apple Maps & tried to manually search for 'Nearest Vons'\\
+Task1 & Apple Maps & hit \textit{Share} button; no response from tapping \\
+Task1 & Google Maps & hit \textit{Near You} button to try to drop pin\\
+Task2 & Apple Maps & tried to manually search for ``Nearest Vons"\\
 Task2 & Google Maps & voice command instead of search\\
-Task3 & Apple Maps & hit 'Current Location' for directions\\
-Task3 & Google Maps & could not find navigate button
+Task3 & Apple Maps & hit \textit{Current Location} for directions\\
+Task3 & Google Maps & found side streets when expecting highway route
 \end{tabular}

### 2014-09-25 usability-study/usability-study.tex
change bolded results to italicized
heuristic evaluation

@@
 \par
-Average time for task \#1 on Apple Maps app: approx. \textbf{19 seconds}
+Average time for task \#1 on Apple Maps app: approx. \textit{19 seconds}
 \par
-Average time for task \#1 on Google Maps app: approx. \textbf{21 seconds}
+Average time for task \#1 on Google Maps app: approx. \textit{21 seconds}
 \end{center}
@@
 \par
-Average time for task \#2 on Apple Maps app: approx. \textbf{24 seconds}
+Average time for task \#2 on Apple Maps app: approx. \textit{24 seconds}
 \par
-Average time for task \#2 on Google Maps app: approx. \textbf{21 seconds}
+Average time for task \#2 on Google Maps app: approx. \textit{21 seconds}
 \end{center}
@@
 \par
-Average time for task \#3 on Apple Maps app: approx. \textbf{36 seconds}
+Average time for task \#3 on Apple Maps app: approx. \textit{36 seconds}
 \par
-Average time for task \#3 on Google Maps app: approx. \textbf{17 seconds}
+Average time for task \#3 on Google Maps app: approx. \textit{17 seconds}
 \end{center}
@@
 \begin{center}
-\vspace{-.6in}
 \includegraphics[keepaspectratio, width=.8\textwidth ]{satisfaction.pdf}
@@
 \par
-\vspace{-.4in}
 Satisfaction ratings were on a scale from 1 to 10.
 \par
-Average rating for Apple Maps: approx. \textbf{6.18}
+Average rating for Apple Maps: approx. \textit{6.18}
 \par
-Average rating for Google Maps: approx. \textbf{7.72}
+Average rating for Google Maps: approx. \textit{7.72}
 
@@
 \section{Heuristic Evaluation}
-All human things are subject to decay. And when fate summons, Monarchs must obey.
+Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
+My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications. I will examine Google Maps first.
 ]
-Hello, here is some text without a meaning.  This text should show what 
-a printed text will look like at this place.
-If you read this text, you will get no information.  Really?  Is there 
-no information?  Is there
+The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, geometric in execution, and with rounded corners.
+Be that as it may, Google Maps completely fails to use any standard iOS icons.
+Though it is not required to use these provided iOS icons, the iOS Human Interface Guidelines state that it is "a good idea to use the built-in icons as much as possible because users already know what they mean".\cite{Apple}
+\includegraphics[width=.45\textwidth]{ios-bar-icons.png}
+\includegraphics[width=.5\textwidth]{google-maps.png}
+
 \end{multicols}
@@
 \bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore,Inc}, 2012.
-\bibitem{HK} Kopka, H., Daly P.W., \emph{A Guide to LaTeX},
-Addison-Wesley, Reading, MA, 1999.
-\bibitem{Pan} Pan, D., ``A Tutorial on MPEG/Audio Compression," \emph{IEEE 
-Multimedia}, Vol.2, pp.60-74, Summer 1998.
+\bibitem{Google} \emph{Material Design}, www.google.com/design/spec/material-design.
+\bibitem{Apple} \emph{iOS Human Interface Guidelines}, https://developer.apple.com/library/ios/documentation/userexperience.
 \end{thebibliography}

### 2014-09-25 usability-study/usability-study.tex
add evaluation format

@@
 Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
-My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications. I will examine Google Maps first.
+My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications.
+\subsection{Google Maps}
 ]
 The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, geometric in execution, and with rounded corners.
-Be that as it may, Google Maps completely fails to use any standard iOS icons.
-Though it is not required to use these provided iOS icons, the iOS Human Interface Guidelines state that it is "a good idea to use the built-in icons as much as possible because users already know what they mean".\cite{Apple}
+Be that as it may, Google Maps completely fails to use any standard iOS icons:
+\par
+\noindent
 \includegraphics[width=.45\textwidth]{ios-bar-icons.png}
+\par
+\noindent
+Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean".\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps.
+There was a higher percent of errors in clicking \underline{wrong icons} in the Google Maps application as there was in the Apple Maps application.
+
 \includegraphics[width=.5\textwidth]{google-maps.png}
@@
 \end{multicols}
+\clearpage
+\begin{multicols}{2}
+[
+Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
+My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications.
+\subsection{Apple Maps}
+]
+\par
+\noindent
+\includegraphics[width=.49\textwidth]{apple-maps.png}
+The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, geometric in execution, and with rounded corners.
+Be that as it may, Google Maps completely fails to use any standard iOS icons:
+\par
+\noindent
+\includegraphics[width=.45\textwidth]{ios-bar-icons.png}
+\par
+\noindent
+Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean".\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps.
+There was a higher percent of errors in clicking \underline{wrong icons} in the Google Maps application as there was in the Apple Maps application.
+
+\end{multicols}
+\clearpage
+\section{Conclusion}
+Usability stuff goes here. blah blah blah
 \clearpage
@@
 \bibitem{Mossberg} Mossberg, Walt., ``The iPhone Takes to the Big Screen," \emph{The Wall Street Journal}, September 2012.
-\bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore,Inc}, 2012.
+\bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore, Inc}, 2012.
 \bibitem{Google} \emph{Material Design}, www.google.com/design/spec/material-design.
-\bibitem{Apple} \emph{iOS Human Interface Guidelines}, https://developer.apple.com/library/ios/documentation/userexperience.
+\bibitem{Apple} \emph{iOS Human Interface Guidelines}, https://developer.apple.com/
 \end{thebibliography}

### 2014-09-25 usability-study/usability-study.tex
add conclusion

@@
 \includegraphics[width=.49\textwidth]{apple-maps.png}
-The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, geometric in execution, and with rounded corners.
-Be that as it may, Google Maps completely fails to use any standard iOS icons:
-\par
-\noindent
-\includegraphics[width=.45\textwidth]{ios-bar-icons.png}
-\par
-\noindent
-Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean".\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps.
-There was a higher percent of errors in clicking \underline{wrong icons} in the Google Maps application as there was in the Apple Maps application.
+One idea that is emphasized in the \textit{iOS Human Interface Guidelines} is to always "defer to content".\cite[Apple] One way in which the guidelines recommend doing this is by taking advantage of the whole screen, which the Apple Maps mobile app seems to be doing.
+The UI helps people understand and interact with the content, but never competes with it.
+Reconsider visual indicators of physicality and realism.
+Bezels, gradients, and drop shadows sometimes lead to heavier UI elements that can overpower or compete with the content.
+Instead, focus on the content and let the UI play a supporting role.
 
@@
 \section{Conclusion}
-Usability stuff goes here. blah blah blah
+The usability metrics of the Google Maps mobile application outranked those of the Apple Maps application in almost every way.
+Though there are some problems with the UI consistency, it does not seem to bother the users (refer back to satisfaction).
+Still, using non-standard icons or processes isn't without drawbacks.
+It's loud.
+It's authorial voice.
+It's greater cognitive load for users.
+Anything inconsistent, first or third party is.
+On an integrated platform like iOS, integrating with the platform offers a lot of benefits, especially when it comes to user experience.
+
+If you use Google apps, either a lot or a little, does the lack of iOS-style buttons make any difference to you?
+
+must recognize that there are limitations to our studies-- didn't test all five metrics; less data for Apple maps; Apple maps is newer; etc.
+Judgment call -- google maps with data that we have now.
 \clearpage

### 2014-09-25 usability-study/usability-study.tex
final touches

@@
 Thus, this study aims to empirically investigate the usability of each application.
-The current consensus within the field of interaction design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics that this particular study focuses on are efficiency, errors, and satisfaction.
+The current consensus within the field of interaction design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
 \section{Contextual Factors}
 Prior to having each test subject perform the tasks, a little contextual data about the users was first collected.
-Users were asked about their previous experience with Apple Maps, their previous experience with Google Maps, and their overall technological proficiency.
+Users were asked about their previous experience with Apple Maps as well as their previous experience with Google Maps.
 \vspace{-.2in}
@@
 The fact that nearly every test subject had significantly more Google Maps experience than Apple Maps experience was a challenge in the conducting of this study.
-Since this study aimed to test the usability metric of efficiency, and not that of learnability, we did not test subjects if they had no prior experience with the application.
+Since this study aimed to test the usability metric of efficiency, and \textit{not} that of learnability, we did not test subjects if they had no prior experience with the application.
 For future studies, however, it might be worthwhile to give inexperienced users a short tutorial so it is possible to test them thereafter.
@@
 \section{The Experiment} \label{sec:Experiment}
-Each person who participated in this study had to carry out three concrete tasks with each of the map applications.
+Each person who participated in this study had to carry out three concrete tasks with the map application[s] with which they were experienced.
 \begin{enumerate}
@@
 \par
-Average time for task \#1 on Apple Maps app: approx. \textit{19 seconds}
+Average time for task \#1 on Google Maps app: $ \approx $ \textit{21.08 seconds}
 \par
-Average time for task \#1 on Google Maps app: approx. \textit{21 seconds}
+Average time for task \#1 on Apple Maps app: $ \approx $ \textit{19.24 seconds}
 \end{center}
@@
 This task's results suggest that users are more efficient using Apple Maps.
-\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it is important to note that there were a couple of anomalies within the set of Google Maps users.}
+\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.}
 \medskip
@@
 \par
-Average time for task \#2 on Apple Maps app: approx. \textit{24 seconds}
+Average time for task \#2 on Google Maps app: $ \approx $ \textit{20.53 seconds}
 \par
-Average time for task \#2 on Google Maps app: approx. \textit{21 seconds}
+Average time for task \#2 on Apple Maps app: $ \approx $ \textit{23.62 seconds}
 \end{center}
@@
 \par
-Average time for task \#3 on Apple Maps app: approx. \textit{36 seconds}
+Average time for task \#3 on Google Maps app: $ \approx $ \textit{15.62 seconds}
 \par
-Average time for task \#3 on Google Maps app: approx. \textit{17 seconds}
+Average time for task \#3 on Apple Maps app: $ \approx $ \textit{35.98 seconds}
 \end{center}
@@
 Task & Application & Specific Error\\\hline
-Task1 & Apple Maps & hit \textit{Share} button; no response from tapping \\
-Task1 & Google Maps & hit \textit{Near You} button to try to drop pin\\
+Task1 & Apple Maps & hit `Share' button to try to map to pin \\
+Task1 & Google Maps & hit `Near You' button to try to drop pin\\
 Task2 & Apple Maps & tried to manually search for ``Nearest Vons"\\
-Task2 & Google Maps & voice command instead of search\\
-Task3 & Apple Maps & hit \textit{Current Location} for directions\\
+Task2 & Google Maps & voice command attempt instead of search\\
+Task3 & Apple Maps & hit `Current Location' for directions\\
 Task3 & Google Maps & found side streets when expecting highway route
@@
 \par
-The \textit{percentage} of subjects with errors is analyzed rather than solely the numbers of errors since more subjects tested the Google Maps app than Apple Maps app.
+The percentage of subjects with errors is analyzed rather than solely the numbers of errors since more subjects tested the Google Maps app than Apple Maps app.
 \begin{table}[ht]
@@
 & \textbf{Task1} & \textbf{Task2} & \textbf{Task3} \\
-Percentage of Apple Users with Errors & 40\% & 50\%  & 20\% \\
-Percentage of Google Users with Errors & 29.4\% & 17.6\%  & 11.7\% \\
+Percentage of Apple Users with Errors & \textit{40\%} & \textit{50\%}  & \textit{20\%} \\
+Percentage of Google Users with Errors & \textit{29.4\%} & \textit{17.6\%}  & \textit{11.7\%} \\
 
@@
 \par
-Average rating for Apple Maps: approx. \textit{6.18}
+Average rating for Apple Maps: $ \approx $ \textit{6.18}
 \par
-Average rating for Google Maps: approx. \textit{7.72}
+Average rating for Google Maps: $ \approx $ \textit{7.72}
 
@@
 ]
-The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, geometric in execution, and with rounded corners.
-Be that as it may, Google Maps completely fails to use any standard iOS icons:
+The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution.
+However, Google Maps completely fails to use any standard iOS icons\cite{Apple}:
 \par
@@
 \noindent
-Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean".\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps.
-There was a higher percent of errors in clicking \underline{wrong icons} in the Google Maps application as there was in the Apple Maps application.
+Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more.
+There was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application than there was in the Apple Maps application.
 
@@
 [
-Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
-My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications.
 \subsection{Apple Maps}
@@
 \noindent
-\includegraphics[width=.49\textwidth]{apple-maps.png}
-One idea that is emphasized in the \textit{iOS Human Interface Guidelines} is to always "defer to content".\cite[Apple] One way in which the guidelines recommend doing this is by taking advantage of the whole screen, which the Apple Maps mobile app seems to be doing.
+\includegraphics[width=.5\textwidth]{apple-maps.png}
+\par
+\noindent
+One idea that is emphasized in the \textit{iOS Human Interface Guidelines} is to always "defer to content".\cite{Apple} One way in which the guidelines recommend doing this is by taking advantage of the whole screen, which the Apple Maps mobile app seems to be doing.
 The UI helps people understand and interact with the content, but never competes with it.

### 2014-09-25 usability-study/usability-study.tex
add graphic for heuristic eval

@@
 \noindent
-One idea that is emphasized in the \textit{iOS Human Interface Guidelines} is to always "defer to content".\cite{Apple} One way in which the guidelines recommend doing this is by taking advantage of the whole screen, which the Apple Maps mobile app seems to be doing.
-The UI helps people understand and interact with the content, but never competes with it.
-Reconsider visual indicators of physicality and realism.
-Bezels, gradients, and drop shadows sometimes lead to heavier UI elements that can overpower or compete with the content.
-Instead, focus on the content and let the UI play a supporting role.
-
+One guideline from iOS HIG is that important content or functionality should be `elevated':\cite{Apple}
+\includegraphics[width=.35\textwidth]{apple-importance.png}
+\par
+\noindent
+This suggests that the icon on the top left-hand corner, which takes you to the directions screen in this case, is the most important.
+I would have to agree that getting directions is very important functionality, so it seems as if the app is complying to the guidelines.
+However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions.
+It is possible that this guideline of important functionality being `elevated' is not intuitive to all users.
 \end{multicols}
+\medskip
+\medskip
+\par
+The lack of standard iOS icons in the Google Maps app may or may not explain some of the efficiency issues as well as some errors that occurred with the tasks on the mobile Google Maps app.
+The guideline of important functionality being towards the upper-left corner of the screen may or may not explain some of the efficiency issues as well as errors that occurred with the tasks on the mobile Apple Maps app.
 \clearpage

### 2014-09-25 usability-study/usability-study.tex
adding files to .gitignore
minor fixes

@@
 Average rating for Google Maps: $ \approx $ \textit{7.72}
-
+\par
+The results suggest that users are more satisfied with Google Maps.
 \begin{multicols}{2}
@@
 Though usability metrics are extremely important in assessing a system, it is also important to heuristically evaluate systems in order to explore \textit{why} they performed the way they do.
-My heuristic evaluations of the mobile map applications will be based on the guidelines documents that correspond to the applications.
-\subsection{Google Maps}
+My heuristic evaluations of the mobile map applications are based on the guidelines documents that correspond to the applications.
+\subsection{Google Maps--- Icons}
 ]
 The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution.
-However, Google Maps completely fails to use any standard iOS icons\cite{Apple}:
-\par
-\noindent
-\includegraphics[width=.45\textwidth]{ios-bar-icons.png}
-\par
-\noindent
+However, Google Maps completely fails to use the standard iOS icons\cite{Apple} (note that the following standard iOS icons are not really in the screen shot of Google Maps on the right) :
+\begin{center}
+\includegraphics[width=.35\textwidth]{ios-bar-icons.png}
+\end{center}
 Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more.
-There was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application than there was in the Apple Maps application.
-
+In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application than there was in the Apple Maps application.
+\begin{center}
 \includegraphics[width=.5\textwidth]{google-maps.png}
-
+\end{center}
 \end{multicols}
+In general, the users of Google Maps were fairly efficient.
+There were, however, a couple of anomalies within the set of users of Google Maps who took significantly longer to complete the first task than the rest of the users.
+These two anomalies both took so long because they initially clicked on the wrong icon.
+In a future study, it would be interesting to see whether or not  the replacement of the Google icons with iOS standard icons would prevent these kinds of anomalies.
 \clearpage
@@
 [
-\subsection{Apple Maps}
+\subsection{Apple Maps--- Placement }
 ]
+\begin{center}
+\includegraphics[width=.51\textwidth]{apple-maps.png}
+\end{center}
 \par
 \noindent
-\includegraphics[width=.5\textwidth]{apple-maps.png}
-\par
-\noindent
-One guideline from iOS HIG is that important content or functionality should be `elevated':\cite{Apple}
+One guideline from the iOS Human Interface Guidelines is that important content or functionality should be `elevated':\cite{Apple}
 \includegraphics[width=.35\textwidth]{apple-importance.png}
@@
 \noindent
-This suggests that the icon on the top left-hand corner, which takes you to the directions screen in this case, is the most important.
+This suggests that the icon on the top left-hand corner ---which happens to be the directions screen in this app--- is the most important.
 I would have to agree that getting directions is very important functionality, so it seems as if the app is complying to the guidelines.
 However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions.
-It is possible that this guideline of important functionality being `elevated' is not intuitive to all users.
+It is possible that this guideline of important functionality being `elevated' may not be intuitive to all users.
 \end{multicols}

### 2014-09-25 usability-study/usability-study.tex
add info on fitts' law

@@
 \noindent
-This suggests that the icon on the top left-hand corner ---which happens to be the directions screen in this app--- is the most important.
-I would have to agree that getting directions is very important functionality, so it seems as if the app is complying to the guidelines.
+This suggests that the icon on the top left-hand corner ---which happens to be the directions screen in the Apple Maps app--- is the most important.
+I would have to agree that getting directions is a very important functionality.
+Thus, it seems as if the app is complying to the guidelines.
 However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions.
@@
 \end{multicols}
-\medskip
-\medskip
 \par
-The lack of standard iOS icons in the Google Maps app may or may not explain some of the efficiency issues as well as some errors that occurred with the tasks on the mobile Google Maps app.
-The guideline of important functionality being towards the upper-left corner of the screen may or may not explain some of the efficiency issues as well as errors that occurred with the tasks on the mobile Apple Maps app.
+The Apple Maps mobile application certainly complies to the guideline of placing important functionality toward the upper-left corner of the screen.
+Although this is a guideline provided by Apple, it is not entirely clear whether or not this is the best placement for important functionality.
+The top corner opposite to the hand the user holds the phone with requires a little extra effort and stretching for many hands.
+This adds an additional variable in an equation known as Fitts' Law\cite{Fitts}, which is indicative of an increase in the movement time.
+The motion between two targets is no longer a fluid, resistance-free motion.
+It is possible that, even though it complies with Apple guidelines, the placement of the directions button caused errors that could be preveted by a different placement.
+In a future study, it would be noteworthy to deduce what placement is optimal for efficiency, error prevention, and satisfaction.
 \clearpage
@@
 \bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore, Inc}, 2012.
+\bibitem{Fitts}Fitts, Paul M. ``The information capacity of the human motor system in controlling the amplitude of movement" \emph{Journal of Experimental Psychology}, June 1954.
 \bibitem{Google} \emph{Material Design}, www.google.com/design/spec/material-design.

### 2014-09-25 usability-study/usability-study.tex
conclusion work

@@
 The usability metrics of the Google Maps mobile application outranked those of the Apple Maps application in almost every way.
-Though there are some problems with the UI consistency, it does not seem to bother the users (refer back to satisfaction).
-Still, using non-standard icons or processes isn't without drawbacks.
+Though there are some inconsistencies with the user interface (UI) of the Google Maps mobile app that are not present in the UI of the Apple Maps mobile app, the users were still more satisfied in general with Google Maps.
+\par
+My heuristic evaluation for Google Maps focuses on the app's decision to not use standard iOS icons.
+One must recognize that using non-standard icons or processes isn't without drawbacks.
 It's loud.
@@
 
-If you use Google apps, either a lot or a little, does the lack of iOS-style buttons make any difference to you?
-
 must recognize that there are limitations to our studies-- didn't test all five metrics; less data for Apple maps; Apple maps is newer; etc.
@@
 %references
-\bibitem{Nielsen} Nielsen, Jakob, \emph{Usability Engineering}, Boston: Academic, 1993.
-\bibitem{Mossberg} Mossberg, Walt., ``The iPhone Takes to the Big Screen," \emph{The Wall Street Journal}, September 2012.
+\bibitem{Apple} \emph{iOS Human Interface Guidelines}, https://developer.apple.com.
 \bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore, Inc}, 2012.
@@
 \bibitem{Google} \emph{Material Design}, www.google.com/design/spec/material-design.
-\bibitem{Apple} \emph{iOS Human Interface Guidelines}, https://developer.apple.com/
+\bibitem{Mossberg} Mossberg, Walt., ``The iPhone Takes to the Big Screen," \emph{The Wall Street Journal}, September 2012.
+\bibitem{Nielsen} Nielsen, Jakob, \emph{Usability Engineering}, Boston: Academic, 1993.
 \end{thebibliography}

### 2014-09-25 usability-study/usability-study.tex
adding conclusion

@@
 My heuristic evaluation for Google Maps focuses on the app's decision to not use standard iOS icons.
-One must recognize that using non-standard icons or processes isn't without drawbacks.
-It's loud.
-It's authorial voice.
-It's greater cognitive load for users.
-Anything inconsistent, first or third party is.
+One must recognize that using non-standard icons or processes is not without drawbacks.
+Anything inconsistent, first or third party, is a greater cognitive load for users.
 On an integrated platform like iOS, integrating with the platform offers a lot of benefits, especially when it comes to user experience.
+\par 
+My heuristic evaluation for Apple Maps focuses on the placement of important functionality.
+My judgment call on which device or system I feel performed best is Google Maps.
 

### 2014-09-25 usability-study/usability-study.tex
formatting tips from class

@@
 \footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.}
-\medskip
-\medskip
+\clearpage
 \par
@@
 \textbf{Second Task: }Find the location of the nearest Vons.
-\begin{figure}[ht]
+\begin{figure}[htbp]
 \begin{center}
-\vspace{-.5in}
+\vspace{-.4in}
 \includegraphics[keepaspectratio, width=.8\textwidth ]{task2.pdf}
@@
 \noindent
-This task's results suggest that users are more efficient using Google Maps. 
 \medskip
 \medskip
+This task's results suggest that users are more efficient using Google Maps. 
 \par
@@
 \textbf{Third Task: }Find the fastest route from the current location to LACMA.
-\begin{figure}[ht]
+\begin{figure}[htbp]
 \begin{center}
@@
 \begin{center}
-\vspace{-.4in}
+\vspace{-.6in}
+
 \par
@@
 This task's results suggest that users are more efficient using Google Maps. 
-
+\clearpage
 \subsection{Errors}
@@
 & \multicolumn{4}{c}{}\\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
-& \textbf{Task1} & \textbf{Task2} & \textbf{Task3} \\
+& \underline{Task1} & \underline{Task2} & \underline{Task3} \\
 Percentage of Apple Users with Errors & \textit{40\%} & \textit{50\%}  & \textit{20\%} \\
@@
 \subsection{Satisfaction}
-\begin{figure}[ht]
+\begin{figure}[htbp]
 \begin{center}
+\vspace{-.4in}
 \includegraphics[keepaspectratio, width=.8\textwidth ]{satisfaction.pdf}
@@
 \end{figure}
+\vspace{-.6in}
 \par
@@
 The results suggest that users are more satisfied with Google Maps.
+\clearpage
 \begin{multicols}{2}

### 2014-09-26 usability-study/usability-study.tex
final touches

@@
 Thus, this study aims to empirically investigate the usability of each application.
-The current consensus within the field of interaction design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
+The current consensus within the field of Interaction Design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
 \section{Contextual Factors}
-Prior to having each test subject perform the tasks, a little contextual data about the users was first collected.
+Prior to having each test subject perform the tasks, a little contextual data about the subject was first collected.
 Users were asked about their previous experience with Apple Maps as well as their previous experience with Google Maps.
@@
 
+\vspace{-.1in}
 \section{Usability Metrics}
+\vspace{-.1in}
 \subsection{Efficiency}
@@
 \noindent
-This task's results suggest that users are more efficient using Apple Maps.
-\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.}
+This task's results suggest that on average, Apple Maps is more efficient for dropping pins and mapping to them.\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.}
 \clearpage
@@
 \noindent
+This task's results suggest that on average, Google Maps is more efficient for finding the desired locations nearest to your current location. 
+\par
 \medskip
-\medskip
-This task's results suggest that users are more efficient using Google Maps. 
-\par
 \noindent
@@
 \noindent
-This task's results suggest that users are more efficient using Google Maps. 
+The third task's results suggest that on average, Google Maps is more efficient for finding the fastest route from your current location to some desired location. 
 \clearpage
@@
 \par
-The only errors collected by this study were incidents wherein the user did something whose result is not what he or she anticipated.
-The table on the following page shows you some examples of errors from this investigation.
+The ``errors" collected by this study were what the field of Interaction Design defines as errors: incidents wherein the user did something whose result is not what he or she anticipated.
+The following table shows you some examples of errors from this investigation.
 
@@
 \par
-The results suggest that Google Maps is more usable with respect to errors.
+The results suggest that errors are less likely with Google Maps .
 
@@
 The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution.
-However, Google Maps completely fails to use the standard iOS icons\cite{Apple} (note that the following standard iOS icons are not really in the screen shot of Google Maps on the right) :
+Thus, Google Maps chooses to not use the standard iOS icons\cite{Apple}:
 \begin{center}
@@
 Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more.
-In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application than there was in the Apple Maps application.
+In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application, which could have possibly been prevented if standard iOS icons were used.
 \begin{center}
@@
 \end{multicols}
-In general, the users of Google Maps were fairly efficient.
-There were, however, a couple of anomalies within the set of users of Google Maps who took significantly longer to complete the first task than the rest of the users.
-These two anomalies both took so long because they initially clicked on the wrong icon.
-In a future study, it would be interesting to see whether or not  the replacement of the Google icons with iOS standard icons would prevent these kinds of anomalies.
+In a future study, it would be interesting to see whether or not  the replacement of the Google icons with iOS standard icons would prevent the Google Maps frequent errors.
 \clearpage
@@
 The motion between two targets is no longer a fluid, resistance-free motion.
-It is possible that, even though it complies with Apple guidelines, the placement of the directions button caused errors that could be preveted by a different placement.
-In a future study, it would be noteworthy to deduce what placement is optimal for efficiency, error prevention, and satisfaction.
+It is possible that, even though it complies with Apple guidelines, the placement of the directions button caused errors that could be prevented by a different placement.
+In a future study, it would be noteworthy to deduce what placement of important functionality is optimal for efficiency, error prevention, and satisfaction.
 \clearpage
@@
 Though there are some inconsistencies with the user interface (UI) of the Google Maps mobile app that are not present in the UI of the Apple Maps mobile app, the users were still more satisfied in general with Google Maps.
-\par
-My heuristic evaluation for Google Maps focuses on the app's decision to not use standard iOS icons.
-One must recognize that using non-standard icons or processes is not without drawbacks.
-Anything inconsistent, first or third party, is a greater cognitive load for users.
-On an integrated platform like iOS, integrating with the platform offers a lot of benefits, especially when it comes to user experience.
-\par 
-My heuristic evaluation for Apple Maps focuses on the placement of important functionality.
-My judgment call on which device or system I feel performed best is Google Maps.
-
-must recognize that there are limitations to our studies-- didn't test all five metrics; less data for Apple maps; Apple maps is newer; etc.
-Judgment call -- google maps with data that we have now.
+Though our data was somewhat limited, the results seem to suggest that the Google Maps mobile app performs better than the Apple Maps mobile app.
 \clearpage
@@
 %references
-\bibitem{Apple} \emph{iOS Human Interface Guidelines}, https://developer.apple.com.
+\bibitem{Apple} ``iOS Human Interface Guidelines- iOS Developer Library", \emph{Apple Inc.} https://developer.apple.com, September 2014.
 \bibitem{ComScore}``Analytics for a Digital World - ComScore, Inc." \emph{ComScore, Inc}, 2012.
 \bibitem{Fitts}Fitts, Paul M. ``The information capacity of the human motor system in controlling the amplitude of movement" \emph{Journal of Experimental Psychology}, June 1954.
-\bibitem{Google} \emph{Material Design}, www.google.com/design/spec/material-design.
+\bibitem{Google} ``Material Design", \emph{Google} www.google.com/design/spec/material-design, September 2014.
 \bibitem{Mossberg} Mossberg, Walt., ``The iPhone Takes to the Big Screen," \emph{The Wall Street Journal}, September 2012.

### 2014-10-12 mental-model-paper/paper2.tex
layout according to acm latex style guide

%
% latex-sample.tex
%
% This LaTeX source file provides a template for a typical research paper.
%

%
% Use the standard article template.
%
\documentclass{article}

% The geometry package allows for easy page formatting.
\usepackage{geometry}
\geometry{letterpaper}

% Load up special logo commands.
\usepackage{doc}

% Package for formatting URLs.
\usepackage{url}

% Packages and definitions for graphics files.
\usepackage{graphicx}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

%
% Set the title, author, and date.
%
\title{Usability of Modern Display Technologies}
\author{Rachel Rivera}
\date{October 16, 2014}

%
% The document proper.
%
\begin{document}

% Add the title section.
\maketitle

% Add an abstract.
\abstract{
This study aims to examine how interaction design concepts specifically map to the usability of modern display technology.
//TODO: write a better abstract
}

% Add various lists on new pages.
\pagebreak
\tableofcontents

% Start the paper on a new page.
\pagebreak

%
% Body text.
%
\section{Introduction}
\label{introduction}

You will almost certainly start with an introductory description of the topic that you investigated in your assignment.  Discuss any goals, motivation, or examples of the subject; the key is to provide the reader with any information that is necessary to understand why your topic was worth investigating.  This descriptive section should also allow the reader to understand the subsequent detail sections on the subject.

\section{Background, Preliminary, and Related Work}

Perhaps the most important functionality to learn for the paper is \LaTeX\ bibliography support.  Citations and references are handled automatically by \LaTeX\ through its companion program, \BibTeX.  All you have to do is provide a bibliography file that provides the reference information and internal keys (very much like variable names) that you use in your document.\footnote{And always remember to run \LaTeX\ \emph{at least twice} after running \BibTeX.}

\BibTeX\ supports virtually all kinds of references, including books \cite{dui,sgg,iokit,palmos}, parts of books \cite{userModeLinux}, articles \cite{nielsen:dui-review,heer-shneiderman,stackableThreads,xpkernel}, and conference proceedings \cite{ux-3d,iring,contextFileSearch,osHaskell,hibernator}, to name a few.  If not already included in your \LaTeX\ distribution, download and install the \texttt{url} package to support formatting of URLs; you can usually mention these in the \emph{note} or \emph{howpublished} fields of your \BibTeX\ file.

Like Section~\ref{introduction}, a background, preliminary, and related work section is also almost certainly needed for your paper.  In this section, describe any history, work, or projects that serve as direct contributors to the subject of your research paper.  Look at other papers in the literature to see how they organized, presented, and discussed prior work.

The Shneiderman/Plaisant text \cite{dui} provide some pointers to seminal or key works; because they made it into the textbook they aren't necessarily ``bleeding edge,'' but they likely provide the foundation for your chosen subject matter.

\section{Main Content Sections}

The outline after the introductory and background, preliminary, and related work sections is more dependent on the specific subject of your research.  Remember to cite references where appropriate, organize the material so that it flows well and is clear to the reader.

\subsection{Multiple Outline Levels}

\LaTeX\ has support for up to three outline levels (\verb!\section!, \verb!\subsection!, and \verb!\subsubsection!).  It also recognizes \verb!\paragraph! and \verb!\subparagraph! directives, though those don't show up in the table of contents.  All of these directives expect a title.

Note also the use of the \verb!\verb! directive for inserting code-like labels or symbols.  It was particularly needed here so that we can include the backslash character in the text.

\subsection{Tables and Figures}

\LaTeX\ has full support for tables and figures.  Table~\ref{table-sample} shows a sample table and Figure~\ref{figure-sample} shows a sample figure.  Note the built-in support for captions and the automated numbering functionality.  Lists of tables and figures can also be automatically generated, as seen at the beginning of this document.

\begin{table}
\centering
\begin{tabular}{|c|c|c|}\hline
Column 1 & Column 2 & Column 3 \\\hline\hline
a & b & c \\
d & e & f \\
g & h & i \\\hline
\end{tabular}

\caption{A sample table}
\label{table-sample}
\end{table}

\begin{figure}
\centering
\includegraphics[width=2in]{space.jpg}

\caption{A sample figure}
\label{figure-sample}
\end{figure}

One very important thing to remember about how \LaTeX\ handles tables and figures by default: you don't have to worry about where they go exactly.  The general rule is that you insert them in the source after your first reference to them, and \LaTeX\ determines their final position.  It also makes decisions on how much page space to devote to them.  This all follows \LaTeX's overall theme of focusing on the content of your paper, and not its format.

Just so you can see a second table, Table~\ref{table-sample2} is provided.

\begin{table}
\centering
\begin{tabular}{|c|c|c|}\hline
Column 1 & Column 2 & Column 3 \\\hline\hline
a & b & c \\
d & e & f \\
g & h & i \\\hline
\end{tabular}

\caption{Another sample table}
\label{table-sample2}
\end{table}

\section{Another Section}

We're adding another section just so you can see how that looks.  Plus there are a few more \LaTeX\ features to illustrate.

\subsection{Bulleted and Numbered Lists}

\LaTeX\ is very good at providing clean lists.  Examples are shown below.

\begin{itemize}
\item Bulleted items come out properly indented and spaced, every time.

\begin{itemize}
\item Sub-bullets are a virtual no-brainer: just nest another \verb!itemize! block.
\item Note how the bullet character automatically changes too.
\end{itemize}

\item Just keep on adding \verb!\item!s\ldots

\item \ldots until you're done.
\end{itemize}

Numbered lists are almost identical, except that you specify \verb!enumerate! instead of \verb!itemize!.  List items are specified in exactly the same way (thus making it easy to change list types).

\begin{enumerate}
\item A list item
\item Another list item
\item A list item with multiple nested lists

\begin{itemize}
\item Nested lists can be of mixed types.
\item That's a lot of power and flexibility for the price of learning a handful of directives.

\begin{enumerate}
\item Like nested bullet lists, nested numbered lists also ``intelligently'' change their numbering schemes.
\item Meanwhile, all \emph{you} have to write is \verb!\item!.  \LaTeX\ does the rest.
\end{enumerate}
\end{itemize}

\item Back to your regularly scheduled list item

\end{enumerate}

\subsection{Subsection with Another Figure}

We may as well include a second figure also, shown in Figure~\ref{figure-sample2}.  The same image file is used, but note how it can be resized.  Again, observe how the positions of the tables and figures do not necessarily match their positions in the source file, reiterating the aforementioned \LaTeX\ functionality for deciding where these items go in the final document.  You provide an approximate location, and \LaTeX\ does the rest.

\begin{figure}
\centering
\includegraphics[width=1in]{space.jpg}

\caption{Another sample figure}
\label{figure-sample2}
\end{figure}

\section{Conclusion}

Wrap up your paper with an ``executive summary'' of the paper itself, reiterating its subject and its major points.  If you want examples, just look at the conclusions from the literature.

% Generate the bibliography.
\bibliography{latex-sample}
\bibliographystyle{unsrt}

\end{document}

### 2014-10-13 mental-model-paper/paper2.tex
finished introduction

@@
 \usepackage{doc}
+\usepackage{cite}
 
@@
 
-You will almost certainly start with an introductory description of the topic that you investigated in your assignment.
-Discuss any goals, motivation, or examples of the subject; the key is to provide the reader with any information that is necessary to understand why your topic was worth investigating.
-This descriptive section should also allow the reader to understand the subsequent detail sections on the subject.
+It is significant to study electronic screens in the field of \textit{human-computer interaction} (HCI) as they are the devices through which information is transferred between the users and the interface.
+Although electronic displays have been around for nearly a century, they are still transforming and evolving in a multitude of ways.
+Screens are diversifying in order to tackle  the limitations that have presented themselves over the years.
+Screen devices have been changing in a variety of ways---including altering in size, shape, flexibility, resolution, and more---in order to increase usability for specific tasks.
 
-\section{Background, Preliminary, and Related Work}
+There exists a plethora of different kinds of screen devices, each designed and built for particular functions.
+For example, \textit{focus plus context screens} (f+c screens), which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents that are too large to fit standard screens.\cite{Baudisch:2002:FPC:506443.506445} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch:2009:BST:1618452.1618505}
+
+ Though each kind of screen has its merits, this study will focus exclusively on touch sensitive screens.
+Many experts within the field of interaction design have discussed how the concepts of the field map specifically to the usability of touch sensitive screens.
+Researchers from Linköping University articulated how ``interaction design on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface".\cite{Albinsson:2003:HPT:642611.642631}
+
+\section{Background/ Prior Work/ Literary Review}
+There is a wealth of previous work pertaining to how interaction design concepts specifically map to the usability of modern display technology.
+
+Some ``Interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface.
+The zero displacement between input and output, control and feedback, hand action and eye gaze, make touch screens very intuitive to use, particularly for novice users.
+Not surprisingly, touch screens have been widely and successfully used in public information kiosks, ticketing machines, bank teller machines, and the like.
+Besides its directness, touch screen interfaces also have additional advantages.
+First, because its control surface is overlaid on the display, no extra input control device or space is necessary for touch screen interaction.
+Second, touch screens are more robust than free moving input devices such as the mouse.
+Being direct between control and display, touch screens also have special limitations.
+First, the user's finger, hand, and arm can obscure part of the screen.
+Second, the human finger as a pointing device has a very low 'resolution'." \cite{Apple} (High Precision Touch Screen Interaction)
+
+Touch screens are ubiquitous on tablets and smartphones, but their flat glass surfaces can hinder close reading and accurate typing.
+People flipping through electronic pages often retain less of what they read than on printed ones, studies suggest.
+Typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention to avoid typos, draining the concentration from the thoughts being expressed.
+Companies are trying to address these problems with new tools adapted from the analog world of three-dimensional typewriter keys, tactile paper pages, and pop quizzes on the blackboard.
+
+Tactus Technology of Fremont, California is developing a keyboard with shape-shifting keys that pop up from the screens surface when needed the reced so that the screen is flat and featureless again.
+Fluid in tiny microchannels raises the keys up and later recedes, making it appear that they keys have melted away.
+The technology will be offered later this year as an accessory to the iPad Mini, co-founder Craig Ciesla said.
+Such tactile features can help build muscle memory and improve accuracy -- skills lost in the rush to touch screes, said Scott MacKenzie, a professor of electrical engineering and computer science at York University in Toronto who specializes in human-computer interaction.
+"As touchscreen phones lack the tactile feel of a physical keyboard, the visual demand on the user is increased." (Eyes-free Text Entry with Error Correction on Touchscreen Mobile Devices) Many people who type on flat glass screen must keep their eyes focused on the surface to hit the correct key, he said.
+"It's not just that visual attention is needed, but a lot of visual attention"
+
+This means less focus on the act of composition.
+Keys that rise and fall don't use up valuable screen real estate when no longer needed, are a good solution.
+Accurate typing isn't the only problem with touch screes.
+Many studies suggest that people's memory and comprehension are often better when they read long passages on paper than on screen, said Mariette DiChristina, editor in chief of Scientific American.
+But electronic textbooks are incorporating ways to compensate for this.
 
@@
 
-% Generate the bibliography.
-\bibliography{latex-sample}
-\bibliographystyle{unsrt}
+\bibliography{mybib}{}
+\bibliographystyle{plain}
+\end{document}
+%%%%%%%%%%%%% end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-10-13 usability-study/usability-study.tex

@@
 \noindent This investigation is motivated by the increasing significance of the usability of systems in today's competitive technology market.
-Usability testing and usability evaluations have become important tools in the assessment of user interfaces.
+Usability testing and evaluation have become important tools in the assessment of user interfaces.
 This paper presents empirical mobile usability studies of two distinct map applications: Google Maps and Apple Maps.
@@
 When the Apple Maps mobile application was first released as part of iOS 6 in September 2012, it received a significant amount of negative feedback.
-Walt Mossberg of the \textit{Wall Street Journal} was one of many who thought that Apple Maps was the biggest drawback of the iPhone 5.\cite{Mossberg} At this time, the Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people.\cite{ComScore} However, Apple Maps grew, matured, and began gaining traction.
+Walt Mossberg of the \textit{Wall Street Journal} was one of many who thought that Apple Maps was the biggest drawback of the iPhone 5 \cite{Mossberg}.
+At this time, the Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people.\cite{ComScore}.
+However, Apple Maps grew, matured, and began gaining traction.
 By September 2013, Apple Maps gained 35 million regular users.
-The number of users of Google Maps dropped to around 58.7 million at this time.\cite{ComScore} Although the Apple Maps mobile app is newer than the Google Maps mobile app,\footnote{The initial release of Google Maps for Mobile was in September of 2008.} the two map applications are comparable with respect to features they provide as well as popularity nowadays.
-\medskip
-\par
+The number of users of Google Maps dropped to around 58.7 million at this time \cite{ComScore}.
+Although the Apple Maps mobile app is newer than the Google Maps mobile app,\footnote{The initial release of Google Maps for Mobile was in September of 2008.} the two map applications are comparable with respect to features they provide as well as popularity nowadays.
+% JD: Excellent contextualization of the study with objective data to back it up.
+
+\medskip % JD: Shouldn't be necessary.  If LaTeX doesn't do it this way, there is likely
+\par     %     a package that changes its behavior, rather than inserting one-offs like this.
 Since both Google Maps apps and Apple Maps apps are widely known and used, it is important that the usability metrics of the two systems are examined.
 Thus, this study aims to empirically investigate the usability of each application.
-The current consensus within the field of Interaction Design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
+The current consensus within the field of Interaction Design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction} \cite{Nielsen}.
+The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
+% JD: I won't change anymore at this point, but citations typically appear *before* the period.
+
 \section{Contextual Factors}
@@
 \end{enumerate}
-In order to keep the study as controlled as possible, all testing was done on iOS devices (no map applications of android devices were tested).
+In order to keep the study as controlled as possible, all testing was done on iOS devices (no map applications of Android devices were tested).
 As test subjects executed the tasks, information was collected pertaining to their efficiency, errors, and satisfaction.
 
+% JD: Yes...shouldn't be necessary to insert this.  LaTeX's intent is to relieve you
+%     of this concern.  If you don't like what you see, leave it be unless it materially
+%     interferes with the communication of your content.
 \vspace{-.1in}
@@
 \noindent
-This task's results suggest that on average, Apple Maps is more efficient for dropping pins and mapping to them.\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.}
+This task's results suggest that on average, Apple Maps is more efficient for dropping pins and mapping to them.\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.} % JD: For more expansive studies, statistical methods help clean this up.
 \clearpage
@@
 \par
-The results suggest that errors are less likely with Google Maps .
+The results suggest that errors are less likely with Google Maps.
 
@@
 ]
-The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution.
-Thus, Google Maps chooses to not use the standard iOS icons\cite{Apple}:
+The icons of the Google Maps mobile application comply with the Google Design Guidelines \cite{Google}: simple, modern, and geometric in execution.
+Thus, Google Maps chooses to not use the standard iOS icons \cite{Apple}:
 \begin{center}
@@
 \end{center}
-Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more.
+Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean \cite{Apple}.'' Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more.
 In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application, which could have possibly been prevented if standard iOS icons were used.
+% JD: Nice, well-connected conclusion.
 \begin{center}
@@
 I would have to agree that getting directions is a very important functionality.
-Thus, it seems as if the app is complying to the guidelines.
+Thus, it seems as if the app is complying with the guidelines.
 However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions.
@@
 The top corner opposite to the hand the user holds the phone with requires a little extra effort and stretching for many hands.
-This adds an additional variable in an equation known as Fitts' Law\cite{Fitts}, which is indicative of an increase in the movement time.
+This adds an additional variable in an equation known as Fitts' Law \cite{Fitts}, which is indicative of an increase in the movement time.
 The motion between two targets is no longer a fluid, resistance-free motion.
@@
 In a future study, it would be noteworthy to deduce what placement of important functionality is optimal for efficiency, error prevention, and satisfaction.
+% JD: Yes, I think this notion of "elevation" needs to be examined alongside the
+%     assorted "navigation models" that iOS describes (hierarchical, tabs, etc.).
+%     There may be some interactions between these two that determine how an element
+%     needs to be placed.
+%
+%     Meanwhile, nice application of Fitts' Law :)
 \clearpage
@@
 %references
+% JD: Not a bad start; next level would be using BibTeX.
 \bibitem{Apple} ``iOS Human Interface Guidelines- iOS Developer Library", \emph{Apple Inc.} https://developer.apple.com, September 2014.

### 2014-10-13 mental-model-paper/paper2.tex
first part of background section complete

@@
 
-It is significant to study electronic screens in the field of \textit{human-computer interaction} (HCI) as they are the devices through which information is transferred between the users and the interface.
-Although electronic displays have been around for nearly a century, they are still transforming and evolving in a multitude of ways.
+The study of electronic screens is significant in the field of human-computer interaction (HCI) as screens are the devices through which information is transferred between the users and the interface.
+Although electronic displays have been around for nearly a century,\footnote{The cathode ray tube (CRT) was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
 Screens are diversifying in order to tackle the limitations that have presented themselves over the years.
-Screen devices have been changing in a variety of ways---including altering in size, shape, flexibility, resolution, and more---in order to increase usability for specific tasks.
 
-There exists a plethora of different kinds of screen devices, each designed and built for particular functions.
-For example, \textit{focus plus context screens} (f+c screens), which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents that are too large to fit standard screens.\cite{Baudisch:2002:FPC:506443.506445} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch:2009:BST:1618452.1618505}
+Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
+For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens.\cite{Baudisch} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch}
 
  Though each kind of screen has its merits, this study will focus exclusively on touch sensitive screens.
-Many experts within the field of interaction design have discussed how the concepts of the field map specifically to the usability of touch sensitive screens.
-Researchers from Linköping University articulated how ``interaction design on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface".\cite{Albinsson:2003:HPT:642611.642631}
+Many experts within the field of interaction design have discussed how the concepts of interaction design map specifically to the usability of touch sensitive screens.
+In a study from Link\"{o}ping University, researchers articulated how ``interaction design on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface."\cite{Albinsson}
 
 \section{Background/ Prior Work/ Literary Review}
-There is a wealth of previous work pertaining to how interaction design concepts specifically map to the usability of modern display technology.
+There exists a fair amount of previous work pertaining to the advantages as well as the limitations of touch sensitive screens.
+One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention to avoid typos.
+A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed.
+\cite{Tinwala:2010:ETE:1868914.1868972} Be that as it may, there exist several text entry methods that attempt to address this problem.
+For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds, and by vibrations.
+\cite{Tinwala:2010:ETE:1868914.1868972} Another proposed solution comes from Tactus Technology, a company in Fremont, California.
+Employees of Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=2.7in]{tactus-keyboard.jpg}
+
+\caption{A Tactus keyboard}
+\label{figure-sample}
+\end{figure}
 
@@
 First, the user's finger, hand, and arm can obscure part of the screen.
-Second, the human finger as a pointing device has a very low 'resolution'." \cite{Apple} (High Precision Touch Screen Interaction)
+Second, the human finger as a pointing device has a very low 'resolution'."
 
@@
 \end{table}
-
-\begin{figure}
-\centering
-\includegraphics[width=2in]{space.jpg}
-
-\caption{A sample figure}
-\label{figure-sample}
-\end{figure}
 
@@
 Wrap up your paper with an ``executive summary'' of the paper itself, reiterating its subject and its major points.  If you want examples, just look at the conclusions from the literature.
+\clearpage
 

### 2014-10-13 mental-model-paper/paper2.tex
add more background and remove placeholder images

@@
 There exists a fair amount of previous work pertaining to the advantages as well as the limitations of touch sensitive screens.
-One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention to avoid typos.
+One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention.
 A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed.
-\cite{Tinwala:2010:ETE:1868914.1868972} Be that as it may, there exist several text entry methods that attempt to address this problem.
-For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds, and by vibrations.
+\cite{Tinwala:2010:ETE:1868914.1868972} There exist several input entry methods, however, that attempt to address this problem.
+For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds.
 \cite{Tinwala:2010:ETE:1868914.1868972} Another proposed solution comes from Tactus Technology, a company in Fremont, California.
-Employees of Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus}
+Employees of Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.
+
 \begin{figure}[ht]
@@
 \includegraphics[width=2.7in]{tactus-keyboard.jpg} 
-
 \caption{A Tactus keyboard}
@@
 \end{figure}
+
+Another limitation of touchscreen devices that is frequently examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated how it is difficult for users to point at targets that are often smaller than the width of their finger.
+A notable approach to tackling this problem was made by Schneiderman, Sears, and colleagues.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
+The user drags the cursor to a desired target and then lifts the finger to select the targeted object.
+Although this technique allows users to perform precise movements with generally less errors, the user's efficiency tends to suffer.
+
+Instead of using a bare finger, in some cases the user may use a stylus (pen) to interact with touch screens.
+A stylus is much ``sharper" pointer than a finger tip, but its resolution may still not be as good as a mouse cursor.
+A study shows taht pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} The drawbacks to the stylus solution, as discussed in a study by Forlines and Balakrishnan, include the occlusion of other elements on the screen.\cite{Forlines}
+
+Zooming is another method that has been investigated in attempts to increase precision of touchscreen interaction.
+It is possible to use zooming to enlarge the information space enough so that the user can comfortably point at the target with a bare finger.
+A common implementation of this zooming technique is called \textit{bounding box zoom}.\cite{Albinsson} With this approach, the user first activates a zooming mode, and then draws a `bounding box' that encloses the area to be zoomed in.
+When the user releases the box, the system zooms and pans to the selected are and the mode switches back to pointing.
+One zooming graphical interface that implements the \textit{bounding box zoom} technique is \textit{Pad++}.\cite{Bederson} Though the \textit{Pad++} helps users effectively touch things on the screen, one cost is that ``designers can no longer rely on the users' familiarity with metaphorical reference and this has learnability consequences." \cite{Bederson}
 
@@
 
-Touch screens are ubiquitous on tablets and smartphones, but their flat glass surfaces can hinder close reading and accurate typing.
-People flipping through electronic pages often retain less of what they read than on printed ones, studies suggest.
-Typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention to avoid typos, draining the concentration from the thoughts being expressed.
-Companies are trying to address these problems with new tools adapted from the analog world of three-dimensional typewriter keys, tactile paper pages, and pop quizzes on the blackboard.
-
-Tactus Technology of Fremont, California is developing a keyboard with shape-shifting keys that pop up from the screens surface when needed the reced so that the screen is flat and featureless again.
-Fluid in tiny microchannels raises the keys up and later recedes, making it appear that they keys have melted away.
-The technology will be offered later this year as an accessory to the iPad Mini, co-founder Craig Ciesla said.
-Such tactile features can help build muscle memory and improve accuracy -- skills lost in the rush to touch screes, said Scott MacKenzie, a professor of electrical engineering and computer science at York University in Toronto who specializes in human-computer interaction.
-"As touchscreen phones lack the tactile feel of a physical keyboard, the visual demand on the user is increased." (Eyes-free Text Entry with Error Correction on Touchscreen Mobile Devices) Many people who type on flat glass screen must keep their eyes focused on the surface to hit the correct key, he said.
-"It's not just that visual attention is needed, but a lot of visual attention"
-
 This means less focus on the act of composition.
 Keys that rise and fall don't use up valuable screen real estate when no longer needed, are a good solution.
-Accurate typing isn't the only problem with touch screes.
+Accurate typing isn't the only problem with touch screens.
 Many studies suggest that people's memory and comprehension are often better when they read long passages on paper than on screen, said Mariette DiChristina, editor in chief of Scientific American.
@@
 All you have to do is provide a bibliography file that provides the reference information and internal keys (very much like variable names) that you use in your document.\footnote{And always remember to run \LaTeX\ \emph{at least twice} after running \BibTeX.}
-
-\BibTeX\ supports virtually all kinds of references, including books \cite{dui,sgg,iokit,palmos}, parts of books \cite{userModeLinux}, articles \cite{nielsen:dui-review,heer-shneiderman,stackableThreads,xpkernel}, and conference proceedings \cite{ux-3d,iring,contextFileSearch,osHaskell,hibernator}, to name a few.
-If not already included in your \LaTeX\ distribution, download and install the \texttt{url} package to support formatting of URLs; you can usually mention these in the \emph{note} or \emph{howpublished} fields of your \BibTeX\ file.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=1in]{space.jpg}
-
-\caption{Another sample figure}
-\label{figure-sample2}
-\end{figure}
-
 \section{Conclusion}

### 2014-10-14 mental-model-paper/paper2.tex
finished literary review section

@@
 \usepackage{cite}
+\usepackage[margin=1cm]{caption}
 
@@
 
-The study of electronic screens is significant in the field of human-computer interaction (HCI) as screens are the devices through which information is transferred between the users and the interface.
+Electronic screens are the devices through which information is transferred between users and the interface.
+In the same way screens are an imperative component of user interaction with a device, the study of screens is an imperative component of the field of interaction design.
+
 Although electronic displays have been around for nearly a century,\footnote{The cathode ray tube (CRT) was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
 Screens are diversifying in order to tackle the limitations that have presented themselves over the years.
-
 Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
@@
  Though each kind of screen has its merits, this study will focus exclusively on touch sensitive screens.
-Many experts within the field of interaction design have discussed how the concepts of interaction design map specifically to the usability of touch sensitive screens.
-In a study from Link\"{o}ping University, researchers articulated how ``interaction design on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface."\cite{Albinsson}
+In a study from Link\"{o}ping University, researchers articulated how ``interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface."\cite{Albinsson} Though it is clear that a connection between the field of interaction design and modern touchscreen devices exists, this study aims to investigate exactly \textit{how} interaction design concepts specifically map to the usability of these screens.
 
 \section{Background/ Prior Work/ Literary Review}
+\subsection{Limitations and Proposed Solutions}
 There exists a fair amount of previous work pertaining to the advantages as well as the limitations of touch sensitive screens.
@@
 \cite{Tinwala:2010:ETE:1868914.1868972} Another proposed solution comes from Tactus Technology, a company in Fremont, California.
-Employees of Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.
+Employees from Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.
 
@@
 \centering
-\includegraphics[width=2.7in]{tactus-keyboard.jpg} 
+\includegraphics[width=2in]{tactus-keyboard.jpg} 
 \caption{A Tactus keyboard}
@@
 
-Another limitation of touchscreen devices that is frequently examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated how it is difficult for users to point at targets that are often smaller than the width of their finger.
+Another limitation of touchscreen devices that is frequently examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
 A notable approach to tackling this problem was made by Schneiderman, Sears, and colleagues.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
 The user drags the cursor to a desired target and then lifts the finger to select the targeted object.
-Although this technique allows users to perform precise movements with generally less errors, the user's efficiency tends to suffer.
+Although this technique generally allowed users to perform precise movements with less errors, the user's efficiency tended to suffer.\cite{Sears}
 
-Instead of using a bare finger, in some cases the user may use a stylus (pen) to interact with touch screens.
-A stylus is much ``sharper" pointer than a finger tip, but its resolution may still not be as good as a mouse cursor.
-A study shows taht pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} The drawbacks to the stylus solution, as discussed in a study by Forlines and Balakrishnan, include the occlusion of other elements on the screen.\cite{Forlines}
+Many studies have purposed using a stylus (pen) to interact with touch screens instead of using a bare finger.
+A stylus has better ``resolution" than a finger tip and recent studies have demonstrated that pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} A study by Forlines and Balakrishnan, which evaluated multiple forms of stylus input, discussed that a major drawback of this approach is the occlusion of other elements on the screen by the pen.\cite{Forlines}
 
-Zooming is another method that has been investigated in attempts to increase precision of touchscreen interaction.
+Zooming is also found in the literature as a proposed solution to increase precision of touchscreen interaction.
 It is possible to use zooming to enlarge the information space enough so that the user can comfortably point at the target with a bare finger.
 A common implementation of this zooming technique is called \textit{bounding box zoom}.\cite{Albinsson} With this approach, the user first activates a zooming mode, and then draws a `bounding box' that encloses the area to be zoomed in.
-When the user releases the box, the system zooms and pans to the selected are and the mode switches back to pointing.
-One zooming graphical interface that implements the \textit{bounding box zoom} technique is \textit{Pad++}.\cite{Bederson} Though the \textit{Pad++} helps users effectively touch things on the screen, one cost is that ``designers can no longer rely on the users' familiarity with metaphorical reference and this has learnability consequences." \cite{Bederson}
+When the user releases the box, the system zooms and pans to the selected area and the mode switches back to pointing.
 
-Some ``Interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface.
-The zero displacement between input and output, control and feedback, hand action and eye gaze, make touch screens very intuitive to use, particularly for novice users.
-Not surprisingly, touch screens have been widely and successfully used in public information kiosks, ticketing machines, bank teller machines, and the like.
-Besides its directness, touch screen interfaces also have additional advantages.
-First, because its control surface is overlaid on the display, no extra input control device or space is necessary for touch screen interaction.
-Second, touch screens are more robust than free moving input devices such as the mouse.
-Being direct between control and display, touch screens also have special limitations.
-First, the user's finger, hand, and arm can obscure part of the screen.
-Second, the human finger as a pointing device has a very low 'resolution'."
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3in]{zoom.png} 
+\caption{An example of \textit{bounding box zoom}. First, the user defines the zoom area by drawing a rectangle (left). Then, the user can the perform direct pointing at a finer scale (right).}
+\label{figure-sample}
+\end{figure}
 
-This means less focus on the act of composition.
-Keys that rise and fall don't use up valuable screen real estate when no longer needed, are a good solution.
-Accurate typing isn't the only problem with touch screens.
-Many studies suggest that people's memory and comprehension are often better when they read long passages on paper than on screen, said Mariette DiChristina, editor in chief of Scientific American.
-But electronic textbooks are incorporating ways to compensate for this.
+An example of a graphical interface that implements the \textit{bounding box zoom} technique is \textit{Pad++}.\cite{Bederson} Though the results of a study by Benjamin Bederson and James Hollan suggested that \textit{Pad++} improved users' overall touchscreen interaction precision, the device's emphasis on expandability and efficiency had certain ``learnability consequences."\cite{Bederson}
 
-Perhaps the most important functionality to learn for the paper is \LaTeX\ bibliography support.
-Citations and references are handled automatically by \LaTeX\ through its companion program, \BibTeX.
-All you have to do is provide a bibliography file that provides the reference information and internal keys (very much like variable names) that you use in your document.\footnote{And always remember to run \LaTeX\ \emph{at least twice} after running \BibTeX.}
+\clearpage
 
-Like Section~\ref{introduction}, a background, preliminary, and related work section is also almost certainly needed for your paper.
-In this section, describe any history, work, or projects that serve as direct contributors to the subject of your research paper.
-Look at other papers in the literature to see how they organized, presented, and discussed prior work.
+\subsection{Advantages}
+Previous work on the subject of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays.
+Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses},  argues that touchscreen devices can be advantageous for multiple reasons: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
+\cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasoning.
+For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} A separate experiment that implemented a touchscreen electronic medical record system supported Schneiderman's assertion that touchscreen display devices are generally more learnable.\cite{Douglas:2011:SUL:2029976.2029990}
 
-The Shneiderman/Plaisant text \cite{dui} provide some pointers to seminal or key works; because they made it into the textbook they aren't necessarily ``bleeding edge,'' but they likely provide the foundation for your chosen subject matter.
+\section{Methods}
 
-\section{Main Content Sections}
+Although there exist several limitations to this study, it seems like the main limitation is the lack of tactile stuff.
+lack of precision can lead to more errors, but the lack of tactile response can make a system completely unusable by someone who has a visual impairment.
+It seems like the main concern is tactile blah.
+this is important because blah.
 
-The outline after the introductory and background, preliminary, and related work sections is more dependent on the specific subject of your research.
-Remember to cite references where appropriate, organize the material so that it flows well and is clear to the reader.
-
-\subsection{Multiple Outline Levels}
+\subsection{What Work is Most Relevant/Important?}
 
@@
 
-\subsection{Tables and Figures}
-
-\LaTeX\ has full support for tables and figures.
-Table~\ref{table-sample} shows a sample table and Figure~\ref{figure-sample} shows a sample figure.
-Note the built-in support for captions and the automated numbering functionality.
-Lists of tables and figures can also be automatically generated, as seen at the beginning of this document.
-
-\begin{table}
-\centering
-\begin{tabular}{|c|c|c|}\hline
-Column 1 & Column 2 & Column 3 \\\hline\hline
-a & b & c \\
-d & e & f \\
-g & h & i \\\hline
-\end{tabular}
-
-\caption{A sample table}
-\label{table-sample}
-\end{table}
-
-One very important thing to remember about how \LaTeX\ handles tables and figures by default: you don't have to worry about where they go exactly.
-The general rule is that you insert them in the source after your first reference to them, and \LaTeX\ determines their final position.
-It also makes decisions on how much page space to devote to them.
-This all follows \LaTeX's overall theme of focusing on the content of your paper, and not its format.
-
-Just so you can see a second table, Table~\ref{table-sample2} is provided.
-
-\begin{table}
-\centering
-\begin{tabular}{|c|c|c|}\hline
-Column 1 & Column 2 & Column 3 \\\hline\hline
-a & b & c \\
-d & e & f \\
-g & h & i \\\hline
-\end{tabular}
-
-\caption{Another sample table}
-\label{table-sample2}
-\end{table}
-
-\section{Another Section}
+\section{Discussion}
 
 We're adding another section just so you can see how that looks.  Plus there are a few more \LaTeX\ features to illustrate.
-
-\subsection{Bulleted and Numbered Lists}
 
@@
 
-\subsection{Subsection with Another Figure}
-
 We may as well include a second figure also, shown in Figure~\ref{figure-sample2}.
@@
 
-\section{Conclusion}
+\section{Conclusions}
 

### 2014-10-14 mental-model-paper/paper2.tex
add first quotes for methods section

@@
 
-Although there exist several limitations to this study, it seems like the main limitation is the lack of tactile stuff.
-lack of precision can lead to more errors, but the lack of tactile response can make a system completely unusable by someone who has a visual impairment.
-It seems like the main concern is tactile blah.
-this is important because blah.
+Though literature on the subject of touchscreen display devices have presented many limitations, the most relevant and important seems to be the lack of tactile features and the requirement for heightened visual attention.
+For people who are blind, accessing touch screen interfaces is nearly impossible.
+Some accessibility features are beginning to be developed though
+
+The principle of feedback
+feedback ---sending back to the user information about what action has actually been done, what result has been accomplished--- is a well-known concept in the science of control and information theory\cite{Norman02} "In the good old days of the telephone", "telephones were designed with much more care and concern for the user.
+Designers at the Bell Telephone Laboratories worried a lot about feedback.
+The push buttons were designed to give an appropriate feel---tactile feedback.
+When a button was pushed, a tone was fed back into the earpiece so the user could tell that the button had been properly pushed.
+When the phone call was being connected, clicks, tones, and other noises gave the user feedback about the progress of the call." ``All this changed." ``Why are modern telephone systems so difficult to use?
+Basically, the problem is that the systems have more features and less feedback"
 

### 2014-10-14 mental-model-paper/paper2.tex
restructure prior work

@@
 \section{Background/ Prior Work/ Literary Review}
-\subsection{Limitations and Proposed Solutions}
-There exists a fair amount of previous work pertaining to the advantages as well as the limitations of touch sensitive screens.
+\subsection{Limitations/Proposed Solutions}
+There exists a fair amount of previous work pertaining to the the limitations of touch sensitive screens.
 One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention.
 A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed.
-\cite{Tinwala:2010:ETE:1868914.1868972} There exist several input entry methods, however, that attempt to address this problem.
-For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds.
-\cite{Tinwala:2010:ETE:1868914.1868972} Another proposed solution comes from Tactus Technology, a company in Fremont, California.
-Employees from Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=2in]{tactus-keyboard.jpg} 
-\caption{A Tactus keyboard}
-\label{figure-sample}
-\end{figure}
+\cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices extremely difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI).
+A study from Virginia Polytechnic Institute and State University examined how, even with the use of \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}} functionality, the user can only ``develop a spatial mental model for the interface or the screen through dead reckoning" as there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
 
 Another limitation of touchscreen devices that is frequently examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
-A notable approach to tackling this problem was made by Schneiderman, Sears, and colleagues.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
+A notable technique created specifically to address this problem was made by Schneiderman, Sears, and colleagues.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
 The user drags the cursor to a desired target and then lifts the finger to select the targeted object.
@@
 Many studies have purposed using a stylus (pen) to interact with touch screens instead of using a bare finger.
-A stylus has better ``resolution" than a finger tip and recent studies have demonstrated that pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} A study by Forlines and Balakrishnan, which evaluated multiple forms of stylus input, discussed that a major drawback of this approach is the occlusion of other elements on the screen by the pen.\cite{Forlines}
+A stylus has better ``resolution" than a finger tip and recent studies have demonstrated that pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} A study by Forlines and Balakrishnan, which evaluated multiple forms of stylus input, articulated that one drawback of this approach is the occlusion of other elements on the screen by the pen.\cite{Forlines}
 
 Zooming is also found in the literature as a proposed solution to increase precision of touchscreen interaction.
-It is possible to use zooming to enlarge the information space enough so that the user can comfortably point at the target with a bare finger.
-A common implementation of this zooming technique is called \textit{bounding box zoom}.\cite{Albinsson} With this approach, the user first activates a zooming mode, and then draws a `bounding box' that encloses the area to be zoomed in.
-When the user releases the box, the system zooms and pans to the selected area and the mode switches back to pointing.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=3in]{zoom.png} 
-\caption{An example of \textit{bounding box zoom}. First, the user defines the zoom area by drawing a rectangle (left). Then, the user can the perform direct pointing at a finer scale (right).}
-\label{figure-sample}
-\end{figure}
-
-An example of a graphical interface that implements the \textit{bounding box zoom} technique is \textit{Pad++}.\cite{Bederson} Though the results of a study by Benjamin Bederson and James Hollan suggested that \textit{Pad++} improved users' overall touchscreen interaction precision, the device's emphasis on expandability and efficiency had certain ``learnability consequences."\cite{Bederson}
-
-\clearpage
+It is possible to use zooming to enlarge the information space of the touchscreen enough so that the user can comfortably point at the target with a bare finger.
+However, a study by P\"{a}r-Anders Albinsson and Shumin Zhai asserted that zooming has a fundamental drawback since ``when zoomed to a sub area of interest, one loses the contextual global view that can be important for the user's task."\cite{Albinsson}
 
@@
 \section{Methods}
+There exist several input entry methods, however, that attempt to address this problem.
+For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds.
+\cite{Tinwala:2010:ETE:1868914.1868972}
+Another proposed solution comes from Tactus Technology, a company in Fremont, California.
+Employees from Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3in]{tactus-keyboard.jpg} 
+\caption{A Tactus keyboard}
+\label{figure-sample}
+\end{figure}
 Though literature on the subject of touchscreen display devices have presented many limitations, the most relevant and important seems to be the lack of tactile features and the requirement for heightened visual attention.
 For people who are blind, accessing touch screen interfaces is nearly impossible.
-Some accessibility features are beginning to be developed though
+Some accessibility features are beginning to be developed though.
+
+``Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility have remained active throughout this time.\cite{Buxton:1986:HID:22339.22386} Touch 
+screen accessibility research has considered various device 
+form factors. These techniques relied upon hardware modifications, 
+such as augmenting a touch screen with physical buttons or 
+placing a physical overlay atop the screen. Such approaches 
+may be expensive to install, may limit the flexibility of the 
+underlying software (by imposing physical structures), and 
+may interfere with use by sighted people; unsurprisingly, 
+such techniques have not been widely adopted. 
+More recent efforts have focused on using gestures to 
+provide blind users with access to touch screens without 
+modifying the underlying hardware. For example, Slide 
+Rule\cite{Kane:2008:SRM:1414471.1414487} used multi-touch gestures to allow blind users to 
+browse and explore content on a touch screen-based smartphone.  Researchers have also explored techniques to 
+address specific aspects of blind touch screen interaction, 
+such as text entry and gesture selection\cite{conf/chi/KaneWL11}. 
+However, these techniques have often focused on small, 
+mobile phone-sized screens. Furthermore, many of these 
+techniques change the fundamental layout of the screen to 
+improve accessibility. Thus, our current work addresses 
+interaction with larger touch screens, and explores methods 
+to preserve users’ understanding of the screen layout.
 
@@
 
-We're adding another section just so you can see how that looks.  Plus there are a few more \LaTeX\ features to illustrate.
-
+Though there is awareness, nothing has been widely accepted yet. More research should be devoted to this problem.
 \LaTeX\ is very good at providing clean lists.  Examples are shown below.

### 2014-10-15 mental-model-paper/paper2.tex
tweaking background section

@@
 Electronic screens are the devices through which information is transferred between users and the interface.
-In the same way screens are an imperative component of user interaction with a device, the study of screens is an imperative component of the field of interaction design.
-
-Although electronic displays have been around for nearly a century,\footnote{The cathode ray tube (CRT) was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
-Screens are diversifying in order to tackle the limitations that have presented themselves over the years.
-Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
+Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
+Screens are diversifying in order to tackle the limitations that have presented themselves over the years.\cite{Eisenberg} Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
 For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens.\cite{Baudisch} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch}
@@
  Though each kind of screen has its merits, this study will focus exclusively on touch sensitive screens.
-In a study from Link\"{o}ping University, researchers articulated how ``interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface."\cite{Albinsson} Though it is clear that a connection between the field of interaction design and modern touchscreen devices exists, this study aims to investigate exactly \textit{how} interaction design concepts specifically map to the usability of these screens.
+In a study from Link\"{o}ping University, researchers articulated how ``interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface."\cite{Albinsson} Though it is clear that a connection between modern touchscreen devices and the field of interaction design exists, this study aims to investigate exactly \textit{how} interaction design concepts specifically map to the usability of touch screens.
 
@@
 There exists a fair amount of previous work pertaining to the the limitations of touch sensitive screens.
-One limitation that is often discussed in the literature is how typing on flat surfaces with no physical keys to guide the fingers requires heightened visual attention.
+One limitation that is discussed in the literature is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention.
 A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed.
 \cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices extremely difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI).
-A study from Virginia Polytechnic Institute and State University examined how, even with the use of \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}} functionality, the user can only ``develop a spatial mental model for the interface or the screen through dead reckoning" as there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
+A study from Virginia Polytechnic Institute and State University examined how, even with the use of \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}} functionality, touch screens do not provide sufficient support for IBSVI.
+These users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning" as there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
 
-Another limitation of touchscreen devices that is frequently examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
-A notable technique created specifically to address this problem was made by Schneiderman, Sears, and colleagues.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
+One purposed solution to this limitation is a physical overlay, developed by researchers specifically to help IBSVI engage their spatial cognition, perception, and sensing resources while interacting with touch screens.
+\cite{El-Glaly:2013:TTF:2460625.2460665} The tactile overlay covers the touch device screen.
+An IBSVI can use the tactile patterns of the overlay to locate the text, icons, or buttons in the touch device interface.
+A study examining the efficacy of this physical overlay suggested IBSVI who used it were typically more efficient than those using voice-over technology or `trail-and-error' exploration.
+\cite{El-Glaly:2013:TTF:2460625.2460665}
+
+Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
+A notable technique was created by Schneiderman, Sears, and colleagues to specifically address this problem.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
 The user drags the cursor to a desired target and then lifts the finger to select the targeted object.
-Although this technique generally allowed users to perform precise movements with less errors, the user's efficiency tended to suffer.\cite{Sears}
+Although this technique generally allowed a user to perform precise movements with less errors, the user's efficiency tended to suffer.\cite{Sears}
 
@@
 Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses},  argues that touchscreen devices can be advantageous for multiple reasons: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
-\cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasoning.
-For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} A separate experiment that implemented a touchscreen electronic medical record system supported Schneiderman's assertion that touchscreen display devices are generally more learnable.\cite{Douglas:2011:SUL:2029976.2029990}
+\cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasons.
+For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Schneiderman's assertion that touchscreen display devices are generally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}
 

### 2014-10-16 mental-model-paper/paper2.tex
more background work/methods work

@@
 \subsection{Limitations/Proposed Solutions}
-There exists a fair amount of previous work pertaining to the the limitations of touch sensitive screens.
+There exists a fair amount of previous work pertaining to the usability of touch sensitive screens.
+The literature often articulates limitations, as well as proposed solutions, with respect to usability.
 One limitation that is discussed in the literature is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention.
 A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed.
-\cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices extremely difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI).
-A study from Virginia Polytechnic Institute and State University examined how, even with the use of \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}} functionality, touch screens do not provide sufficient support for IBSVI.
+\cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI).
+A study from Virginia Polytechnic Institute and State University examined how, even with the use of features such as  \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}}, touch screens still do not provide sufficient support for IBSVI.
 These users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning" as there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
@@
 \section{Methods}
+The prior work in the field demonstrates how there is a variety of different advantages as well as limitations that can be examined when evaluating the usability of touchscreen devices.
+This study will focus exclusively on the limitation of the minimal tactile feedback provided by touchscreen devices, and what this means with respect to usability.
+
 There exist several input entry methods, however, that attempt to address this problem.

### 2014-10-21 mental-model-paper/paper2.tex
work on methods

@@
 
-Electronic screens are the devices through which information is transferred between users and the interface.
+Electronic screens are the devices through which information is transferred between the user and the interface.
 Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
@@
 There exists a fair amount of previous work pertaining to the usability of touch sensitive screens.
-The literature often articulates limitations, as well as proposed solutions, with respect to usability.
+The literature often evaluates usability by examining the limitations of these screen devices.
 One limitation that is discussed in the literature is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention.
@@
 A study from Virginia Polytechnic Institute and State University examined how, even with the use of features such as  \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}}, touch screens still do not provide sufficient support for IBSVI.
-These users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning" as there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
+These users were only able to ``develop a spatial mental model for the interface or the screen through dead reckoning" since there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
 
@@
 An IBSVI can use the tactile patterns of the overlay to locate the text, icons, or buttons in the touch device interface.
-A study examining the efficacy of this physical overlay suggested IBSVI who used it were typically more efficient than those using voice-over technology or `trail-and-error' exploration.
+A study examining the efficacy of this physical overlay illustrated that IBSVI who used it were generally more efficient than those using voice-over technology or `trail-and-error' exploration.
 \cite{El-Glaly:2013:TTF:2460625.2460665}
@@
 Previous work on the subject of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays.
-Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses},  argues that touchscreen devices can be advantageous for multiple reasons: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
+Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, provides a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
 \cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasons.
@@
 \section{Methods}
-The prior work in the field demonstrates how there is a variety of different advantages as well as limitations that can be examined when evaluating the usability of touchscreen devices.
-This study will focus exclusively on the limitation of the minimal tactile feedback provided by touchscreen devices, and what this means with respect to usability.
+The prior work in the field demonstrates that there is a variety of different advantages as well as limitations that can be examined when evaluating the usability of touchscreen devices.
+Since this paper is too short to discuss each advantage and limitation, the paper will focus exclusively on the limitation of the minimal tactile feedback provided by touchscreen devices.
+
+Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility have remained active throughout this time.
+\cite{Buxton:1986:HID:22339.22386} Previous research has explored techniques that make touch screens more accessible by modifying the hardware.
+This paper mentioned one such technique when explaining physical overlays.\cite{El-Glaly:2013:TTF:2460625.2460665} However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people." \cite{Kane:2011:AOI:2047196.2047232}  The study asserts that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
+
+\clearpage
+More recent efforts have focused on using gestures to provide IBSVI with access to touch screens without modifying the underlying hardware.
+For example, Slide Rule \cite{Kane:2008:SRM:1414471.1414487} used multi-touch gestures to allow blind users to browse and explore content on a touch screen-based smartphone.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=4.5in]{slide-rule.jpg} 
+\caption{An example of some multi-touch gestures recognized by Slide Rule.(1) A one-finger scan to browse lists; (2) A two-finger tap to select items; (3) A flick gesture used to flip between pages; (4) An L-select gesture to browse hierarchy of menu options.
+\cite{Kane:2008:SRM:1414471.1414487} }
+\label{figure-sample}
+\end{figure}
+
+  Researchers have also explored techniques to 
+address specific aspects of blind touch screen interaction, 
+such as text entry and gesture selection\cite{conf/chi/KaneWL11}. 
+However, these techniques have often focused on small, 
+mobile phone-sized screens. Furthermore, many of these 
+techniques change the fundamental layout of the screen to 
+improve accessibility. Thus, our current work addresses 
+interaction with larger touch screens, and explores methods 
+to preserve users’ understanding of the screen layout.
 
@@
 
-``Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility have remained active throughout this time.\cite{Buxton:1986:HID:22339.22386} Touch 
-screen accessibility research has considered various device 
-form factors. These techniques relied upon hardware modifications, 
-such as augmenting a touch screen with physical buttons or 
-placing a physical overlay atop the screen. Such approaches 
-may be expensive to install, may limit the flexibility of the 
-underlying software (by imposing physical structures), and 
-may interfere with use by sighted people; unsurprisingly, 
-such techniques have not been widely adopted. 
-More recent efforts have focused on using gestures to 
-provide blind users with access to touch screens without 
-modifying the underlying hardware. For example, Slide 
-Rule\cite{Kane:2008:SRM:1414471.1414487} used multi-touch gestures to allow blind users to 
-browse and explore content on a touch screen-based smartphone.  Researchers have also explored techniques to 
-address specific aspects of blind touch screen interaction, 
-such as text entry and gesture selection\cite{conf/chi/KaneWL11}. 
-However, these techniques have often focused on small, 
-mobile phone-sized screens. Furthermore, many of these 
-techniques change the fundamental layout of the screen to 
-improve accessibility. Thus, our current work addresses 
-interaction with larger touch screens, and explores methods 
-to preserve users’ understanding of the screen layout.
-
+\section{Discussion}
 The principle of feedback
@@
 Basically, the problem is that the systems have more features and less feedback"
-
-\subsection{What Work is Most Relevant/Important?}
-
-\LaTeX\ has support for up to three outline levels (\verb!\section!, \verb!\subsection!, and \verb!\subsubsection!).
-It also recognizes \verb!\paragraph!
-and \verb!\subparagraph!
-directives, though those don't show up in the table of contents.
-All of these directives expect a title.
-
-Note also the use of the \verb!\verb! directive for inserting code-like labels or symbols.  It was particularly needed here so that we can include the backslash character in the text.
-
-\section{Discussion}
-
-Though there is awareness, nothing has been widely accepted yet. More research should be devoted to this problem.
-\LaTeX\ is very good at providing clean lists.  Examples are shown below.
-
-\begin{itemize}
-\item Bulleted items come out properly indented and spaced, every time.
-
-\begin{itemize}
-\item Sub-bullets are a virtual no-brainer: just nest another \verb!itemize! block.
-\item Note how the bullet character automatically changes too.
-\end{itemize}
-
-\item Just keep on adding \verb!\item!s\ldots
-
-\item \ldots until you're done.
-\end{itemize}
-
-Numbered lists are almost identical, except that you specify \verb!enumerate! instead of \verb!itemize!.  List items are specified in exactly the same way (thus making it easy to change list types).
-
-\begin{enumerate}
-\item A list item
-\item Another list item
-\item A list item with multiple nested lists
-
-\begin{itemize}
-\item Nested lists can be of mixed types.
-\item That's a lot of power and flexibility for the price of learning a handful of directives.
-
-\begin{enumerate}
-\item Like nested bullet lists, nested numbered lists also ``intelligently'' change their numbering schemes.
-\item Meanwhile, all \emph{you} have to write is \verb!\item!.  \LaTeX\ does the rest.
-\end{enumerate}
-\end{itemize}
-
-\item Back to your regularly scheduled list item
-
-\end{enumerate}
-
-We may as well include a second figure also, shown in Figure~\ref{figure-sample2}.
-The same image file is used, but note how it can be resized.
-Again, observe how the positions of the tables and figures do not necessarily match their positions in the source file, reiterating the aforementioned \LaTeX\ functionality for deciding where these items go in the final document.
-You provide an approximate location, and \LaTeX\ does the rest.
 

### 2014-10-21 mental-model-paper/paper2.tex
general structure complete

@@
 
-Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility have remained active throughout this time.
-\cite{Buxton:1986:HID:22339.22386} Previous research has explored techniques that make touch screens more accessible by modifying the hardware.
+\subsection{Prevailing Set of Views}
+Most research pertaining to this limitation focuses on how the lack of tactile feedback affects the usability of the device for Individuals with Blindness or Severe Visual Impairment.
+Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility for IBSVI have remained active throughout this time.
+\cite{Buxton:1986:HID:22339.22386} The prevailing set of views on the subject recognize that the touch screens are not effective for the IBSVI, and propose solutions specifically to address needs of the IBSVI.
+
+Previous research has explored techniques that make touch screens more accessible by modifying the hardware.
 This paper mentioned one such technique when explaining physical overlays.\cite{El-Glaly:2013:TTF:2460625.2460665} However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people." \cite{Kane:2011:AOI:2047196.2047232}  The study asserts that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
 
-\clearpage
 More recent efforts have focused on using gestures to provide IBSVI with access to touch screens without modifying the underlying hardware.
@@
 
+There exist several input entry methods, however, that attempt to address this problem.
+For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds.
+\cite{Tinwala:2010:ETE:1868914.1868972}
+
   Researchers have also explored techniques to 
@@
 
-There exist several input entry methods, however, that attempt to address this problem.
-For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds.
-\cite{Tinwala:2010:ETE:1868914.1868972}
+\clearpage
+\subsection{Emerging technology}
+There is technology emerging that is not necessarily aimed to specifically help the IBSVI.
+
 Another proposed solution comes from Tactus Technology, a company in Fremont, California.
@@
 \section{Discussion}
+Research discusses usability for IBSVI but does has not delved into technology that can be effective for the sighted and IBSVI alike.
 The principle of feedback

### 2014-10-21 mental-model-paper/paper2.tex
very rough draft minus conclusion

@@
-%
-% latex-sample.tex
-%
-% This LaTeX source file provides a template for a typical research paper.
-%
-
-%
-% Use the standard article template.
-%
-\documentclass{article}
+\documentclass[11pt]{article}
 
@@
 \author{Rachel Rivera}
-\date{October 16, 2014}
+\date{October 21, 2014}
 
@@
 
-Electronic screens are the devices through which information is transferred between the user and the interface.
+Electronic screens are devices through which information is transferred between the user and the interface.
 Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3in]{TactileGrid.jpg} 
+\caption{Tactile grid}
+\label{figure-sample}
+\end{figure}
+
 Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
@@
 \subsection{Advantages}
-Previous work on the subject of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays.
-Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, provides a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
+Previous work on the usability of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays.
+Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, suggests a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
 \cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasons.
-For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Schneiderman's assertion that touchscreen display devices are generally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}
+For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Schneiderman's assertion that touchscreen display devices are genertally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}
+
+Studies suggest that there also exist usability advantages to using the graphically rendered, non-tactile buttons on touch screens, which are often called \textit{soft buttons}.
+A study by Seungyon Lee and Shumin Zhai articulates that some advantages to soft buttons compared to their hard counterparts include their ability to appear or disappear according to the interaction context, their ability to change size depending on screen-space availability, and the ease with which they can be stylized or updated.\cite{Lee:2009:PTS:1518701.1518750} The results from the experiments presented in Lee and Zhai's study suggest that for tasks such as  dialing phone numbers and using calculator functions soft buttons ``can offer a level of performance that is similar, and in some cases even superior, to the level of performance offered by typical hard buttons embedded in today's handheld devices." \cite{Lee:2009:PTS:1518701.1518750}
 
 \section{Methods}
-The prior work in the field demonstrates that there is a variety of different advantages as well as limitations that can be examined when evaluating the usability of touchscreen devices.
-Since this paper is too short to discuss each advantage and limitation, the paper will focus exclusively on the limitation of the minimal tactile feedback provided by touchscreen devices.
+The prior work in the field proves that there is a variety of different advantages, limitations, and proposed solutions to those limitations that can be examined when evaluating the usability of touchscreen devices.
+Since this paper is too limited in both scope and depth to discuss each advantage and limitation, the paper will focus exclusively on the limitation of the minimal tactile feedback provided by touchscreen devices.
 
 \subsection{Prevailing Set of Views}
-Most research pertaining to this limitation focuses on how the lack of tactile feedback affects the usability of the device for Individuals with Blindness or Severe Visual Impairment.
+Most research pertaining to this limitation focuses on how the lack of tactile feedback affects the usability of devices for Individuals with Blindness or Severe Visual Impairment---as previously touched on in this study.
+The prevailing set of views is that the lack of accessibility for IBSVI is the main problem with touch screen devices.
 Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility for IBSVI have remained active throughout this time.
-\cite{Buxton:1986:HID:22339.22386} The prevailing set of views on the subject recognize that the touch screens are not effective for the IBSVI, and propose solutions specifically to address needs of the IBSVI.
+\cite{Buxton:1986:HID:22339.22386} Thus, a significant portion of the proposed solutions for this limitation are developed specifically to address the needs of IBSVI.
 
-Previous research has explored techniques that make touch screens more accessible by modifying the hardware.
+Previous research has explored techniques that make touch screens more accessible to IBSVI by modifying the hardware.
 This paper mentioned one such technique when explaining physical overlays.\cite{El-Glaly:2013:TTF:2460625.2460665} However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people." \cite{Kane:2011:AOI:2047196.2047232}  The study asserts that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
 
-More recent efforts have focused on using gestures to provide IBSVI with access to touch screens without modifying the underlying hardware.
+Researchers have also explored techniques using access overlays, which are implemented as semi-transparent windows that reside above a standard application.
+When activated, an access overlay gathers information about the location and content of all targets on the screen, and provides access to these targets through a combination of speech and audio feedback, alternative gesture input, and additional user interface controls.
+As access overlays are entirely software-based, they do not require alterations to the underlying touch screen hardware.
+Results from a recent study suggested that access overlays enabled users with visual impairments to locate on-screen targets faster than traditional techniques, improved their spatial understanding, and were preferred.\cite{Kane:2011:AOI:2047196.2047232} However,  access overlays have also not been widely adopted.
+
+Recent efforts have focused on using gestures as another way to provide IBSVI with access to touch screens.
 For example, Slide Rule \cite{Kane:2008:SRM:1414471.1414487} used multi-touch gestures to allow blind users to browse and explore content on a touch screen-based smartphone.
@@
 
-There exist several input entry methods, however, that attempt to address this problem.
-For instance, Tinwala and MacKenzie present a system that includes auditory and tactile feedback to guide eyes-free entry using speech and non-speech sounds.
-\cite{Tinwala:2010:ETE:1868914.1868972}
+Slide Rule's design is ``based on interviews with blind mobile device users and on user-centered design with blind people." \cite{Kane:2008:SRM:1414471.1414487} The results from a study pertaining to the usability of Slide Rule showed that users were able to complete tasks more quickly with Slide Rule than with a button-based mobile screen reader.
+However, they did make more errors.
+\cite{Kane:2008:SRM:1414471.1414487} Similar interfaces now appear in consumer devices such as Apple's iPhone\cite{AppleMultiTouch} and Android-based smartphones.\cite{AndroidMultiTouch}
 
-  Researchers have also explored techniques to 
-address specific aspects of blind touch screen interaction, 
-such as text entry and gesture selection\cite{conf/chi/KaneWL11}. 
-However, these techniques have often focused on small, 
-mobile phone-sized screens. Furthermore, many of these 
-techniques change the fundamental layout of the screen to 
-improve accessibility. Thus, our current work addresses 
-interaction with larger touch screens, and explores methods 
-to preserve users’ understanding of the screen layout.
+Researchers have also explored techniques to address specific aspects of blind touch screen interaction such as text entry\cite{Tinwala:2010:ETE:1868914.1868972, Bonner:2010:NNA:2166616.2166649} and gesture selection.\cite{Kane:2011:UGB:1978942.1979001}
 
-\clearpage
-\subsection{Emerging technology}
-There is technology emerging that is not necessarily aimed to specifically help the IBSVI.
+The significance of all of these proposed solutions is that they were each developed by researchers specifically to address the needs of Individuals with Blindness or Severe Visual Impairment.
+Many studies, including each study mentioned in this section, only tested solutions improved usability for IBSVI.
+The extent to which these solutions helped sighted users was not examined.
 
-Another proposed solution comes from Tactus Technology, a company in Fremont, California.
+\subsection{Relevant Emerging Technology}
+There exists some relevant emerging technology that makes touch screen devices more tactile.
+Though these emerging devices may not have been the direct result of research pertaining to usability, they do provide some noteworthy possible solutions to the usability limitation.
+
+Tactus Technology, a company in Fremont, California.
 Employees from Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.

### 2014-10-22 mental-model-paper/paper2.tex
almost final finish

@@
 
-% The geometry package allows for easy page formatting.
 \usepackage{geometry}
@@
 
-% Load up special logo commands.
 \usepackage{doc}
@@
 
-% Package for formatting URLs.
 \usepackage{url}
 
-% Packages and definitions for graphics files.
 \usepackage{graphicx}
@@
 
-%
-% Set the title, author, and date.
-%
 \title{Usability of Modern Display Technologies}
@@
 
-%
-% The document proper.
-%
 \begin{document}
 
-% Add the title section.
 \maketitle
 
-% Add an abstract.
 \abstract{
 This study aims to examine how interaction design concepts specifically map to the usability of modern display technology.
-//TODO: write a better abstract
+This paper discusses the limitation of the lack of tactile feedback for touchscreen devices as well as how this limitation affects usability.
+This paper examines some current proposed solutions for making touch surfaces truly tangible.
+This limitation and each proposed solution has a variety of implications with respect to learnability, efficiency, satisfaction and errors.
+
 }
 
-% Add various lists on new pages.
 \pagebreak
@@
 
-% Start the paper on a new page.
 \pagebreak
 
-%
-% Body text.
-%
 \section{Introduction}
@@
 Electronic screens are devices through which information is transferred between the user and the interface.
-Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still transforming and evolving in a multitude of ways.
-Screens are diversifying in order to tackle the limitations that have presented themselves over the years.\cite{Eisenberg} Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
+Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still diversifying in a multitude of ways to tackle the limitations that have presented themselves over the years.\cite{Eisenberg} Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
 For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens.\cite{Baudisch} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch}
@@
  Though each kind of screen has its merits, this study will focus exclusively on touch sensitive screens.
-In a study from Link\"{o}ping University, researchers articulated how ``interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface."\cite{Albinsson} Though it is clear that a connection between modern touchscreen devices and the field of interaction design exists, this study aims to investigate exactly \textit{how} interaction design concepts specifically map to the usability of touch screens.
+In a study from Link\''{o}ping University, researchers articulated how ``interaction on touch sensitive screens is literally the most `direct' form of HCI, where information display and control are but one surface.''\cite{Albinsson} Though it is clear that a connection between modern touchscreen devices and the field of interaction design exists, this study aims to investigate exactly \textit{how} interaction design concepts specifically map to the usability of touch screens.
 
 \section{Background/ Prior Work/ Literary Review}
+\label{background}
 \subsection{Limitations/Proposed Solutions}
 There exists a fair amount of previous work pertaining to the usability of touch sensitive screens.
-The literature often evaluates usability by examining the limitations of these screen devices.
-One limitation that is discussed in the literature is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention.
+The literature often evaluates usability by examining the limitations of these screen devices, as well as the proposed solutions for the limitations.
+One frequently discussed limitation is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention.
 A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed.
 \cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI).
-A study from Virginia Polytechnic Institute and State University examined how, even with the use of features such as  \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}}, touch screens still do not provide sufficient support for IBSVI.
-These users were only able to ``develop a spatial mental model for the interface or the screen through dead reckoning" since there is an ``absence of any landmark other than the boundary of the device." \cite{El-Glaly:2013:TTF:2460625.2460665}
+A study from Virginia Polytechnic Institute and State University examined how touch screens---even those that include features such as  \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}}---do not provide sufficient support for IBSVI.
+These users were only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' since there was an ``absence of any landmark other than the boundary of the device.'' \cite{El-Glaly:2013:TTF:2460625.2460665}
 
-One purposed solution to this limitation is a physical overlay, developed by researchers specifically to help IBSVI engage their spatial cognition, perception, and sensing resources while interacting with touch screens.
-\cite{El-Glaly:2013:TTF:2460625.2460665} The tactile overlay covers the touch device screen.
-An IBSVI can use the tactile patterns of the overlay to locate the text, icons, or buttons in the touch device interface.
+One purposed solution to this limitation is the use of a tactile overlay.
+(See Figure~\ref{tactile-grid}) One such overlay was recently developed by researchers specifically to help IBSVI engage their spatial cognition, perception, and sensing resources while interacting with touch screens.
+\cite{El-Glaly:2013:TTF:2460625.2460665} The tactile overlay covers the touch device screen and IBSVI can use the tactile patterns of the overlay to locate the text, icons, or buttons in the touch device interface.
 A study examining the efficacy of this physical overlay illustrated that IBSVI who used it were generally more efficient than those using voice-over technology or `trail-and-error' exploration.
@@
 \includegraphics[width=3in]{TactileGrid.jpg} 
-\caption{Tactile grid}
-\label{figure-sample}
+\caption{A Tactile Grid}
+\label{tactile-grid}
 \end{figure}
 
-Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution."\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
+Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution.''\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
 A notable technique was created by Schneiderman, Sears, and colleagues to specifically address this problem.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
@@
 Many studies have purposed using a stylus (pen) to interact with touch screens instead of using a bare finger.
-A stylus has better ``resolution" than a finger tip and recent studies have demonstrated that pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} A study by Forlines and Balakrishnan, which evaluated multiple forms of stylus input, articulated that one drawback of this approach is the occlusion of other elements on the screen by the pen.\cite{Forlines}
+A stylus has better ``resolution'' than a finger tip and recent studies have demonstrated that pen input is emerging as a promising interaction modality for touch screens.\cite{Bi} A study by Forlines and Balakrishnan, which evaluated multiple forms of stylus input, articulated that one drawback of this approach is the occlusion of other elements on the screen by the pen.\cite{Forlines}
 
@@
 It is possible to use zooming to enlarge the information space of the touchscreen enough so that the user can comfortably point at the target with a bare finger.
-However, a study by P\"{a}r-Anders Albinsson and Shumin Zhai asserted that zooming has a fundamental drawback since ``when zoomed to a sub area of interest, one loses the contextual global view that can be important for the user's task."\cite{Albinsson}
+However, a study by P\''{a}r-Anders Albinsson and Shumin Zhai asserted that zooming has a fundamental drawback since ``when zoomed to a sub area of interest, one loses the contextual global view that can be important for the user's task.''\cite{Albinsson}
 
@@
 
-Studies suggest that there also exist usability advantages to using the graphically rendered, non-tactile buttons on touch screens, which are often called \textit{soft buttons}.
-A study by Seungyon Lee and Shumin Zhai articulates that some advantages to soft buttons compared to their hard counterparts include their ability to appear or disappear according to the interaction context, their ability to change size depending on screen-space availability, and the ease with which they can be stylized or updated.\cite{Lee:2009:PTS:1518701.1518750} The results from the experiments presented in Lee and Zhai's study suggest that for tasks such as  dialing phone numbers and using calculator functions soft buttons ``can offer a level of performance that is similar, and in some cases even superior, to the level of performance offered by typical hard buttons embedded in today's handheld devices." \cite{Lee:2009:PTS:1518701.1518750}
+Studies suggest that there also exist usability advantages to using the graphically rendered, non-tactile buttons on touch screens, which are referred to as \textit{soft buttons}.
+A study by Seungyon Lee and Shumin Zhai outlined some advantages of soft buttons compared to their hard counterparts, including their ability to appear or disappear according to the interaction context, their ability to change size depending on screen-space availability, and the ease with which they can be stylized or updated.\cite{Lee:2009:PTS:1518701.1518750} The results from the experiments presented in Lee and Zhai's study suggest that for tasks such as  dialing phone numbers and using calculator functions, soft buttons ``can offer a level of performance that is similar, and in some cases even superior, to the level of performance offered by typical hard buttons embedded in today's handheld devices.'' \cite{Lee:2009:PTS:1518701.1518750}
 
 \section{Methods}
-The prior work in the field proves that there is a variety of different advantages, limitations, and proposed solutions to those limitations that can be examined when evaluating the usability of touchscreen devices.
+The prior work in the field illustrates that there is a variety of different advantages, limitations, and proposed solutions that can be examined when evaluating the usability of touchscreen devices.
 Since this paper is too limited in both scope and depth to discuss each advantage and limitation, the paper will focus exclusively on the limitation of the minimal tactile feedback provided by touchscreen devices.
 
-\subsection{Prevailing Set of Views}
-Most research pertaining to this limitation focuses on how the lack of tactile feedback affects the usability of devices for Individuals with Blindness or Severe Visual Impairment---as previously touched on in this study.
-The prevailing set of views is that the lack of accessibility for IBSVI is the main problem with touch screen devices.
+\subsection{Prevailing Views}
+Most research pertaining to this limitation focuses on how the lack of tactile feedback affects the usability of devices for Individuals with Blindness or Severe Visual Impairment.
 Touch screen interfaces have been popular for over 20 years, and concerns about touch screen accessibility for IBSVI have remained active throughout this time.
-\cite{Buxton:1986:HID:22339.22386} Thus, a significant portion of the proposed solutions for this limitation are developed specifically to address the needs of IBSVI.
+\cite{Buxton:1986:HID:22339.22386} The prevailing view in the literature is that the lack of accessibility for IBSVI is the main problem with touch screen devices.
+A majority of solutions developed to tackle this usability limitation were created specifically to address the needs of IBSVI.
 
 Previous research has explored techniques that make touch screens more accessible to IBSVI by modifying the hardware.
-This paper mentioned one such technique when explaining physical overlays.\cite{El-Glaly:2013:TTF:2460625.2460665} However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people." \cite{Kane:2011:AOI:2047196.2047232}  The study asserts that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
+This paper previously mentioned one such technique when discussing a physical overlay.
+(See Section~\ref{background}) However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people.'' \cite{Kane:2011:AOI:2047196.2047232}  The study asserted that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
 
@@
 As access overlays are entirely software-based, they do not require alterations to the underlying touch screen hardware.
-Results from a recent study suggested that access overlays enabled users with visual impairments to locate on-screen targets faster than traditional techniques, improved their spatial understanding, and were preferred.\cite{Kane:2011:AOI:2047196.2047232} However,  access overlays have also not been widely adopted.
+Results from a recent study proved that access overlays enabled users with visual impairments to locate on-screen targets faster than traditional techniques, improved their spatial understanding, and were preferred.\cite{Kane:2011:AOI:2047196.2047232} However,  like physical overlays, access overlays have yet to become widely adopted.
 
@@
 For example, Slide Rule \cite{Kane:2008:SRM:1414471.1414487} used multi-touch gestures to allow blind users to browse and explore content on a touch screen-based smartphone.
+(see Figure~\ref{slide-rule}).
+Similar interfaces now appear in consumer devices such as Apple's iPhone\cite{AppleMultiTouch} and Android-based smartphones \cite{AndroidMultiTouch} as well.
 
@@
 \cite{Kane:2008:SRM:1414471.1414487} }
-\label{figure-sample}
+\label{slide-rule}
 \end{figure}
 
-Slide Rule's design is ``based on interviews with blind mobile device users and on user-centered design with blind people." \cite{Kane:2008:SRM:1414471.1414487} The results from a study pertaining to the usability of Slide Rule showed that users were able to complete tasks more quickly with Slide Rule than with a button-based mobile screen reader.
-However, they did make more errors.
-\cite{Kane:2008:SRM:1414471.1414487} Similar interfaces now appear in consumer devices such as Apple's iPhone\cite{AppleMultiTouch} and Android-based smartphones.\cite{AndroidMultiTouch}
+Slide Rule's design is ``based on interviews with blind mobile device users and on user-centered design with blind people.'' \cite{Kane:2008:SRM:1414471.1414487} The results from a study pertaining to the usability of Slide Rule showed that users were able to complete tasks more quickly with Slide Rule than with a button-based mobile screen reader.
+However, the users did make more errors.
+\cite{Kane:2008:SRM:1414471.1414487}
 
@@
 The significance of all of these proposed solutions is that they were each developed by researchers specifically to address the needs of Individuals with Blindness or Severe Visual Impairment.
-Many studies, including each study mentioned in this section, only tested solutions improved usability for IBSVI.
+Many studies, including each study mentioned in this section, only tested how the solutions improved usability for IBSVI.
 The extent to which these solutions helped sighted users was not examined.
@@
 There exists some relevant emerging technology that makes touch screen devices more tactile.
-Though these emerging devices may not have been the direct result of research pertaining to usability, they do provide some noteworthy possible solutions to the usability limitation.
+Though these emerging devices may not have been created with the usability limitation directly in mind, the devices still appear to be possible solutions to the problem at hand.
 
-Tactus Technology, a company in Fremont, California.
+One noteworthy device comes from Tactus Technology, a company in Fremont, California.
 Employees from Tactus Technology are creating a keyboard with shape-shifting keys that pop up from the screen's surface when needed and recede again when no longer necessary.\cite{Tactus} Craig Ciesla, the co-founder of the company, said that the keyboard will be offered later this year.
+(see Figure~\ref{tactus-keyboard}).
 
@@
 \caption{A Tactus keyboard}
-\label{figure-sample}
+\label{tactus-keyboard}
 \end{figure}
-Though literature on the subject of touchscreen display devices have presented many limitations, the most relevant and important seems to be the lack of tactile features and the requirement for heightened visual attention.
-For people who are blind, accessing touch screen interfaces is nearly impossible.
-Some accessibility features are beginning to be developed though.
 
 \section{Discussion}
-Research discusses usability for IBSVI but does has not delved into technology that can be effective for the sighted and IBSVI alike.
-The principle of feedback
-feedback ---sending back to the user information about what action has actually been done, what result has been accomplished--- is a well-known concept in the science of control and information theory\cite{Norman02} "In the good old days of the telephone", "telephones were designed with much more care and concern for the user.
-Designers at the Bell Telephone Laboratories worried a lot about feedback.
-The push buttons were designed to give an appropriate feel---tactile feedback.
-When a button was pushed, a tone was fed back into the earpiece so the user could tell that the button had been properly pushed.
-When the phone call was being connected, clicks, tones, and other noises gave the user feedback about the progress of the call." ``All this changed." ``Why are modern telephone systems so difficult to use?
-Basically, the problem is that the systems have more features and less feedback"
+Prior work pertaining to the usability of touch screen devices has focused primarily on how Individuals with Blindness or Severe Visual Impairment (IBSVI) interact with the screens.
+Many proposed solutions to the lack of tactile feedback have been developed specifically for IBSVI.
+There also exist certain studies that examine usability of touch screen devices for sighted users.
+\cite{Tinwala:2010:ETE:1868914.1868972} However, there does not seem to be any research in the field that examines how tactile feedback affects usability for IBSVI \textit{and} sighted users alike.
+
+Usability research evaluating touchscreen efficacy for both IBSVI and sighted users still needs to be covered in the literature.
+I completely agree that the lack of accessibility for IBSVI is a major issue with touchscreen devices.
+However, providing more tactile feedback can increase usability not just for IBSVI, but also for sighted users.
+In Donald Norman's book, \textit{The Design of Everyday Things}, Norman touches on this principle of feedback.
+Norman describes how telephones in the past ``were designed with much more care and concern for the user'' as ``the push buttons were designed to give an appropriate feel---tactile feedback.''\cite{Norman02} Norman asserts that modern telephone systems are ``so difficult to use'' since they have ``more features and less feedback.''\cite{Norman02} As tactile feedback is significant for IBSVI and sighted users alike, research should examine how both groups of users respond to proposed solutions.
+I think that this could possibly lead to a solution that becomes more widely accepted.
 
 \section{Conclusions}
+This study examined the usability of touchscreen devices by looking specifically at the limitation of the lack of tactile feedback.
+The literature focuses primarily on how this limitation reduces learnability and efficiency of touchscreen devices for Individuals with Blindness or Severe Visual Impairment.
+Many proposed solutions to this usability limitation are centered around the needs of IBSVI.
+Though there exist some studies pertaining to how this limitation affects usability for sighted users, the findings are not as comprehensive.
+There also exist a few devices that were not necessarily centered around the needs of IBSVI, such as the Tactus keyboard.
+The main concept missing from this field of study is solutions to the limitation that are effective for IBSVI and sighted users alike.
 
-Wrap up your paper with an ``executive summary'' of the paper itself, reiterating its subject and its major points.  If you want examples, just look at the conclusions from the literature.
 \clearpage
@@
 \end{document}
-%%%%%%%%%%%%% end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-10-22 mental-model-paper/talk-questionnaire.tex
questionnaire work

\documentclass{article}
\usepackage{lipsum}
\newcounter{question}
\setcounter{question}{0}
\begin{document}

\newcommand\Que[1]{%
   \leavevmode\par
   \stepcounter{question}
   \noindent
   \thequestion. Q --- #1\par}

\newcommand\Ans[2][]{%
    \leavevmode\par\noindent
   {\leftskip37pt
    A --- \textbf{#1}#2\par}}

\par
\noindent
\textit{Rachel Rivera}

\Que{Give at least two specific examples of objects or behaviors that cognitive psychologists study.}
\Ans{ \\ 1. The behaviors of patients with localized brain injury. \\ 2. The behaviors of split-brain patients.}

\Que{How do these subjects of study relate to what interaction design studies?}
\Ans{These subjects relate to a plethora of different interaction design topics. One prominent topic that is relevant in both fields is the notion of \textit{learnability}. }

\Que{Give at least two specific \textit{methods} used by cognitive psychologists to study or learn about their subjects.}
\Ans{\\ 1. Having split-brain patients draw the same object twice--- once with the left hand and once with the right. \\ 2. Examining what each hemisphere can accomplish in isolation by testing split-brain subjects' responses to touching objects versus saying the names of the objects. }

\Que{How do these methods of study relate to how interaction design research is performed?}
\Ans{ The manner in which these methods isolate portions of the brain and examine the abilities of those portions in isolation parallels the general manner in which interaction design research is performed. }
\end{document}

### 2014-10-26 rgp/rpg.js
landing page

@@ -1,0 +1,27 @@
+$(function() {
+    $.getJSON(
+        "http://lmu-diabolical.appspot.com/characters",
+        function (characters) {
+            $("tbody").append(characters.map(function (character) {
+    /*
+                return $("<tr></tr>")
+                    .append($("<td></td>")
+                        .append($("<img/>")
+                            .attr({ src: "http://icons.iconarchive.com/icons/hopstarter/superhero-avatar/256/Avengers-Black-Widow-icon.png" })
+                            .height(24)))
+                    .append($("<td></td>").text(character.name))
+                    .append($("<td></td>").text(character.classType))
+                    .append($("<td></td>").text(character.gender))
+                    .append($("<td></td>").text(character.level));
+    */
+                var tr = $(".character-template").clone();
+                tr.find(".name").text(character.name);
+                tr.find(".class").text(character.classType);
+                tr.find(".gender").text(character.gender);
+                tr.find(".level").text(character.level);
+                return tr;
+            }));
+        }
+    );
+})
+

### 2014-10-27 rgp/search.js
add responsive search

@@ -1,0 +1,9 @@
+$('#search').keyup(function() {
+	var searchField = $('#search').val();
+	var searchRegExp = new RegExp(searchField, "i");
+    $(".character-template").each(function(index, row){
+    	var allCells = $(row).find('td');
+    	var nameCell = allCells[0];
+    	($(nameCell).text().search(searchRegExp) !== -1) ? $(row).show() : $(row).hide();
+    });
+});

### 2014-10-27 rgp/rpg.js
re-naming for clarity

@@ -3,17 +3,22 @@ $(function() {
         "http://lmu-diabolical.appspot.com/characters",
         function (characters) {
             $("tbody").append(characters.map(function (character) {
-    /*
-                return $("<tr></tr>")
-                    .append($("<td></td>")
-                        .append($("<img/>")
-                            .attr({ src: "http://icons.iconarchive.com/icons/hopstarter/superhero-avatar/256/Avengers-Black-Widow-icon.png" })
-                            .height(24)))
-                    .append($("<td></td>").text(character.name))
-                    .append($("<td></td>").text(character.classType))
-                    .append($("<td></td>").text(character.gender))
-                    .append($("<td></td>").text(character.level));
-    */
+                var tr = $(".character-template").clone();
+                tr.find(".name").text(character.name);
+                tr.find(".class").text(character.classType);
+                tr.find(".gender").text(character.gender);
+                tr.find(".level").text(character.level);
+                return tr;
+            }));
+        }
+    );
+})
+
+$(function() {
+    $.getJSON(
+        "http://lmu-diabolical.appspot.com/characters",
+        function (characters) {
+            $("tbody").append(characters.map(function (character) {
                 var tr = $(".character-template").clone();
                 tr.find(".name").text(character.name);
                 tr.find(".class").text(character.classType);

### 2014-10-27 rgp/rpg.js
fix slowness bug

@@ -2,6 +2,7 @@ $(function() {
     $.getJSON(
         "http://lmu-diabolical.appspot.com/characters",
         function (characters) {
+            console.log("initial GET with characters " + characters);
             $("tbody").append(characters.map(function (character) {
                 var tr = $(".character-template").clone();
                 tr.find(".name").text(character.name);
@@ -12,21 +13,5 @@ $(function() {
             }));
         }
     );
-})
-
-$(function() {
-    $.getJSON(
-        "http://lmu-diabolical.appspot.com/characters",
-        function (characters) {
-            $("tbody").append(characters.map(function (character) {
-                var tr = $(".character-template").clone();
-                tr.find(".name").text(character.name);
-                tr.find(".class").text(character.classType);
-                tr.find(".gender").text(character.gender);
-                tr.find(".level").text(character.level);
-                return tr;
-            }));
-        }
-    );
-})
+});
 

### 2014-11-18 rpg/rpg.js
fixes

@@ -4,16 +4,18 @@ $(function() {
         "http://lmu-diabolical.appspot.com/characters",
         function (characters) {
             $("tbody").append(characters.map(function (character) {
-                var filledInTemplate = fillInCharacterInfo('.character-template', character);
+                var tableOfCharacters     = fillTable('.character-template', character);
+                var editModalForCharacter = $('.edit-character-modal-template').clone()
 
-                filledInTemplate.find(".edit").bind("click", function(){
-                    showEditCharacterModal(character);
+                tableOfCharacters.find(".edit").bind("click", function() {
+                    editModalForCharacter = fillEditModal(editModalForCharacter, character);
+                    edit(character, editModalForCharacter);
                 });
 
-                filledInTemplate.find(".delete").bind("click", function(){
+                tableOfCharacters.find(".delete").bind("click", function(){
                     showDeleteCharacterModal(character);
                 });
-                return filledInTemplate;
+                return tableOfCharacters;
             }));
         }
     );
@@ -52,143 +54,141 @@ $(function() {
     };
 
     var showDeleteCharacterModal = function(character) {
-        $('#deleteModal').find('.modal-message').text("Are you sure you want to delete " + character.name + "?");
-        $('#deleteModal').modal('show');
-        var deleteCharBtn = $('#deleteModal').find('.delete-character');
-        $(deleteCharBtn).bind('click', function(){
-          $(deleteCharBtn).button('loading');
-          setTimeout(function () {
-            $(deleteCharBtn).button('reset');
-            $(deleteCharBtn).text('Try Again?');
-          }, 5000);
-          deleteCharacter(character);
+        var deleteModal = $('.delete-modal-template');
+        $(deleteModal).find('.modal-message')
+          .text('Are you sure you want to delete ' + character.name + '?');
+
+        var deleteBtn = $(deleteModal).find('.btn-delete-character');
+        $(deleteBtn).bind('click', function(){
+            deleteCharacter(character, deleteModal);
+            $(deleteBtn).button('loading');
+            setTimeout(function (){
+                $(deleteBtn).button('reset');
+            }, 5000);
         })
 
+        $(deleteModal).modal('show');
 
-
-        // BootstrapDialog.show({
-        //     message: 'Are you sure you want to delete this character?',
-        //     buttons: [{
-        //         type: BootstrapDialog.TYPE_DANGER,
-        //         icon: 'glyphicon glyphicon-remove',
-        //         label: 'Delete',
[…]

### 2014-11-18 rpg/search.js

@@ -1,9 +1,11 @@
-$('#search').keyup(function() {
-	var searchField = $('#search').val();
-	var searchRegExp = new RegExp(searchField, "i");
-    $(".character-template").each(function(index, row){
-    	var allCells = $(row).find('td');
-    	var nameCell = allCells[0];
-    	($(nameCell).text().search(searchRegExp) !== -1) ? $(row).show() : $(row).hide();
+$(function(){
+    $('#search').keyup(function() {
+        var searchField = $('#search').val();
+        var searchRegExp = new RegExp(searchField, "i");
+        $(".character-template").each(function(index, row){
+            var allCells = $(row).find('td');
+            var nameCell = allCells[0];
+            ($(nameCell).text().search(searchRegExp) !== -1) ? $(row).show() : $(row).hide();
+        });
     });
 });

### 2014-11-18 rpg/rpg.js
GET before PUT

@@ -3,6 +3,7 @@ $(function() {
     window.onload = $.getJSON(
         "http://lmu-diabolical.appspot.com/characters",
         function (characters) {
+            console.log("initial GET " + JSON.stringify(characters));
             $("tbody").append(characters.map(function (character) {
                 var tableOfCharacters     = fillTable('.character-template', character);
                 var editModalForCharacter = $('.edit-character-modal-template').clone()
@@ -109,7 +110,7 @@ $(function() {
         var editBtn = $('.edit-check-modal').find('.btn-edit');
         $(editBtn).bind("click", function(){
           resetModal(editModalForCharacter);
-          sendCharacterUpdates(originalCharacter, updatedCharacter);
+          getCharacter(originalCharacter, updatedCharacter);
           $(editBtn).button('loading');
           setTimeout(function () {
             $(editBtn).button('reset');
@@ -126,22 +127,33 @@ $(function() {
     var sendCharacterUpdates = function(originalCharacter, updatedCharacter){
         $.ajax({
             type: 'PUT',
-            url: "http://www.lmu-diabolical.appspot.com/characters/" + updatedCharacter.id,
-            data: JSON.stringify(updatedCharacter),
+            url: "http://lmu-diabolical.appspot.com//characters/" + originalCharacter.id,
             contentType: "application/json",
             dataType: "json",
             accept: "application/json",
+            data: JSON.stringify(updatedCharacter),
             success: function (data, textStatus, jqXHR) {
                 console.log('data' + data);
                 console.log('textStatus: ' + textStatus);
+                console.log('jqXHR ' + JSON.stringify(jqXHR));
                 updateOriginalCharacter(originalCharacter, updatedCharacter);
                 $('.edit-check-modal').modal('hide');
                 $('#successful-edit').show();
                 setTimeout(function() {
                     $('#successful-edit').hide();
                 }, 5000);
-            }
+            },
         });
+    }
+
+    var getCharacter = function(originalCharacter, updatedCharacter){
+        var url = "http://lmu-diabolical.appspot.com/characters/" + originalCharacter.id;
+        $.getJSON(
+            url,
+            function (character) {
+                sendCharacterUpdates(character, updatedCharacter);
+            // Do something with the character.
+            });
     }
 
 

### 2014-11-19 rpg/rpg.js
fixes with edit/delete

@@ -5,17 +5,18 @@ $(function() {
         function (characters) {
             console.log("initial GET " + JSON.stringify(characters));
             $("tbody").append(characters.map(function (character) {
-                var tableOfCharacters     = fillTable('.character-template', character);
-                var editModalForCharacter = $('.edit-character-modal-template').clone()
+                var tableOfCharacters     = fillTableCharacterInfo('.character-template', character);
+                var editModalForCharacter = $('.edit-character-modal-template').clone();
+                var deleteModalForCharacter = $('.delete-character-modal-template').clone();
 
-                tableOfCharacters.find(".edit").bind("click", function() {
-                    editModalForCharacter = fillEditModal(editModalForCharacter, character);
-                    edit(character, editModalForCharacter);
+                tableOfCharacters.find(".btn-edit").bind("click", function() {
+                    editCharacter(character.id, editModalForCharacter);
                 });
 
-                tableOfCharacters.find(".delete").bind("click", function(){
-                    showDeleteCharacterModal(character);
+                tableOfCharacters.find(".btn-remove").bind("click", function(){
+                    deleteCharacter(character.id, deleteModalForCharacter);
                 });
+
                 return tableOfCharacters;
             }));
         }
@@ -54,9 +55,8 @@ $(function() {
         })
     };
 
-    var showDeleteCharacterModal = function(character) {
-        var deleteModal = $('.delete-modal-template');
-        $(deleteModal).find('.modal-message')
+    var deleteCharacter = function(character) {
+        $('.remove-character-modal-template').find('.modal-message')
           .text('Are you sure you want to delete ' + character.name + '?');
 
         var deleteBtn = $(deleteModal).find('.btn-delete-character');
@@ -72,14 +72,15 @@ $(function() {
 
     }
 
-    var deleteCharacter = function(character, deleteCharacterModal) {
+    var sendDeleteRequest = function(character, deleteCharacterModal) {
         $.ajax({
             type: 'DELETE',
             url: "http://lmu-diabolical.appspot.com/characters/" + character.id,
             success: function (data, textStatus, jqXHR) {
                 $(deleteCharacterModal).modal('hide');
                 updateTableWithDelete(character);
-                $('#successful-delete').show()
+                $('#successful-delete').show();
+                $('html, body').animate({ scrollTop: 0 }, 0);
                 setTimeout(function() {
                     $('#successful-delete').hide();
[…]

### 2014-11-19 rpg/rpg.js
creating characters

@@ -22,9 +22,66 @@ $(function() {
         }
     );
 
-    $(".createCharacter").bind("click", function(){
-        showCreateCharacterModal();
-    })
+    $(".btn-create-character").bind("click", function(){
+        createCharacter();
+    });
+
+    $(".btn-spawn-item").bind("click", function(){
+        spawnItem();
+    });
+
+    var spawnItem = function(){
+        console.log("blah");
+    }
+
+
+
+    var createCharacter = function() {
+        var createCharacterModal = $('.create-character-modal-template');
+        createCharacterModal.modal('show');
+        var createBtn = $(createCharacterModal).find('.btn-create-character');
+        createBtn.bind('click', function(){
+            console.log("clicked");
+            sendPostRequest({
+                name: createCharacterModal.find('#new-name').val(),
+                classType: createCharacterModal.find('#new-class').val(),
+                gender: createCharacterModal.find('#new-gender').val().toUpperCase(),
+                level: createCharacterModal.find('#new-level').val()
+            });
+            $(createBtn).button('loading');
+            setTimeout(function (){
+                $(createBtn).button('reset');
+            }, 5000);
+        })
+    }
+
+    var sendPostRequest = function(newCharacter){
+        console.log('sending post request');
+        $.ajax({
+            type: 'POST',
+            url: "http://lmu-diabolical.appspot.com/characters",
+            contentType: "application/json",
+            dataType: "json",
+            accept: "application/json",
+            data: JSON.stringify(newCharacter),
+            complete: function (jqXHR, textStatus) {
+                updateTableWithAddition(newCharacter);
+                resetModal('.create-character-modal-template');
+                $('.create-character-modal-template').modal('hide');
+                $('html, body').animate({ scrollTop: 0 }, 0);
+                $('#successful-create').show();
+                setTimeout(function() {
+                    $('#successful-create').hide();
+                }, 5000);
+            }
+        });
+    }
+
+    var updateTableWithAddition = function(newCharacter){
+        var row = $('tbody').children('tr').eq(1).clone();
+        var characterRow = fillTableCharacterInfo(row, newCharacter);
+        $('tbody').prepend(characterRow);
+    }
 
     var showCreateCharacterModal = function(character) {
         BootstrapDialog.show({
@@ -119,8 +176,8 @@ $(function() {
                     $(editModal).modal('hide');
                     var updatedCharacter = getUpdatedCharacter(character, editModal);
[…]

### 2014-11-20 mental-model-paper/paper2.tex
Inline comments and feedback for HW 1021.

@@
 
-\abstract{
+\begin{abstract} % JD: My bad; this was the right way.
 This study aims to examine how interaction design concepts specifically map to the usability of modern display technology.
@@
 This limitation and each proposed solution has a variety of implications with respect to learnability, efficiency, satisfaction and errors.
-
-}
+\end{abstract}
 
@@
 Electronic screens are devices through which information is transferred between the user and the interface.
-Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still diversifying in a multitude of ways to tackle the limitations that have presented themselves over the years.\cite{Eisenberg} Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
-For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens.\cite{Baudisch} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch}
+Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922 \cite{Cathode}.} they are still diversifying in a multitude of ways to tackle the limitations that have presented themselves over the years \cite{Eisenberg}.
+Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks.
+For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens \cite{Baudisch}.
+Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields \cite{Hirsch}.
+% JD: I won't change it for the rest of the document, but just note that you want your cites *before*
+%     the periods of sentences.
 
@@
 \subsection{Limitations/Proposed Solutions}
-There exists a fair amount of previous work pertaining to the usability of touch sensitive screens.
+There exists a fair amount of previous work pertaining to the usability of touch sensitive screens
+\cite{Tinwala:2010:ETE:1868914.1868972,El-Glaly:2013:TTF:2460625.2460665}. % JD: etc.
+% JD: A sentence like the above typically concludes with a list of citations---representing the
+%     aforementioned "fair amount of previous work." :)  Then these citations are cherry-picked
+%     and discussed in further detail..the way you do next.
 The literature often evaluates usability by examining the limitations of these screen devices, as well as the proposed solutions for the limitations.
@@
 Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution.''\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger.
-A notable technique was created by Schneiderman, Sears, and colleagues to specifically address this problem.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
+A notable technique was created by Shneiderman, Sears, and colleagues to specifically address this problem.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen.
 The user drags the cursor to a desired target and then lifts the finger to select the targeted object.
@@
 Previous work on the usability of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays.
-Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, suggests a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
-\cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasons.
-For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Schneiderman's assertion that touchscreen display devices are genertally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}
+Ben Shneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, suggests a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc.
+\cite{73754} Many experiments have been conducted that provide evidence supporting Shneiderman's reasons.
+For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Shneiderman's assertion that touchscreen display devices are genertally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}
 
@@
 Previous research has explored techniques that make touch screens more accessible to IBSVI by modifying the hardware.
-This paper previously mentioned one such technique when discussing a physical overlay.
-(See Section~\ref{background}) However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people.'' \cite{Kane:2011:AOI:2047196.2047232}  The study asserted that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
+This paper previously mentioned one such technique when discussing a physical overlay (See Section~\ref{background}).
+% JD: Parentheticals like the one above also typically precede the period.
+However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people.'' \cite{Kane:2011:AOI:2047196.2047232}  The study asserted that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.
+
+% JD: You don't do this very often, but have done it enough times to get noticed:
+%     sometimes you quote your references directly, but the quote itself does not
+%     seem particularly compelling.  i.e., You would have done just as well to
+%     paraphrase the author.  The reason this is preferred is that it gives your
+%     writing a more cohesive voice---yours.  Unless a quote has such compelling
+%     word choice and phrasing that its point loses value when paraphrased, things
+%     generally read more smoothly if you don't disrupt the flow with quotes.
+%
+%     And note that you are still citing your source, so the paraphrasing does not
+%     constitute any form of unfair use.
 
@@
 
+% JD: Nicely done!  Please consider entering this in ULRA.  Additional figures might be
+%     needed, I think, to communicate with a broader audience, but overall this is a
+%     well-investigated, well-executed piece of work.
+
 \bibliography{mybib}{}

### 2014-11-20 rpg/rpg.js
spawn item

@@ -31,7 +31,22 @@ $(function() {
     });
 
     var spawnItem = function(){
-        console.log("blah");
+        $.getJSON(
+            "http://lmu-diabolical.appspot.com/items/spawn",
+            {
+                level: 50,
+                slot: "body"
+            },
+            function (item) {
+                var spawnItemModal = $('.spawn-item-modal-template');
+                for (key in item){
+                    var currentClass = '.' + key;
+                    console.log("current class " + currentClass);
+                    spawnItemModal.find(currentClass).html(item[key]);
+                }
+                spawnItemModal.modal('show');
+            }
+        );
     }
 
 
@@ -246,7 +261,7 @@ $(function() {
     /* Hide removed character from table without having to reload the page */
     var updateTableWithDelete = function(deletedCharacter){
         var rowToHide = '#' + deletedCharacter.id;
-        $(rowToHide).hide();
+        $(rowToHide).remove();
     }
 
     /* Update table without having to reload the page. */

### 2014-11-24 dream-design/dream-design.tex
introduction

\documentclass[11pt]{article}

\usepackage{geometry}
\geometry{letterpaper}

\usepackage{doc}
\usepackage{cite}
\usepackage[margin=1cm]{caption}

\usepackage{url}

\usepackage{graphicx}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{A ``Dream'' Interface Design}
\author{Rachel Rivera}
\date{November 25, 2014}

\begin{document}

\maketitle

\begin{abstract}
This investigation purposes a ``dream'' user interface for touch screen devices. 
\end{abstract}

\pagebreak
\tableofcontents

\pagebreak

\section{System Description}
\label{System Description}

Mobile devices are ubiquitous in contemporary society. The latest figures from the United Nations' telecommunications agency estimate that there are around 6.8 billion cell-phone subscriptions in the world \cite{UNTelecommunications	}. Thus, the investigation and development of usable interfaces for mobile devices is a significant research topic.

Many mobile devices use touchscreen keyboards instead of physical keyboards. This allows for a larger display without increasing the size of the device. Furthermore, touchscreen keyboards have the ability to change their layout based on user input and disappear when not needed. To compensate for the lack of tactile feedback provided by physical keys, touchscreen keyboards often include aural feedback in the form of audible clicks from a speaker. Haptic feedback in the form of device vibrations is often included as well.

Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen mobile devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for others. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) as the users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

I have designed a ``dream'' interface for touchscreen mobile devices that aims to tackle some of these usability issues. My design provides a significant amount of tactile feedback that will be helpful for individuals without visual impairments and IBSVI alike.

\section{Top-Level Design}
\label{background}

\section{Usage Scenarios}

\section{Rationale}

\section{Usability Metric ``Forecast''}
\clearpage

\bibliography{mybib}{}
\bibliographystyle{plain}
\end{document}

\end{document}

### 2014-11-24 dream-design/dream-design.tex
system overview

@@
 
-\section{System Description}
-\label{System Description}
+\section{Introduction}
+\label{Introduction}
 
@@
 A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}.
-The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for others.
-A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) as the users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
+The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users.
+A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) as these users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
 
-I have designed a ``dream'' interface for touchscreen mobile devices that aims to tackle some of these usability issues.
-My design provides a significant amount of tactile feedback that will be helpful for individuals without visual impairments and IBSVI alike.
+Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses some of these usability issues.
+
+\section{System Description}
+\label{System Description}
+
+In this investigation, I propose a ``dream'' interface design for touchscreen mobile devices that provides a significant amount of tactile feedback.
+This feedback is designed in a way that can be helpful for individuals with and without visual impairments alike.
+This design was developed with the usability metrics of learnability, efficiency, and errors in mind.
+
+This design makes use of technology that is currently being created by Tactus, a company in Fremont California \cite{Tactus}.
+Tactus brings real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus}).
+Small fluid channels are routed throughout the Tactile Layer and enable fluid ot expand the top polymer layer to create the physical buttons \cite{Tactus}.
+
+Although this technology from Tactus is extremely bleeding-edge, it has already received a fair amount of recognition \cite{CNN, I-Zone, PCMag}.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=4.5in]{tactus1.jpg} 
+\caption{Tactus Technology keyboard}
+\label{tactus}
+\end{figure}
 
 \section{Top-Level Design}
+The design aims to assist IBSVI engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+The tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback that informs the users of \textit{what} is at the location on the screen while the tactile buttons heps the user know \textit{where} it is.
 \label{background}

### 2014-11-24 dream-design/dream-design.tex
add wireframe

@@
 This feedback is designed in a way that can be helpful for individuals with and without visual impairments alike.
-This design was developed with the usability metrics of learnability, efficiency, and errors in mind.
+This design was developed with the usability metrics of learnability, efficiency, and error rate in mind.
 
-This design makes use of technology that is currently being created by Tactus, a company in Fremont California \cite{Tactus}.
+My ``dream'' interface design makes use of technology that is currently being created by Tactus, a company in Fremont, California \cite{Tactus}.
 Tactus brings real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus}).
-Small fluid channels are routed throughout the Tactile Layer and enable fluid ot expand the top polymer layer to create the physical buttons \cite{Tactus}.
+Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.
 
-Although this technology from Tactus is extremely bleeding-edge, it has already received a fair amount of recognition \cite{CNN, I-Zone, PCMag}.
+Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition and positive feedback \cite{CNN, I-Zone, PCMag, Wired}.
+Reviewers articulate how the technology seems to be ``downright magical'' \cite{CNN}.
+Though the look and feel of Tactus technology is totally futuristic, the technology is supposed to appear in consumer devices early next year \cite{Tactus}.
+Companies will also be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote.
+This ``dream'' user interface design makes use of the fact that the possibilities for the panel are varied.
 
@@
 \section{Top-Level Design}
-The design aims to assist IBSVI engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+The design of this ``dream'' interface aims to assist users with and without visual impairments alike.
+The design will use Tactus technology to create a grid-like system for IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1in]{wireframe.png} 
+\caption{Tactus Technology keyboard}
+\label{tactus}
+\end{figure}
+
 The tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback that informs the users of \textit{what} is at the location on the screen while the tactile buttons heps the user know \textit{where} it is.

### 2014-11-25 dream-design/dream-design.tex
more top-level overview

@@
 
-Many mobile devices use touchscreen keyboards instead of physical keyboards.
+Many mobile devices utilize touch screens instead of physical keyboards or buttons.
 This allows for a larger display without increasing the size of the device.
-Furthermore, touchscreen keyboards have the ability to change their layout based on user input and disappear when not needed.
-To compensate for the lack of tactile feedback provided by physical keys, touchscreen keyboards often include aural feedback in the form of audible clicks from a speaker.
+Furthermore, touchscreen keyboards and buttons have the ability to change their layout based on user input and disappear when not needed.
+To compensate for the lack of tactile feedback provided by physical keys and buttons, touchscreen devices often include aural feedback in the form of audible clicks from a speaker.
 Haptic feedback in the form of device vibrations is often included as well.
@@
 
-In this investigation, I propose a ``dream'' interface design for touchscreen mobile devices that provides a significant amount of tactile feedback.
+In this investigation, I propose a ``dream'' interface design for touchscreen mobile devices that focuses on providing as much as feedback as possible.
 This feedback is designed in a way that can be helpful for individuals with and without visual impairments alike.
-This design was developed with the usability metrics of learnability, efficiency, and error rate in mind.
+The design in its entirety was created with the usability metrics of learnability, efficiency, and error rate in mind.
 
-My ``dream'' interface design makes use of technology that is currently being created by Tactus, a company in Fremont, California \cite{Tactus}.
-Tactus brings real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus}).
+My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology, a company in Fremont, California \cite{Tactus}.
+Tactus Technology creates real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus}).
 Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.
 
-Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition and positive feedback \cite{CNN, I-Zone, PCMag, Wired}.
-Reviewers articulate how the technology seems to be ``downright magical'' \cite{CNN}.
-Though the look and feel of Tactus technology is totally futuristic, the technology is supposed to appear in consumer devices early next year \cite{Tactus}.
+Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition and awards \cite{CNN, I-Zone, PCMag, Wired}.
+Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}.
+Though the look and feel of Tatctus technology is totally futuristic, the technology is already beginning to appear in consumer devices.
+In 2013, Touch Revolution, the ``largest-volume glass projected capacitive multi-touch screen manufacturer in the world'', announced a partnership with Tactus Technology \cite{TactusAvailability}.
+The technologies are being combined and manufactured into consumer devices today.
 Companies will also be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote.
@@
 \section{Top-Level Design}
-The design of this ``dream'' interface aims to assist users with and without visual impairments alike.
-The design will use Tactus technology to create a grid-like system for IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+The design of this ``dream'' interface aims to assist users with visual impairments and users without visual impairments alike.
+The design includes an optional physical grid, tactile buttons of varying sizes, and auditory feedback.
+
+\subsection{Grid-like Tactile Support}
+The design will use Tactus technology to create a physical grid-like layout on the screen.
+This layout IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
 The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.

### 2014-11-25 dream-design/dream-design.tex
new wireframes

@@
 
-In this investigation, I propose a ``dream'' interface design for touchscreen mobile devices that focuses on providing as much as feedback as possible.
-This feedback is designed in a way that can be helpful for individuals with and without visual impairments alike.
+In this investigation, I propose a ``dream'' interface design for touchscreen mobile devices that focuses on providing the user with as much as feedback as possible.
+This feedback is designed in a way that aims to be helpful for users with visual impairments and users without visual impairments alike.
 The design in its entirety was created with the usability metrics of learnability, efficiency, and error rate in mind.
 
-My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology, a company in Fremont, California \cite{Tactus}.
-Tactus Technology creates real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus}).
+My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}.
+Tactus Technology, a company based in Fremont, California, creates real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus1}).
 Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.
 
-Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition and awards \cite{CNN, I-Zone, PCMag, Wired}.
+Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition as well as several awards \cite{CNN, I-Zone, PCMag, Wired}.
 Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}.
-Though the look and feel of Tatctus technology is totally futuristic, the technology is already beginning to appear in consumer devices.
-In 2013, Touch Revolution, the ``largest-volume glass projected capacitive multi-touch screen manufacturer in the world'', announced a partnership with Tactus Technology \cite{TactusAvailability}.
-The technologies are being combined and manufactured into consumer devices today.
-Companies will also be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote.
-This ``dream'' user interface design makes use of the fact that the possibilities for the panel are varied.
+Though the look and feel of Tatctus technology seems totally futuristic, the technology is already beginning to appear in consumer devices.
+In 2013, Touch Revolution, the largest-volume glass projected capacitive multi-touch screen manufacturer in the world, announced a partnership with Tactus Technology \cite{TactusAvailability}.
+The technologies are being combined and manufactured into consumer devices today \cite{TactusAvailability}.
+Not only is this technology becoming more and more available, but it is becoming more customizable as well.
+Companies will soon be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote \cite{CNN}.
 
@@
 \centering
-\includegraphics[width=4.5in]{tactus1.jpg} 
+\includegraphics[width=4.5in]{tactus2.jpg} 
 \caption{Tactus Technology keyboard}
-\label{tactus}
+\label{tactus1}
 \end{figure}
 
+My ``dream'' user interface design makes use of the tactile feedback as well as the flexibility that this technology from Tactus provides.
+
 \section{Top-Level Design}
-The design of this ``dream'' interface aims to assist users with visual impairments and users without visual impairments alike.
-The design includes an optional physical grid, tactile buttons of varying sizes, and auditory feedback.
+The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike.
+The three main features designed to accomplish this goal are the optional physical grid, the tactile buttons of varying sizes, and the auditory feedback.
 
-\subsection{Grid-like Tactile Support}
-The design will use Tactus technology to create a physical grid-like layout on the screen.
-This layout IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
-The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.
+\subsection{Optional Physical Grid}
+The design incorporates Tactus technology to pa physical grid-like layout on the screen.
 
@@
 \centering
-\includegraphics[width=1in]{wireframe.png} 
-\caption{Tactus Technology keyboard}
-\label{tactus}
+\includegraphics[width=1.5in]{wireframe-grid.png} 
+\caption{Example Grid Layout}
+\label{wireframe-grid}
 \end{figure}
@@
 The tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback that informs the users of \textit{what} is at the location on the screen while the tactile buttons heps the user know \textit{where} it is.
-\label{background}
+
+\subsection{Tactile Buttons}
+
+\subsection{Auditory Feedback}
 
@@
 \section{Rationale}
+The design incorporates Tactus technology to pa physical grid-like layout on the screen.
+This layout IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.
 

### 2014-11-25 dream-design/dream-design.tex
finish top-level design

@@
 \usepackage{geometry}
+\usepackage{subcaption} 
 \geometry{letterpaper}
@@
 The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike.
-The three main features designed to accomplish this goal are the optional physical grid, the tactile buttons of varying sizes, and the auditory feedback.
+The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the tactile buttons, and the auditory feedback.
 
 \subsection{Optional Physical Grid}
-The design incorporates Tactus technology to pa physical grid-like layout on the screen.
+The design incorporates Tactus technology to provide an optional tactile grid for the device (see Figure~\ref{wireframe-grid}).
+The grid layout would consist of a set of three-dimensional horizontal lines and vertical lines, which could be enabled by the user in the device settings .
+Both sets of lines are arranged with equal space between each line.
+The main purpose of this grid is to help IBSVI formulate a mental reference for location awareness to maintain place on the screen.
 
@@
 \centering
-\includegraphics[width=1.5in]{wireframe-grid.png} 
-\caption{Example Grid Layout}
+\includegraphics[width=2in]{wireframe-grid.png} 
+\caption{A simple wireframe demonstrating the grid layout idea.
+Each line in the figure would actually be a three-dimensional line that appears on the screen with the use of microfluidic technology from Tactus.}
 \label{wireframe-grid}
@@
 
-The tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback that informs the users of \textit{what} is at the location on the screen while the tactile buttons heps the user know \textit{where} it is.
+\subsection{Tactile Buttons}
+The design also includes tactile buttons that rise up and recede as needed (see Figure~\ref{buttons}).
+This aspect of the design would also be made possible with the use of technology from Tactus.
+Ideally, the mobile device would be ``intelligent'' about which buttons rose up on the flat screen depending on input from the user.
 
-\subsection{Tactile Buttons}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=6in]{buttons.jpg}
+\caption{A simple representation demonstrating the idea of different subsets of buttons rising up on the same surface at different times.}
+\label{buttons}
+\end{figure}
+
+In instances when less options are available to the user, less buttons should rise up from the screen.
+Similarly, when more options are available to the user, more buttons should rise up from the screen.
+This design aspect should provide more clarity with respect to control flow and what the user is capable of doing at any moment.
 
 \subsection{Auditory Feedback}
+The design of this interface was created in such a way that the tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback.
+This feedback can inform the user of \textit{what} is at the location on the screen whereas the tactile buttons and grid helps the user know \textit{where} it is.
 
@@
 \section{Rationale}
+The tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback that informs the users of \textit{what} is at the location on the screen while the tactile buttons heps the user know \textit{where} it is.
 The design incorporates Tactus technology to pa physical grid-like layout on the screen.

### 2014-11-25 dream-design/dream-design.tex
usage scenarios

@@
 \subsection{Auditory Feedback}
-The design of this interface was created in such a way that the tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback.
+The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide auditory feedback.
 This feedback can inform the user of \textit{what} is at the location on the screen whereas the tactile buttons and grid helps the user know \textit{where} it is.
@@
 \section{Usage Scenarios}
+There exist several usage scenarios for this interface design.
+Sighted users employ vision to tell the user what and where to touch, but this interface lacks accessibility for IBSVI because it lacks physical landmarks and tangible buttons.
+A typical solution is to provide voice over to provide this information, but there is nothing to guide IBSVI to that location in the first place, and there are no landmarks other than dead-reckoning and pure spatial memory to support navigation for the IBSVI.
+For example, the Apple operating system, iOS, has VoiceOver function, and the Android OS has a Talkbalk function.
+iOS VoiceOver reads aloud the icon, the user should double tap on that location without the benefit of vision to ensure that the finger does not stray from the touch point.
+
+The problem with such accessibility modes is that the user can develop a spatial mental model for the interface or screed only through dead reckoning to find VoiceOver locations, in the absence of any landmark other than the boundary of the device.
+The optional tactile grid as well as the three-dimensional buttons that appear with this design solves this problem.
+The additional tactile feedback and physical landmarks can guide IBSVI to interact efficiently and effectively with the touch device and its voice over functionality.
+
+Another use case for this interface pertains specifically to the ``intelligent'' tactile buttons.
+Users often may get overwhelmed by the number of options that are always available to them.
+It also may be frustrating for users to try to perform a task and not be able to.
+To examine a specific example, let's say that the user cannot take any more photos on their phone since they are out of memory.
+In this case, the interface would not even raise the button used to capture photos.
+This is a good indicator to the user that they need to perform some kind of an action before proceeding.
 

### 2014-11-25 dream-design/dream-design.tex
rationale

@@
 The design incorporates Tactus technology to provide an optional tactile grid for the device (see Figure~\ref{wireframe-grid}).
-The grid layout would consist of a set of three-dimensional horizontal lines and vertical lines, which could be enabled by the user in the device settings .
+The grid layout would consist of a set of very thin, three-dimensional horizontal and vertical lines, which could be enabled by the user in the device settings .
 Both sets of lines are arranged with equal space between each line.
@@
 
-\subsection{Tactile Buttons}
+\subsection{``Intelligent'' Tactile Buttons}
 The design also includes tactile buttons that rise up and recede as needed (see Figure~\ref{buttons}).
@@
 \section{Rationale}
-The tactile stuff is intended to operate in conjunction with other technologies that provide auditory feedback that informs the users of \textit{what} is at the location on the screen while the tactile buttons heps the user know \textit{where} it is.
-The design incorporates Tactus technology to pa physical grid-like layout on the screen.
-This layout IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+This layout is designed so that IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
 The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.
 
+Feedback.
+In design, it is important to show the effect of an action.
+Without feedback, one is always wondering whether anything has hap- pened.
+Maybe the button wasn't pushed hard enough; maybe the machine has stopped working; maybe it is doing the wrong thing.
+Without feed- back, we turn equipment off at improper times or restart unnecessarily, losing all our recent work.
+Or we repeat the command and end up having the operation done twice, often to our detriment.
+Feedback is critical (the design of everyday things).
+
+In the good old days of the telephone, before the American tele- phone system was divided among competing companies, before tele- phones were fancy and had so many features, telephones were de- signed with much more care and concern for the user.
+Designers at the Bell Telephone Laboratories worried a lot about feedback.
+The push buttons were designed to give an appropriate feel—tactile feedback.
+When a button was pushed, a tone was fed back into the earpiece so the user could tell that the button had been properly pushed.
+When the phone call was being connected, clicks, tones, and other noises gave the user feedback about the progress of the call.
+And the speaker's voice was always fed back to the earpiece in a carefully controlled amount, because the auditory feedback (called "sidetone") helped the person regulate how loudly to talk.
+All this has changed.
+We now have tele- phones that are much more powerful and often cheaper than those that existed just a few years ago—more function for less money.
+To be fair, these new designs are pushing hard on the paradox of technology: added functionality generally comes along at the price of added com- plexity.
+But that does not justify backward progress.
+Why are the modern telephone systems so difficult to learn and to use?
+Basically, the problem is that the systems have more features and less feedback.
+Suppose all telephones had a small display screen, not unlike the ones on small, inexpensive calculators.
+The display could be used to present, upon the push of a button, a brief menu of all the features of the telephone, one by one.
+When the desired one was encountered, the user would push another button to indicate that it should be invoked.
+If further action was required, the display could tell the person what to do.
+The display could even be auditory, with speech instead of a visual display.
+Only two buttons need be added to the
+telephone: one to change the display, one to accept the option on display. Of course, the telephone would be slightly more expensive. The tradeoff is cost versus usability.
+
+Usability research evaluating touchscreen efficacy for both IBSVI and sighted users
+still needs to be covered in the literature. I completely agree that the lack of accessibility
+for IBSVI is a major issue with touchscreen devices. However, providing more tactile
+feedback can increase usability not just for IBSVI, but also for sighted users. In Donald
+Norman’s book, The Design of Everyday Things, Norman touches on this principle of
+feedback. Norman describes how telephones in the past “were designed with much
+more care and concern for the user” as “the push buttons were designed to give an
+appropriate feel—tactile feedback.” [20] Norman asserts that modern telephone systems
+are “so difficult to use” since they have “more features and less feedback.” [20] As tactile
+feedback is significant for IBSVI and sighted users alike, research should examine how
+both groups of users respond to proposed solutions. I think that this could possibly
+lead to a solution that becomes more widely accepted.
+
 \section{Usability Metric ``Forecast''}
+Predict to see an increase in learnability for sighted users and IBSVI alike.
+Should be more efficient for IBSVI but may be slightly less efficient for sighted users.
+They can no longer press a button quite as easily.
+Now that buttons are physical, they will have to lift their finger slightly more and press down slightly harder.
+This should not take any major hits with efficiency.
+Rate of errors should be less for both parties.
 \clearpage

### 2014-11-25 dream-design/dream-design.tex
add abstract

@@
 \begin{abstract}
-This investigation purposes a ``dream'' user interface for touch screen devices. 
+This investigation presents a ``dream'' user interface design for touch screen mobile devices.
+This design aims to tackle the usability issue of insufficient feedback as well as provide an interface that is user-friendly for individuals with visual impairments and individuals without visual impairments alike.
+Of the five prevalent usability metrics within the field of Interaction Design, the three metrics of learnability, efficiency, and errors were particularly significant in the creation and development of this interface design \cite{Nielsen:1993:UE:529793}.
 \end{abstract}
@@
 
-Many mobile devices utilize touch screens instead of physical keyboards or buttons.
+Many of these mobile devices today utilize touch screens instead of physical keyboards or buttons.
 This allows for a larger display without increasing the size of the device.
@@
 The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users.
-A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) as these users are only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
+A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the interface or the screen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
 
-Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses some of these usability issues.
+Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses this usability issue.
 
@@
 
-In this investigation, I propose a ``dream'' interface design for touchscreen mobile devices that focuses on providing the user with as much as feedback as possible.
+In this investigation, I present a ``dream'' interface design for touchscreen mobile devices that focuses on providing the user with as much as feedback as possible.
 This feedback is designed in a way that aims to be helpful for users with visual impairments and users without visual impairments alike.
@@
 In 2013, Touch Revolution, the largest-volume glass projected capacitive multi-touch screen manufacturer in the world, announced a partnership with Tactus Technology \cite{TactusAvailability}.
-The technologies are being combined and manufactured into consumer devices today \cite{TactusAvailability}.
+The technologies from the two companies are being combined and manufactured into consumer devices today \cite{TactusAvailability}.
 Not only is this technology becoming more and more available, but it is becoming more customizable as well.
@@
 
-My ``dream'' user interface design makes use of the tactile feedback as well as the flexibility that this technology from Tactus provides.
+My ``dream'' user interface design makes use of the varied tactile feedback that this technology from Tactus can provide.
 
@@
 The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike.
-The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the tactile buttons, and the auditory feedback.
+The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the ``intelligent'' tactile buttons, and the auditory feedback.
 
 \subsection{Optional Physical Grid}
-The design incorporates Tactus technology to provide an optional tactile grid for the device (see Figure~\ref{wireframe-grid}).
-The grid layout would consist of a set of very thin, three-dimensional horizontal and vertical lines, which could be enabled by the user in the device settings .
-Both sets of lines are arranged with equal space between each line.
-The main purpose of this grid is to help IBSVI formulate a mental reference for location awareness to maintain place on the screen.
+The design incorporates Tactus technology in order to provide an optional tactile grid for the device (see Figure~\ref{wireframe-grid}).
+The grid layout would consist of a set of very thin, three-dimensional horizontal and vertical lines.
+If the grid is enabled by the user, the set of horizontal and vertical lines appear on the screen.
+If the grid is disabled by the user, the lines recede.
+The horizontal and vertical lines are arranged with equal space between each line.
+The main purpose of this grid is to help IBSVI formulate a mental reference for location awareness.
+The fact that the grid can easily be disabled makes this feature unobtrusive for users who do not find benefit from it.
 
@@
 \includegraphics[width=2in]{wireframe-grid.png} 
-\caption{A simple wireframe demonstrating the grid layout idea.
-Each line in the figure would actually be a three-dimensional line that appears on the screen with the use of microfluidic technology from Tactus.}
+\caption{A simple wireframe demonstrating the concept of the grid layout.
+Each line within the screen would actually be a physical line that appears on the screen with the use of Tactus's microfluidic technology.}
 \label{wireframe-grid}

### 2014-11-25 dream-design/dream-design.tex
editing

@@
 The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike.
-The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the ``intelligent'' tactile buttons, and the auditory feedback.
+The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the ``intelligent'' tactile buttons, and the aural feedback.
 
@@
 \caption{A simple wireframe demonstrating the concept of the grid layout.
-Each line within the screen would actually be a physical line that appears on the screen with the use of Tactus's microfluidic technology.}
+Each line on the screen of the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
 \label{wireframe-grid}
@@
 \subsection{``Intelligent'' Tactile Buttons}
-The design also includes tactile buttons that rise up and recede as needed (see Figure~\ref{buttons}).
-This aspect of the design would also be made possible with the use of technology from Tactus.
-Ideally, the mobile device would be ``intelligent'' about which buttons rose up on the flat screen depending on input from the user.
+The design also includes tactile buttons that would rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}).
+Like with the optional physical grid, this aspect of the interface design would also make use of Tactus's microfluidic technology.
+
+Ideally, only buttons applicable to the input of the user or current state of the device would appear.
+For example, in an instance where the user is provided with a binary yes/no choice, only two tactile buttons should appear---one for yes and one for no.
+Similarly, if all possible buttons were applicable for a given situation, then all possible buttons should appear.
 
@@
 
-In instances when less options are available to the user, less buttons should rise up from the screen.
-Similarly, when more options are available to the user, more buttons should rise up from the screen.
-This design aspect should provide more clarity with respect to control flow and what the user is capable of doing at any moment.
+\subsection{Aural Feedback}
+The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide aural feedback.
+Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} the application is.
 
-\subsection{Auditory Feedback}
-The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide auditory feedback.
-This feedback can inform the user of \textit{what} is at the location on the screen whereas the tactile buttons and grid helps the user know \textit{where} it is.
+This portion of the interface design should not be too difficult to implement as there already exists technology that provides this kind of aural feedback.
+For example, the Apple operating system, iOS, has the VoiceOver screen reader that provides a spoken description of everything happening on the screen \cite{VoiceOver}.
+The Android OS has a similar service called TalkBack, which adds spoken and audible feedback to the device \cite{TalkBack}.
 

### 2014-11-25 dream-design/dream-design.tex
more edits

@@
 \begin{abstract}
-This investigation presents a ``dream'' user interface design for touch screen mobile devices.
-This design aims to tackle the usability issue of insufficient feedback as well as provide an interface that is user-friendly for individuals with visual impairments and individuals without visual impairments alike.
+This investigation presents a ``dream'' user interface design for touchscreen mobile devices.
+This design aims to tackle the touchscreen usability issue of insufficient feedback as well as provide an interface that is user-friendly for individuals with visual impairments and individuals without visual impairments alike.
 Of the five prevalent usability metrics within the field of Interaction Design, the three metrics of learnability, efficiency, and errors were particularly significant in the creation and development of this interface design \cite{Nielsen:1993:UE:529793}.
@@
 
-Many of these mobile devices today utilize touch screens instead of physical keyboards or buttons.
+Many of these mobile devices today utilize touchscreens instead of physical keyboards or buttons.
 This allows for a larger display without increasing the size of the device.
@@
 
-Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen mobile devices.
+Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen devices.
 \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}.
@@
 The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users.
-A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the interface or the screen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
+A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
 
-Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses this usability issue.
+Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses the usability issue of insufficient feedback.
 
@@
 My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}.
-Tactus Technology, a company based in Fremont, California, creates real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus1}).
+Tactus Technology, a company based in Fremont, California, creates technology that can reshape the surface of touchscreens on demand.
+The technology allows real physical buttons to dynamically appear and disappear into a flat touchscreen (see Figure~\ref{tactus1}).
 Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.
 
-Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition as well as several awards \cite{CNN, I-Zone, PCMag, Wired}.
+Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition \cite{CNN, I-Zone, PCMag, Wired}.
 Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}.
@@
 The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike.
-The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the ``intelligent'' tactile buttons, and the aural feedback.
+The three main features of this design are the optional physical grid, the ``intelligent'' tactile buttons, and the aural feedback.
 
@@
 \caption{A simple wireframe demonstrating the concept of the grid layout.
-Each line on the screen of the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
+Each line in the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
 \label{wireframe-grid}
@@
 \subsection{``Intelligent'' Tactile Buttons}
-The design also includes tactile buttons that would rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}).
-Like with the optional physical grid, this aspect of the interface design would also make use of Tactus's microfluidic technology.
-
-Ideally, only buttons applicable to the input of the user or current state of the device would appear.
+The interface design also includes tactile buttons that use Tactus's microfluidic technology to rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}).
+Ideally, only buttons applicable to the input of the user or current state of the device would appear at any given time.
 For example, in an instance where the user is provided with a binary yes/no choice, only two tactile buttons should appear---one for yes and one for no.
@@
 \subsection{Aural Feedback}
+\label{aural-feedback}
 The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide aural feedback.
-Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} the application is.
+Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} an application is.
 
-This portion of the interface design should not be too difficult to implement as there already exists technology that provides this kind of aural feedback.
+This type of feedback should not be too difficult to implement as there already exists technology that provides similar aural feedback.
 For example, the Apple operating system, iOS, has the VoiceOver screen reader that provides a spoken description of everything happening on the screen \cite{VoiceOver}.
@@
 There exist several usage scenarios for this interface design.
-Sighted users employ vision to tell the user what and where to touch, but this interface lacks accessibility for IBSVI because it lacks physical landmarks and tangible buttons.
-A typical solution is to provide voice over to provide this information, but there is nothing to guide IBSVI to that location in the first place, and there are no landmarks other than dead-reckoning and pure spatial memory to support navigation for the IBSVI.
-For example, the Apple operating system, iOS, has VoiceOver function, and the Android OS has a Talkbalk function.
-iOS VoiceOver reads aloud the icon, the user should double tap on that location without the benefit of vision to ensure that the finger does not stray from the touch point.
+One such scenario is when an Individual with Blindness of Servere Visual Impairment (IBSVI) attempts to locate an application on a touchscreen mobile device.
+Sighted users employ vision to figure out what and where to touch.
+Since IBSVI cannot do this, typical touchscreen devices lack accessibility for IBSVI.
+As previously touched on in Section~\ref{aural-feedback}, a typical solution to this problem is to use voice over technology to provide information about what is on the screen at any given time.
+However, there is nothing to guide IBSVI to the location of a specific application in the first place.
+Since there are no landmarks to support navigation for the IBSVI, the user must use dead-reckoning and trial-and-error methods in order to find applications specified by the voice over technology.
 
-The problem with such accessibility modes is that the user can develop a spatial mental model for the interface or screed only through dead reckoning to find VoiceOver locations, in the absence of any landmark other than the boundary of the device.
-The optional tactile grid as well as the three-dimensional buttons that appear with this design solves this problem.
-The additional tactile feedback and physical landmarks can guide IBSVI to interact efficiently and effectively with the touch device and its voice over functionality.
+The design of this interface addresses this problem as the tactile grid and the three-dimensional buttons provide additional landmarks for IBSVI.
+These landmarks aim to allow IBSVI interact more efficiently and effectively with the touch device and its voice over functionality.
+Ideally, the voice over functionality would be able to tell the user the exact cell in the grid in which the desired application is located.
+The user could then find the application much more easily.
 
-Another use case for this interface pertains specifically to the ``intelligent'' tactile buttons.
-Users often may get overwhelmed by the number of options that are always available to them.
-It also may be frustrating for users to try to perform a task and not be able to.
-To examine a specific example, let's say that the user cannot take any more photos on their phone since they are out of memory.
-In this case, the interface would not even raise the button used to capture photos.
-This is a good indicator to the user that they need to perform some kind of an action before proceeding.
+Another use case for this interface is when some functionality of an application is disabled.
+Say, for example, that a user cannot take any more photos on their mobile device because they are out of memory.
+Some ``intelligent'' tactile picture-taking button could recede when this happens to communicate to the user that photo taking functionality is disabled.
+More generally, the``intelligent'' tactile buttons can be useful in indicating to the user what choices they have at any given time.
 

### 2014-11-26 dream-design/dream-design.tex
replace button picture
more edits

@@
 
-Many of these mobile devices today utilize touchscreens instead of physical keyboards or buttons.
+Many mobile devices today utilize touchscreens instead of physical keyboards or buttons.
 This allows for a larger display without increasing the size of the device.
@@
 
-Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen devices.
+However, the literature suggests that overall feedback is still insufficient and remains a major usability issue with touchscreen devices.
 \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}.
@@
 \centering
-\includegraphics[width=6in]{buttons.jpg}
+\includegraphics[width=6in]{buttons.png}
 \caption{A simple representation demonstrating the idea of different subsets of buttons rising up on the same surface at different times.}
@@
 \section{Rationale}
-This layout is designed so that IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
+The driving force behind the creation and development of this interface design is feedback.
+Feedback is considered a critical concept in the field of interaction design.
+Donald Norman, the author of \textit{The Design of Everyday Things}, repeatedly articulates the importance of showing the effect of an action \cite{Norman02}.
+Without feedback, the user is always wondering whether anything has happened.
+In my ``dream'' interface design, the tactile feedback from the physical buttons serve to help the user understand when an action has been performed.
+Furthermore, the additional aural feedback of the design also assists in communicating the current state of the device to the user.
+
+Another interaction design concept that Norman discusses is how systems with more options and less feedback can be very difficult to use \cite{Norman02}.
+My design includes ``intelligent'' physical buttons so that the device provides the user with the minimum number of options any given time.
+Since the buttons will only appear in situations when they are needed, it narrows the options of the user.
+This ultimately decreases the cognitive load of the user as they have less buttons from which to choose.
+
+This layout is designed so that IBSVI can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
 The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.
@@
 In design, it is important to show the effect of an action.
-Without feedback, one is always wondering whether anything has hap- pened.
+Without feedback, one is always wondering whether anything has happened.
 Maybe the button wasn't pushed hard enough; maybe the machine has stopped working; maybe it is doing the wrong thing.
-Without feed- back, we turn equipment off at improper times or restart unnecessarily, losing all our recent work.
+Without feedback, we turn equipment off at improper times or restart unnecessarily, losing all our recent work.
 Or we repeat the command and end up having the operation done twice, often to our detriment.
-Feedback is critical (the design of everyday things).
-
-In the good old days of the telephone, before the American tele- phone system was divided among competing companies, before tele- phones were fancy and had so many features, telephones were de- signed with much more care and concern for the user.
-Designers at the Bell Telephone Laboratories worried a lot about feedback.
-The push buttons were designed to give an appropriate feel—tactile feedback.
-When a button was pushed, a tone was fed back into the earpiece so the user could tell that the button had been properly pushed.
-When the phone call was being connected, clicks, tones, and other noises gave the user feedback about the progress of the call.
-And the speaker's voice was always fed back to the earpiece in a carefully controlled amount, because the auditory feedback (called "sidetone") helped the person regulate how loudly to talk.
-All this has changed.
-We now have tele- phones that are much more powerful and often cheaper than those that existed just a few years ago—more function for less money.
-To be fair, these new designs are pushing hard on the paradox of technology: added functionality generally comes along at the price of added com- plexity.
-But that does not justify backward progress.
-Why are the modern telephone systems so difficult to learn and to use?
-Basically, the problem is that the systems have more features and less feedback.
-Suppose all telephones had a small display screen, not unlike the ones on small, inexpensive calculators.
-The display could be used to present, upon the push of a button, a brief menu of all the features of the telephone, one by one.
-When the desired one was encountered, the user would push another button to indicate that it should be invoked.
-If further action was required, the display could tell the person what to do.
-The display could even be auditory, with speech instead of a visual display.
-Only two buttons need be added to the
-telephone: one to change the display, one to accept the option on display. Of course, the telephone would be slightly more expensive. The tradeoff is cost versus usability.
+Feedback is critical.
 

### 2014-11-26 dream-design/dream-design.tex
final touches

@@
 
-However, the literature suggests that overall feedback is still insufficient and remains a major usability issue with touchscreen devices.
+Although these alternative aural and haptic signals are provided for the user, the literature suggests that overall feedback is still insufficient and remains a major usability issue with touchscreen devices.
 \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}.
-A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}.
-The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users.
-A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
+A study by Hussain Tinwala and Scott MacKenzie submits that the lack of tactile feedback requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}.
+The lack of tactile feedback not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users.
+A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI).
+These users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.
 
@@
 
-My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}.
+This ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}.
 Tactus Technology, a company based in Fremont, California, creates technology that can reshape the surface of touchscreens on demand.
@@
 The technologies from the two companies are being combined and manufactured into consumer devices today \cite{TactusAvailability}.
-Not only is this technology becoming more and more available, but it is becoming more customizable as well.
+Not only is this technology becoming more  available, but it is becoming more customizable as well.
 Companies will soon be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote \cite{CNN}.
@@
 
-My ``dream'' user interface design makes use of the varied tactile feedback that this technology from Tactus can provide.
+This ``dream'' user interface design makes use of the varied tactile feedback that the technology from Tactus can provide.
 
@@
 The grid layout would consist of a set of very thin, three-dimensional horizontal and vertical lines.
-If the grid is enabled by the user, the set of horizontal and vertical lines appear on the screen.
+If the grid is enabled by the user, the set of horizontal and vertical lines appears on the screen.
 If the grid is disabled by the user, the lines recede.
@@
 \caption{A simple wireframe demonstrating the concept of the grid layout.
-Each line in the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
+Each line in the figure would actually be a physical line that appears on the device's screen with the use of Tactus' microfluidic technology.}
 \label{wireframe-grid}
@@
 \subsection{``Intelligent'' Tactile Buttons}
-The interface design also includes tactile buttons that use Tactus's microfluidic technology to rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}).
+The interface design also includes tactile buttons that utilize Tactus' microfluidic technology in order to rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}).
 Ideally, only buttons applicable to the input of the user or current state of the device would appear at any given time.
@@
 \label{aural-feedback}
-The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide aural feedback.
+The design of this interface was created in such a way that the tactile aspects can operate in conjunction with other technologies that provide aural feedback.
 Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} an application is.
@@
 There exist several usage scenarios for this interface design.
-One such scenario is when an Individual with Blindness of Servere Visual Impairment (IBSVI) attempts to locate an application on a touchscreen mobile device.
+One such scenario is when an Individual with Blindness of Severe Visual Impairment (IBSVI) attempts to locate an application on a touchscreen mobile device.
 Sighted users employ vision to figure out what and where to touch.
@@
 Donald Norman, the author of \textit{The Design of Everyday Things}, repeatedly articulates the importance of showing the effect of an action \cite{Norman02}.
-Without feedback, the user is always wondering whether anything has happened.
-In my ``dream'' interface design, the tactile feedback from the physical buttons serve to help the user understand when an action has been performed.
+Without feedback, the user is always wondering whether or not anything has happened.
+In this ``dream'' interface design, the tactile feedback from the physical buttons serves to help the user understand when an action has been performed.
 Furthermore, the additional aural feedback of the design also assists in communicating the current state of the device to the user.
@@
 My design includes ``intelligent'' physical buttons so that the device provides the user with the minimum number of options any given time.
-Since the buttons will only appear in situations when they are needed, it narrows the options of the user.
-This ultimately decreases the cognitive load of the user as they have less buttons from which to choose.
+Since the buttons will only appear in situations when they are needed, it narrows the options for the user.
+This ultimately decreases the cognitive load on the user as they have less buttons from which to choose.
 
-This layout is designed so that IBSVI can engage their spatial cognition, perception and sensing resources while interacting with touch screens.
-The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.
-
-Feedback.
-In design, it is important to show the effect of an action.
-Without feedback, one is always wondering whether anything has happened.
-Maybe the button wasn't pushed hard enough; maybe the machine has stopped working; maybe it is doing the wrong thing.
-Without feedback, we turn equipment off at improper times or restart unnecessarily, losing all our recent work.
-Or we repeat the command and end up having the operation done twice, often to our detriment.
-Feedback is critical.
-
-Usability research evaluating touchscreen efficacy for both IBSVI and sighted users
-still needs to be covered in the literature. I completely agree that the lack of accessibility
-for IBSVI is a major issue with touchscreen devices. However, providing more tactile
-feedback can increase usability not just for IBSVI, but also for sighted users. In Donald
-Norman’s book, The Design of Everyday Things, Norman touches on this principle of
-feedback. Norman describes how telephones in the past “were designed with much
-more care and concern for the user” as “the push buttons were designed to give an
-appropriate feel—tactile feedback.” [20] Norman asserts that modern telephone systems
-are “so difficult to use” since they have “more features and less feedback.” [20] As tactile
-feedback is significant for IBSVI and sighted users alike, research should examine how
-both groups of users respond to proposed solutions. I think that this could possibly
-lead to a solution that becomes more widely accepted.
+This design was also created in a way that allows IBSVI to engage their spatial cognition, perception and sensing resources while interacting with touchscreen mobile devices.
+A previous study from Virginia Polytechnic Institute and State University demonstrates how touchscreen devices with some kind of a physical overlay typically improve learnability and efficiency while decreasing error rates for IBSVI \cite{El-Glaly:2013:TTF:2460625.2460665}.
+The optional grid of this ``dream'' interface is very similar to the physical overlay from the study, and should thus improve usability metrics for IBSVI with touchscreen phones.
 
 \section{Usability Metric ``Forecast''}
-Predict to see an increase in learnability for sighted users and IBSVI alike.
-Should be more efficient for IBSVI but may be slightly less efficient for sighted users.
-They can no longer press a button quite as easily.
-Now that buttons are physical, they will have to lift their finger slightly more and press down slightly harder.
-This should not take any major hits with efficiency.
-Rate of errors should be less for both parties.
+If implemented then tested, I would expect this interface to have better usability metrics with respect to learnability than conventional touchscreen mobile devices.
+More physical landmarks allow all users to engage their spatial cognition more fully when interacting with touchscreens.
+Furthermore, evidence in the past has demonstrated  that more tactile feedback can make a system more learnable \cite{El-Glaly:2013:TTF:2460625.2460665, conf/chi/KaneWL11}.
+
+Studies in the past have also suggested that more tactile feedback makes IBSVI more efficient when using the system \cite{El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}.
+This interface might, however, be slightly less efficient for users who do not have visual impairments.
+These users will have to lift their fingers more and press down significantly harder with the tactile buttons.
+I would expect this difference in efficiency to be negligible though.
+
+I would also expect all users to be more perform less errors as well as be more satisfied with this ``dream'' interface.
+As tactile buttons are harder to press, there would be less likelihood that one would be pressed accidentally.
+Furthermore, studies have shown that most users enjoy tactile feedback \cite{Freeman:2014:TFA:2663204.2663280}.
+
 \clearpage

### 2014-11-26 dream-design/dream-design.tex
small fixes

@@
 
-I would also expect all users to be more perform less errors as well as be more satisfied with this ``dream'' interface.
-As tactile buttons are harder to press, there would be less likelihood that one would be pressed accidentally.
-Furthermore, studies have shown that most users enjoy tactile feedback \cite{Freeman:2014:TFA:2663204.2663280}.
+I would also expect all users to perform less errors as well as be more satisfied with this ``dream'' interface.
+As tactile buttons are harder to press, there would be less likelihood that a button would be pressed accidentally.
+Furthermore, studies have shown that most users prefer tactile feedback \cite{Freeman:2014:TFA:2663204.2663280}.
 

### 2014-11-27 direct-manipulation/boxes-touch.js
add squares

@@ -1,0 +1,98 @@
+var BoxesTouch = {
+    /**
+     * Sets up the given jQuery collection as the drawing area(s).
+     */
+    setDrawingArea: function(jQueryElements) {
+        // Set up any pre-existing box elements for touch behavior.
+        jQueryElements
+            .addClass("drawing-area")
+
+        // Event handler setup must be low-level because jQuery
+        // doesn't relay touch-specific event properties.
+        .each(function(index, element) {
+            element.addEventListener("touchmove", BoxesTouch.trackDrag, false);
+            element.addEventListener("touchend", BoxesTouch.endDrag, false);
+        })
+
+        .find("div.box").each(function(index, element) {
+            element.addEventListener("touchstart", BoxesTouch.startMove, false);
+            element.addEventListener("touchend", BoxesTouch.unhighlight, false);
+        });
+        $('#create-square').bind('click', function() {
+            var newSquare = $('#boxy-box').clone();
+            $(newSquare).css('top', '50px');
+            $("#drawing-area").append(newSquare);
+            jQueryElements.find("div.box").each(function(index, element) {
+                element.addEventListener("touchstart", BoxesTouch.startMove, false);
+                element.addEventListener("touchend", BoxesTouch.unhighlight, false);
+            });
+
+        });
+
+    },
+
+    /**
+     * Tracks a box as it is rubberbanded or moved across the drawing area.
+     */
+    trackDrag: function(event) {
+        $.each(event.changedTouches, function(index, touch) {
+            // Don't bother if we aren't tracking anything.
+            if (touch.target.movingBox) {
+                // Reposition the object.
+                touch.target.movingBox.offset({
+                    left: touch.pageX - touch.target.deltaX,
+                    top: touch.pageY - touch.target.deltaY
+                });
+            }
+        });
+
+        // Don't do any touch scrolling.
+        event.preventDefault();
+    },
+
+
+    /**
+     * Concludes a drawing or moving sequence.
+     */
+    endDrag: function(event) {
+        $.each(event.changedTouches, function(index, touch) {
+            if (touch.target.movingBox) {
+                // Change state to "not-moving-anything" by clearing out
+                // touch.target.movingBox.
+                touch.target.movingBox = null;
+            }
+        });
+    },
+
+    /**
+     * Indicates that an element is unhighlighted.
+     */
+    unhighlight: function() {
+        $(this).removeClass("box-highlight");
+    },
+
+    /**
+     * Begins a box move sequence.
+     */
[…]

### 2014-11-27 direct-manipulation/boxes-touch.js
remove functionality

@@ -57,6 +57,10 @@ var BoxesTouch = {
     endDrag: function(event) {
         $.each(event.changedTouches, function(index, touch) {
             if (touch.target.movingBox) {
+                if (touch.pageX < 50 && touch.pageY < 50){
+                    console.log("yo");
+                    touch.target.movingBox.remove();
+                }
                 // Change state to "not-moving-anything" by clearing out
                 // touch.target.movingBox.
                 touch.target.movingBox = null;

### 2014-11-28 direct-manipulation/boxes-touch.js
initial work on dpad widget

@@ -22,7 +22,7 @@ var BoxesTouch = {
             var newSquare = $('#boxy-box').clone();
             $(newSquare).css('top', '50px');
             $("#drawing-area").append(newSquare);
-            jQueryElements.find("div.box").each(function(index, element) {
+            jQueryElements.find("#" + "boxy-box").each(function(index, element) {
                 element.addEventListener("touchstart", BoxesTouch.startMove, false);
                 element.addEventListener("touchend", BoxesTouch.unhighlight, false);
             });

### 2014-11-28 widget-from-scratch/js/directionalpad.js

@@ -1,0 +1,312 @@
+(function(root, factory) {
+  if (typeof exports === 'object') {
+     module.exports = factory(require('jquery'));
+  } else if (typeof root.define === 'function' && root.define.amd) {
+    define(['jquery'], factory);
+  } else {
+    factory(root.jQuery || root.Zepto);
+  }
+}(this,
+function($) {
+  var CellFactory = {
+    createFrom: function($cell) {
+      var position = $cell.position();
+
+      return {
+        leftPosition: $cell.position().left,
+        topPosition: $cell.position().top,
+        $cell: $cell
+      };
+    }
+  };
+
+
+  var CellTable = function($nodes) {
+    this.table = this.buildTable($nodes);
+    this.rows = this.buildRows();
+    this.columns = this.buildColumns();
+  };
+
+  CellTable.prototype = {
+    buildTable: function($nodes) {
+      return $nodes.map(function() {
+        return CellFactory.createFrom($(this));
+      });
+    },
+
+    buildColumns: function() {
+      var columns = {},
+          self = this;
+
+      $.each(this.table, function(index, cell) {
+        columns[cell.leftPosition] = self.getColumnElements(cell);
+      });
+
+      return columns;
+    },
+
+    buildRows: function() {
+      var rows = {},
+          self = this;
+
+      $.each(this.table, function(i, cell) {
+        rows[cell.topPosition] = self.getRowElements(cell);
+      });
+
+      return rows;
+    },
+
+    getRowElements: function(compareCell) {
+      var self = this;
+
+      return $.map(this.table, function(cell) {
+        if (self.isSameRow(cell, compareCell)) {
+          return cell;
+        }
+
+        return null;
+      });
+    },
+
+    getColumnElements: function(compareCell) {
+      var self = this;
+
+      return $.map(this.table, function(cell) {
+        if (self.isSameColumn(cell, compareCell)) {
+          return cell;
+        }
+
+        return null;
+      });
+    },
+
+    getCurrent: function($cell) {
+      var cell = CellFactory.createFrom($cell);
+
+      return this.findPosition(
+        this.getCell(cell)
+      );
+    },
+
+    isSameColumn: function(cell, compareCell) {
+      if (!compareCell) {
+        throw 'cell';
+      }
+
+      return cell.leftPosition === compareCell.leftPosition;
+    },
+
+    isSameRow: function(cell, compareCell) {
[…]

### 2014-11-28 widget-from-scratch/js/jquery.directionalpad.js
more work with widget

@@ -1,17 +1,7 @@
-(function(root, factory) {
-  if (typeof exports === 'object') {
-     module.exports = factory(require('jquery'));
-  } else if (typeof root.define === 'function' && root.define.amd) {
-    define(['jquery'], factory);
-  } else {
-    factory(root.jQuery || root.Zepto);
-  }
-}(this,
-function($) {
+(function ( $ ) {
   var CellFactory = {
     createFrom: function($cell) {
       var position = $cell.position();
-
       return {
         leftPosition: $cell.position().left,
         topPosition: $cell.position().top,
@@ -20,88 +10,77 @@ function($) {
     }
   };
 
-
   var CellTable = function($nodes) {
-    this.table = this.buildTable($nodes);
-    this.rows = this.buildRows();
-    this.columns = this.buildColumns();
+    this.table = this.makeTable($nodes);
+    this.rows = this.makeRows();
+    this.columns = this.makeColumns();
   };
 
   CellTable.prototype = {
-    buildTable: function($nodes) {
+    makeTable: function($nodes) {
       return $nodes.map(function() {
         return CellFactory.createFrom($(this));
       });
     },
 
-    buildColumns: function() {
+    makeColumns: function() {
       var columns = {},
           self = this;
-
       $.each(this.table, function(index, cell) {
         columns[cell.leftPosition] = self.getColumnElements(cell);
       });
-
       return columns;
     },
 
-    buildRows: function() {
+    makeRows: function() {
       var rows = {},
           self = this;
-
       $.each(this.table, function(i, cell) {
         rows[cell.topPosition] = self.getRowElements(cell);
       });
-
       return rows;
     },
 
     getRowElements: function(compareCell) {
       var self = this;
-
       return $.map(this.table, function(cell) {
-        if (self.isSameRow(cell, compareCell)) {
+        if (self.inSameRow(cell, compareCell)) {
           return cell;
         }
-
         return null;
       });
     },
 
     getColumnElements: function(compareCell) {
       var self = this;
-
       return $.map(this.table, function(cell) {
-        if (self.isSameColumn(cell, compareCell)) {
+        if (self.inSameCol(cell, compareCell)) {
           return cell;
         }
-
         return null;
       });
     },
 
     getCurrent: function($cell) {
       var cell = CellFactory.createFrom($cell);
-
[…]

### 2014-11-28 widget-from-scratch/js/jquery.directionalpad.js
small fixes

@@ -1,4 +1,4 @@
-(function ( $ ) {
+(function($) {
   var CellFactory = {
     createFrom: function($cell) {
       var position = $cell.position();
@@ -10,13 +10,13 @@
     }
   };
 
-  var CellTable = function($nodes) {
+  var DPadTable = function($nodes) {
     this.table = this.makeTable($nodes);
     this.rows = this.makeRows();
     this.columns = this.makeColumns();
   };
 
-  CellTable.prototype = {
+  DPadTable.prototype = {
     makeTable: function($nodes) {
       return $nodes.map(function() {
         return CellFactory.createFrom($(this));
@@ -168,14 +168,14 @@
         next = cells[info.firstIndex ? 0 : cells.length - 1];
       }
 
-      this.setActive(next.$cell);
+      this.setToActive(next.$cell);
     },
 
     down: function($cell, cellIndex) {
       this.move({
         cellPosition: ($cell).position().left,
         index: cellIndex.rowIndex + 1,
-        cells: this.cellTable.columns,
+        cells: this.table.columns,
         firstIndex: true
       });
     },
@@ -184,7 +184,7 @@
       this.move({
         cellPosition: ($cell).position().left,
         index: cellIndex.rowIndex - 1,
-        cells: this.cellTable.columns
+        cells: this.table.columns
       });
     },
 
@@ -192,7 +192,7 @@
       this.move({
         cellPosition: ($cell).position().top,
         index: cellIndex.colIndex - 1,
-        cells: this.cellTable.rows
+        cells: this.table.rows
       });
     },
 
@@ -200,7 +200,7 @@
       this.move({
         cellPosition: ($cell).position().top,
         index: cellIndex.colIndex + 1,
-        cells: this.cellTable.rows,
+        cells: this.table.rows,
         firstIndex: true
       });
     },
@@ -216,13 +216,13 @@
 
       // find the current selected cell
       var $selected = this.$parent.find('.' + this.options.activeClass);
-      var cell = this.cellTable.getCurrent($selected);
+      var cell = this.table.getCurrent($selected);
 
       // apply the function that corresponds to the arrow key pressed
       return this[fn].apply(this, [$selected, cell, event]);
     },
 
-    setActive: function($cell) {
+    setToActive: function($cell) {
       this.$nodes.removeClass(this.options.activeClass);
       $cell.addClass(this.options.activeClass);
     },
@@ -233,21 +233,15 @@
 
[…]
